### 2017-04-20 main.tex
First draft

\documentclass[a4paper,11pt]{article}

% Wider margins (compared to default)
\usepackage[margin=2.5cm]{geometry}

% English support (typography and hyphenation)
\usepackage[english]{babel}

% Unicode encoding
\usepackage[utf8]{inputenc}

% Better default font (Libertine and Inconsolata)
\usepackage[ttscale=.875]{libertine}
\usepackage[scaled=0.96]{zi4}

% Biber
\usepackage[backend=biber, style=numeric, natbib=true]{biblatex}
\bibliography{library.bib}

% Graphics
\usepackage{graphicx}

% Code Listing
\usepackage[framemethod=tikz]{mdframed}
\usepackage{listings}
\surroundwithmdframed[
%  hidealllines=true,
  linewidth=0.25pt,
  linecolor=black!50,
  backgroundcolor=black!4,
  innerleftmargin=8pt,
  innertopmargin=3pt,
  innerbottommargin=3pt]{lstlisting}
\lstset{ %
  basicstyle=\small\tt\linespread{0.75},
  language=Python,
  commentstyle=\color{gray},
  columns=fullflexible,
}

% Hyperref
\usepackage{xcolor}
\definecolor{blendedblue}{rgb}{0.2, 0.2, 0.6}
\definecolor{blendedred}{rgb}{0.8, 0.2, 0.2}
\usepackage[bookmarks=true,
            breaklinks=true,
            pdfborder={0 0 0},
            citecolor=blendedblue,
            colorlinks=true,
            linkcolor=blendedblue,
            urlcolor=blendedblue,
            citecolor=blendedblue,
            linktocpage=false,
            hyperindex=true,
            linkbordercolor=white]{hyperref}
\usepackage{hyperref}
\hypersetup{colorlinks=true}

\title{The R Quintuplet (R$^5$)}
\author{Nicolas P. Rougier and Fabien C. Y. Benureau}
\date{\today}

\begin{document}
\maketitle

% -------------------------
\section*{Introduction (R$^{\mathbf 0}$)}

Scientific reproducibility and replicability issues have received increased attention in the past few years and computer science is no exception.
One could think that computer science would mostly be shielded from those issues. 
But precisely because it is easy to believe that if a program runs once (and gives expected results) it will do so forever, crucial steps to transform working code into meaningful scientific contributions are rarely undertaken; computer science is victim of reproducibility and replicability issues precisely because it seems impervious to them.

In fact, a program can fail in many different ways at many different levels for many different reasons.
Borrowing the terms coined by Ann Gobble, for your program to survive for some noticeable time, you have to make sure it is re-runnable, repeatable, reproducible, replicate and re-usable.\\

Let us illustrate this with a very simple example, a random walk written in Python:

\begin{lstlisting}
# A random walker (R0)
import numpy as np

# randomly choose ten instances of either 1 or -1. 
steps = np.random.choice([-1,+1], 10)
y = steps.cumsum() # compute the cumulative sum (the sum of all preceeding elements)
print y
\end{lstlisting}

Executed, this program would display 
\begin{lstlisting}
[-1  0 -1 -2 -1 -2 -3 -2 -1 -2] # with `steps` being [-1 1 -1 -1  1 -1 -1  1  1 -1]
\end{lstlisting}

What could go wrong with such a simple program?\\
\vfill
Well...
\vfill

% -------------------------
\clearpage
\section*{Re-runnable (R$^{\mathbf 1}$)}

Did you ever wonder if you would be able to run a program you wrote some time
ago? Let's consider a program you wrote yesterday. Most probably you can still
run it today, but would this still be true in a week, in a month, in a year or
in a decade? Probably you would legitimately answer positively for weeks
or even months if you're optimistic, but you may have doubts when considering
longer time periods such as years or even decades. Part of the problem is that
computer science is evolving at a very fast pace and you cannot know how the
system, the software and the libraries your program depends on will evolve.
If you want your program to have a tiny chance to be runnable ten years from
now, you better make it clear what are the dependencies in terms of systems,
software and libraries.\\
% TODO: cite Collberg 2014 study
\cite{Collberg:2016}

\begin{lstlisting}
# A random walker (R1)
# Tested with Python 3.6.0 / Numpy 1.12.0
import numpy as np

# Compute random walk for 10 steps
steps = 2*(np.random.uniform(-1, +1, 10) > 0) - 1
y = steps.cumsum()
print(y)
\end{lstlisting}

In our case, the R$^0$ version of our tiny walker uses the print {\em instruction} that is available in Python 2.x but has been deprecated in Python 3 in favor or a  print {\em function}.
This means this program cannot be run with Python 3 and needs to be fixed.
Since Python 2 will soon be discontinued, it may be best to target directly Python 3 and to use the print function (and as a matter of fact, it will work with Python 2).
Another potential problem is the numpy `choice` function that has been introduced only in numpy v1.7.
If for some reason, the installed version of numpy is older, the program will also fail.
One quick fix is to use a simpler way to generate the steps (even if it is less efficient).
Last and most important, we need to specify all the dependencies of our program.

\clearpage
\section*{Repeatable (R$^{\mathbf 2}$)}

After having taken care of specifying dependencies, you can legitimately hope to be able to run the program without too much trouble.
However this does not guarantee you will get the same results at each run.
For example, if you're using a (pseudo) random number generator, you have to take care of initializing the generator using a specific seed.
If not, you're doomed and you won't be able to get twice the same output. But still, there are actually many more pitfalls ahead.
For example, you can play with the seed value and find one that produce an interesting result that you will save in a file.
And the, you can continue to play with the seed in order get new results.
But what if you want to come back to one specific result?
How do you find the seed related to this specific output?
Again, if you did not foresee this potential problem, you're doomed.\\

Turning back to our random walker example, we can fix it quite easily by initializing the random number generator and produce a dictionary holding seed and result that allows us to reach the R² level.

\begin{lstlisting}
# A random walker (R2)
# Tested with Python 3.6 / Numpy 1.12.0
import numpy as np

# Random generator initialization
seed = 1
np.random.seed(seed)

# Compute random walk for 10 steps
steps = np.random.choice([-1,+1], 10)
y = steps.cumsum()
print(y)
\end{lstlisting}

\clearpage
\section*{Reproducible (R$^{\mathbf 3}$)}

% Adding license for diffusion
% Lorena Barba example (Science blog)

\begin{lstlisting}
# A random walker (R3)
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
# Release under the BSD 2 clauses license
# Tested with Python 3.6 / Numpy 1.12.0 using macOS 10.12.4 on a 64 bits architecture
import numpy as np

def walk(n):
    """ Random walk for n steps """

    steps = 2*(np.random.uniform(-1,+1,n) > 0) - 1
    return steps.cumsum()

if __name__ == '__main__'
    # Unit test
    np.random.seed(1)
    assert walk(10) == [-1, 0, -1, -2, -3, -4, -5, -6, -7, -6]

    # Random generator initialization
    seed = 1
    np.random.seed(seed)

    # Compute random walk for n steps
    y = walk(n=10)

    # Save and display results
    results = {"seed": seed, "y": y}
    print(results["y"])
\end{lstlisting}

\clearpage
\section*{Replicable (R$^{\mathbf 4}$)}

There are some discrepancies between the numpy and the pure Python version due to subtle differences in the random generator.
As explained in the \href{https://docs.python.org/3.6/library/random.html}{Python documentation}, {\em Python uses the Mersenne Twister as the core generator. It produces 53-bit precision floats
and has a period of 2**19937-1.
The underlying implementation in C is both fast and threadsafe.}
But the numpy random generator is also based on the same Mersenne Twister generator and the difference comes from the way seed is interpreted to initialize the generator.
Fortunately, Python gives access to the internal state and we can {\em fix} it to make it match the behavior of the numpy generator (which is the default).

\begin{lstlisting}
# A random walker program (R4)
# A replication of the random walker using pure Python
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
# Release under the BSD 2 clauses license
# Tested with Python 3.6 using macOS 10.12.4 on a 64 bits architecture
import random
from itertools import accumulate

def state(seed):
    """ Compute the Mersenne Twister keys from given seed """
    mt = [seed] * (624+1)
    mt[-1] = 624
    for i in range(1, 624):
        mt[i] = int(0xFFFFFFFF & (1812433253 * (mt[i-1]^mt[i-1]>>30)+i))
    return (3, tuple(mt), None)

def walk(n=10):
    """ Random walk for n steps """

    steps = [-1 if random.uniform(-1,+1) < 0 else +1 for i in range(n)]
    return list(accumulate(steps))

if __name__ == '__main__'
    # Unit test
    random.setstate(state(seed=1))
    assert walk(n=10) == [-1, 0, -1, -2, -3, -4, -5, -6, -7, -6]

    # random generator initialization
    seed = 1
    random.setstate(state(seed))

    # Compute random walk for n steps
    y = walk(n=10)

    # Save and display results
    results = {"seed": seed, "y": y}
    print(results["y"])
\end{lstlisting}

\clearpage
\section*{Reusable (R$^{\mathbf 5}$)}

\clearpage
\section*{Conclusion}

% underscore that compared to psy/bio/etc. the replication issues of CS are easily (reasonably) managed: good solutions that (mostly) works exist right now.

\renewcommand*{\bibfont}{\small}
\printbibliography[title=References]

\end{document}

### 2017-04-21 main.tex
Code updated

@@
 % Code Listing
+\usepackage{listings}
 \usepackage[framemethod=tikz]{mdframed}
-\usepackage{listings}
-\surroundwithmdframed[
-%  hidealllines=true,
-  linewidth=0.25pt,
-  linecolor=black!50,
-  backgroundcolor=black!4,
-  innerleftmargin=8pt,
-  innertopmargin=3pt,
-  innerbottommargin=3pt]{lstlisting}
+\makeatletter
+\def\mdf@@codeheading{Code Listings}
+\define@key{mdf}{title}{\def\mdf@@codeheading{#1}}
+\mdfdefinestyle{lstlisting}{%
+  backgroundcolor=black!2.5,
+  innertopmargin=2pt,
+  middlelinewidth=0.75pt,
+  outerlinewidth=9pt,
+  outerlinecolor=white,
+  innerleftmargin=10pt,
+  innerrightmargin=10pt,
+  leftmargin=0pt,
+  rightmargin=0pt,
+  rightline=false,
+  leftline=false,
+  bottomline=false,
+  skipabove=\topskip,
+  skipbelow=\topskip,
+  roundcorner=0pt,
+  singleextra={
+    \node[text=black, % fill=white, draw,
+          anchor=south west, yshift=-0.25pt, xshift=5pt,
+          font=\sc\footnotesize] at (O|-P) {\csname mdf@@codeheading\endcsname};},
+  firstextra={
+    \node[text=black, % fill=white, draw,
+          anchor=south west, yshift=-0.5pt, xshift=5pt,
+          font=\sc\footnotesize] at (O|-P) {\csname mdf@@codeheading\endcsname};}
+}
+
 \lstset{ %
@@
 }
+
+\lstnewenvironment{code}[2][]{%
+  \lstset{#1}%
+  \mdframed[style=lstlisting,title={#2}]%
+}{\endmdframed}
+
+%\usepackage{listings}
+%\surroundwithmdframed[
+%%  hidealllines=true,
+%  linewidth=0.25pt,
+%  linecolor=black!50,
+%  backgroundcolor=black!4,
+%  innerleftmargin=8pt,
+%  innertopmargin=3pt,
+%  innerbottommargin=3pt]{lstlisting}
+%\lstset{ %
+%  basicstyle=\small\tt\linespread{0.75},
+%  language=Python,
+%  commentstyle=\color{gray},
+%  columns=fullflexible,
+%}
 
@@
 One could think that computer science would mostly be shielded from those issues. 
-But precisely because it is easy to believe that if a program runs once (and gives expected results) it will do so forever, crucial steps to transform working code into meaningful scientific contributions are rarely undertaken; computer science is victim of reproducibility and replicability issues precisely because it seems impervious to them.
+But precisely because it is easy to believe that if a program runs once (and gives expected results) it will do so forever, crucial steps to transform working code into meaningful scientific contributions are rarely undertaken; computer science is plagued by replication problems precisely because it seems impervious to them.\\
 
 In fact, a program can fail in many different ways at many different levels for many different reasons.
-Borrowing the terms coined by Ann Gobble, for your program to survive for some noticeable time, you have to make sure it is re-runnable, repeatable, reproducible, replicate and re-usable.\\
+Borrowing the terms coined by Carole Anne Gobble, for your program to survive for some noticeable time, you have to make sure it is re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$), replicable (R$^4$) and re-usable (R$^5$).\\
 
@@
 
-\begin{lstlisting}
-# A random walker (R0)
-import numpy as np
+\begin{code}{Random walk (R$^0$)}
+# Random walk (R0: worst you can do)
+import random
 
-# randomly choose ten instances of either 1 or -1. 
-steps = np.random.choice([-1,+1], 10)
-y = steps.cumsum() # compute the cumulative sum (the sum of all preceeding elements)
-print y
-\end{lstlisting}
+x = 0
+for i in xrange(10):
+    step = random.choice([-1,+1])
+    x = x + step
+    print x,
+\end{code}
+% That would be too easy if the program was commented...
+% # randomly choose ten instances of either 1 or -1. 
+% # compute the cumulative sum (the sum of all preceeding elements)
 
 Executed, this program would display 
-\begin{lstlisting}
-[-1  0 -1 -2 -1 -2 -3 -2 -1 -2] # with `steps` being [-1 1 -1 -1  1 -1 -1  1  1 -1]
-\end{lstlisting}
+\begin{code}{Output}
+[-1  0 -1 -2 -1 -2 -3 -2 -1 -2] # with `steps` being [-1 +1 -1 -1 +1 -1 -1 +1 +1 -1]
+\end{code}
 
@@
 
-Did you ever wonder if you would be able to run a program you wrote some time
-ago? Let's consider a program you wrote yesterday. Most probably you can still
-run it today, but would this still be true in a week, in a month, in a year or
-in a decade? Probably you would legitimately answer positively for weeks
-or even months if you're optimistic, but you may have doubts when considering
-longer time periods such as years or even decades. Part of the problem is that
-computer science is evolving at a very fast pace and you cannot know how the
-system, the software and the libraries your program depends on will evolve.
-If you want your program to have a tiny chance to be runnable ten years from
-now, you better make it clear what are the dependencies in terms of systems,
-software and libraries.\\
+Have you ever try to re-run a program you wrote some years ago? Frustratingly, it is often surprisingly hard. 
+Part of the problem is that computer science is evolving at a fast pace and you cannot know how the system, the software and the libraries your program depends on will evolve.
+Since you wrote the code, you may have changed your computer or your operating system. The compiler, interpreter or set of libraries installed may have changed or been upgraded. 
+You may find yourself battling with arcane issues of library compatibility---thoroughly orthogonal to your immediate research goals---to execute again a code \emph{that worked perfectly before}. 
+% "thoroughly orthogonal to your immediate research goals": possibly problematic
+
+To be clear, writing code that will run flawlessly months from now---without maintenance---is challenging.
+When considering years or decades, the best efforts can be stymied by the smallest changes, outside of the programmer's control.
+And modernizing a unmaintained ten-year-old code can reveal itself to be so complex and expensive that recreating the old execution environment is easier.
+For this however, the dependencies in terms of systems, software and libraries must be made clear enough.
+
+A \emph{re-runnable} code is one that describes an execution environment in which it is executable and produce the expected results with enough details to be recreated.
+
 % TODO: cite Collberg 2014 study
@@
 
-\begin{lstlisting}
-# A random walker (R1)
-# Tested with Python 3.6.0 / Numpy 1.12.0
-import numpy as np
+\begin{code}{Random walk (R$^1$)}
+# Random walk (R2: re-runnable)
+# Tested with Python 3
+import random
+from itertools import accumulate
 
-# Compute random walk for 10 steps
-steps = 2*(np.random.uniform(-1, +1, 10) > 0) - 1
-y = steps.cumsum()
-print(y)
-\end{lstlisting}
+steps = random.choices([-1,+1], k=10)
+x = list(accumulate(steps))
+print(x) 
+\end{code}
 
-In our case, the R$^0$ version of our tiny walker uses the print {\em instruction} that is available in Python 2.x but has been deprecated in Python 3 in favor or a  print {\em function}.
-This means this program cannot be run with Python 3 and needs to be fixed.
+In our case, the R$^0$ version of our tiny walker seems to imply that any version of Python would be fine. This not the case for two reasons.
+
+First, the R$^0$ version uses the print {\em instruction} that is available in Python 2 (this widly used, and installed by default in many operating systems) but has been deprecated in Python 3 (the current version, first released in 2008, almost a decade ago) in favor or a  print {\em function}.
+This means that R$^0$ cannot be run with Python 3.
+
 Since Python 2 will soon be discontinued, it may be best to target directly Python 3 and to use the print function (and as a matter of fact, it will work with Python 2).
@@
 
-\begin{lstlisting}
-# A random walker (R2)
-# Tested with Python 3.6 / Numpy 1.12.0
-import numpy as np
+\begin{code}{Random walk (R$^2$)}
+# Random walk (R2: repeatable)
+# Tested with Python 3
+import random
+from itertools import accumulate
 
-# Random generator initialization
-seed = 1
-np.random.seed(seed)
+# Random number generator initialization
+random.seed(1)
 
 # Compute random walk for 10 steps
-steps = np.random.choice([-1,+1], 10)
-y = steps.cumsum()
-print(y)
-\end{lstlisting}
+steps = random.choices([-1,+1], k=10)
+x = list(accumulate(steps))
+print(x) 
+\end{code}
 
@@
 \section*{Reproducible (R$^{\mathbf 3}$)}
+As nicely explained by Mesnard and Barba \cite{Mesnard:2016}, reproducibility is harder than you may think.
 
@@
 
-\begin{lstlisting}
-# A random walker (R3)
+\begin{code}{Random walk (R$^3$)}
+# Random walk (R3: reproducible)
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
-# Release under the BSD 2 clauses license
-# Tested with Python 3.6 / Numpy 1.12.0 using macOS 10.12.4 on a 64 bits architecture
-import numpy as np
+# Release under the BSD 2-clause license
+# Tested with Python 3.6 / macOS 10.12.4 / 64 bits architecture
+import random
+from itertools import accumulate
 
@@
 
-    steps = 2*(np.random.uniform(-1,+1,n) > 0) - 1
-    return steps.cumsum()
+    steps = [-1 if random.uniform(-1,+1) < 0 else +1 for i in range(n)]
+    return list(accumulate(steps))
 
-if __name__ == '__main__'
+if __name__ == '__main__':
     # Unit test
-    np.random.seed(1)
-    assert walk(10) == [-1, 0, -1, -2, -3, -4, -5, -6, -7, -6]
+    random.seed(1)
+    assert walk(10) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
 
-    # Random generator initialization
-    seed = 1
-    np.random.seed(seed)
-
-    # Compute random walk for n steps
-    y = walk(n=10)
-
-    # Save and display results
-    results = {"seed": seed, "y": y}
-    print(results["y"])
-\end{lstlisting}
+    # Actual results
+    random.seed(2)
+    print(walk(10))
+\end{code}
 
@@
 
-\begin{lstlisting}
-# A random walker program (R4)
-# A replication of the random walker using pure Python
+\begin{code}{Random walk (R$^4$)}
+# Random walk (R4: replicable)
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
-# Release under the BSD 2 clauses license
-# Tested with Python 3.6 using macOS 10.12.4 on a 64 bits architecture
+# Release under the BSD 2-clause license
+# Tested with Python 3.6 / Numpy 1.12.0 / macOS 10.12.4 / 64 bits architecture
 import random
-from itertools import accumulate
+import numpy as np
 
-def state(seed):
-    """ Compute the Mersenne Twister keys from given seed """
-    mt = [seed] * (624+1)
-    mt[-1] = 624
-    for i in range(1, 624):
-        mt[i] = int(0xFFFFFFFF & (1812433253 * (mt[i-1]^mt[i-1]>>30)+i))
-    return (3, tuple(mt), None)
-
-def walk(n=10):
+def walk(rng, n):
     """ Random walk for n steps """
 
-    steps = [-1 if random.uniform(-1,+1) < 0 else +1 for i in range(n)]
-    return list(accumulate(steps))
+    steps = 2*(rng.uniform(-1,+1,n) > 0) - 1
+    return steps.cumsum().tolist()
 
-if __name__ == '__main__'
+def rng(seed):
+    """ Return a random number generator initialized with seed """
+
+    rng = random.Random()
+    rng.seed(seed)
+    _, keys, _ = rng.getstate()
+    rng = np.random.RandomState()
+    state = rng.get_state()
+    rng.set_state((state[0], keys[:-1], state[2], state[3], state[4]))
+    return rng
+
+if __name__ == '__main__':
     # Unit test
-    random.setstate(state(seed=1))
-    assert walk(n=10) == [-1, 0, -1, -2, -3, -4, -5, -6, -7, -6]
+    assert walk(rng(seed=1), 10) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
 
-    # random generator initialization
-    seed = 1
-    random.setstate(state(seed))
-
-    # Compute random walk for n steps
-    y = walk(n=10)
-
-    # Save and display results
-    results = {"seed": seed, "y": y}
-    print(results["y"])
-\end{lstlisting}
+    # Actual results
+    print(walk(rng(seed=2), 10))
+\end{code}
 
@@
 \section*{Reusable (R$^{\mathbf 5}$)}
+
+\begin{code}{Random walk (R$^5$)}
+# Random walk (R5: re-usable)
+# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
+# Release under the BSD 2-clause license
+# Tested with Python 3.6 / Numpy 1.12.0 / macOS 10.12.4 / 64 bits architecture
+import random
+import numpy as np
+
+def walk(rng, n):
+    """ Random walk for n steps """
+
+    steps = 2*(rng.uniform(-1,+1,n) > 0) - 1
+    return steps.cumsum().tolist()
+
+def rng(seed):
+    """ Return a random number generator initialized with seed """
+
+    rng = random.Random()
+    rng.seed(seed)
+    _, keys, _ = rng.getstate()
+    rng = np.random.RandomState()
+    state = rng.get_state()
+    rng.set_state((state[0], keys[:-1], state[2], state[3], state[4]))
+    return rng
+
+if __name__ == '__main__':
+    # Unit test
+    # assert walk(rng(seed=1), 10) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+
+    import argparse
+    parser = argparse.ArgumentParser("Random walk")
+    parser.add_argument('--seed', type=int, default=1,
+                        help='Seed for random number generator ')
+    parser.add_argument('n', type=int, default=10,
+                        help='Number of step(s) to walk')
+    args = parser.parse_args()
+
+    print("Seed:", args.seed)
+    print("Number of steps:", args.n)
+    print("Result:",  walk(rng(args.seed), args.n))
+\end{code}
 

### 2017-04-24 main.tex
Run Python, run!

@@
 
-\title{The R Quintuplet (R$^5$)}
+\title{Run Python, run!}
+% \title{Re-run, Repeat, Reproduce, Replicate, Re-use.}
+% \title{The R Quintuplet (R$^5$)}
+
 \author{Nicolas P. Rougier and Fabien C. Y. Benureau}
@@
 
-Scientific reproducibility and replicability issues have received increased attention in the past few years and computer science is no exception.
+Scientific reproducibility and replicability issues have received increased attention in the past few years and computer science is no exception \cite{Collberg:2016}.
 One could think that computer science would mostly be shielded from those issues.
-But precisely because it is easy to believe that if a program runs once (and gives expected results) it will do so forever, crucial steps to transform working code into meaningful scientific contributions are rarely undertaken; computer science is plagued by replication problems precisely because it seems impervious to them.\\
+But precisely because it is easy to believe that if a program runs once and gives expected results it will do so forever, 
+crucial steps to transform working code into meaningful scientific contributions are rarely undertaken. 
+In other words, computer science is plagued by replication problems precisely because it seems impervious to them.\\
 
-In fact, a program can fail in many different ways at many different levels for many different reasons.
-Borrowing the terms coined by Carole Anne Gobble, for your program to survive for some noticeable time, you have to make sure it is re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$), replicable (R$^4$) and re-usable (R$^5$).\\
+In fact, a program can fail as a scientific object in many different ways for many different reasons.
+Borrowing the terms coined by Carole Goble \cite{Goble:2016}, for your program to survive for some noticeable time, you have to make sure it is re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$), replicable (R$^4$) and re-usable (R$^5$).\\
 
@@
 \begin{code}{Random walk (R$^0$)}
-# Random walk (R0: worst you can do)
 import random
@@
 And modernizing a unmaintained ten-year-old code can reveal itself to be so complex and expensive that recreating the old execution environment is easier.
-For this however, the dependencies in terms of systems, software and libraries must be made clear enough.
+For this to happen however, the dependencies in terms of systems, software and libraries must be made clear enough.
 
 A \emph{re-runnable} code is one that describes an execution environment in which it is executable and produce the expected results with enough details to be recreated.
-
-% TODO: cite Collberg 2014 study
-\cite{Collberg:2016}
+As shown by Colberg and Proebsting \cite{Collberg:2016}, this is far from being neither obvious nor easy.
 
 \begin{code}{Random walk (R$^1$)}
-# Random walk (R2: re-runnable)
 # Tested with Python 3
@@
 In our case, the R$^0$ version of our tiny walker seems to imply that any version of Python would be fine. This not the case for two reasons.
-
 First, the R$^0$ version uses the print {\em instruction} that is available in Python 2 (this widly used, and installed by default in many operating systems) but has been deprecated in Python 3 (the current version, first released in 2008, almost a decade ago) in favor or a  print {\em function}.
 This means that R$^0$ cannot be run with Python 3.
-
-Since Python 2 will soon be discontinued, it may be best to target directly Python 3 and to use the print function (and as a matter of fact, it will work with Python 2).
-Another potential problem is the numpy `choice` function that has been introduced only in numpy v1.7.
-If for some reason, the installed version of numpy is older, the program will also fail.
-One quick fix is to use a simpler way to generate the steps (even if it is less efficient).
-Last and most important, we need to specify all the dependencies of our program.
+Second problem is the use of the {\tt xrange} operator that does not exist anymore in Python 3.
+It has been replaced by the ubiquituous {\tt range} operator.
+Since Python 2 will soon be discontinued, it may be best to target Python 3 and write a  new version that takes advantage of Python 3.
+Last and most important, we need to specify all the dependencies of our new program.
 
@@
 However this does not guarantee you will get the same results at each run.
-For example, if you're using a (pseudo) random number generator, you have to take care of initializing the generator using a specific seed.
-If not, you're doomed and you won't be able to get twice the same output. But still, there are actually many more pitfalls ahead.
-For example, you can play with the seed value and find one that produce an interesting result that you will save in a file.
-And the, you can continue to play with the seed in order get new results.
-But what if you want to come back to one specific result?
-How do you find the seed related to this specific output?
-Again, if you did not foresee this potential problem, you're doomed.\\
+For example, if you're using a (pseudo) random number generator (RNG), you have to take care of initializing the generator using a specific seed.
+If you don't do it, you won't be able to get twice the same output.
+But still, there are actually many more pitfalls ahead.
+For example, you can play with some parameters and find a specific set that produce interesting results that you save to a file.
+In the meantime, you continue playing with parameters in order get new results.
+But what if you want to come back to this one specific result you just saved? Did you save as well the set of parameters tied to this specific results? 
+This kind of situation happens quite regularly in the scientific literature \cite{Claerbout:2000} such that for some authors, it is virtually impossible to find the set of parameters that has been used to produce this or that figure.
+If you're note careful enough, this will happen to you as well.\\
 
-Turning back to our random walker example, we can fix it quite easily by initializing the random number generator and produce a dictionary holding seed and result that allows us to reach the R² level.
+Turning back to our random walk example, we can fix it quite easily by initializing the random number generator and reach the R² level.
 
 \begin{code}{Random walk (R$^2$)}
-# Random walk (R2: repeatable)
-# Tested with Python 3
+# Tested with Python 3.6
 import random
@@
 # Random number generator initialization
+seed = 1
 random.seed(1)
 
-# Compute random walk for 10 steps
+# Random walk for 10 steps
 steps = random.choices([-1,+1], k=10)
 x = list(accumulate(steps))
+
+# Display & save results
 print(x)
+with open("results-R2-%d.txt" % seed, "w") as file:
+    file.write(str(x))
 \end{code}
@@
 \section*{Reproducible (R$^{\mathbf 3}$)}
-As nicely explained by Mesnard and Barba \cite{Mesnard:2016}, reproducibility is harder than you may think.
+As nicely explained by Mesnard and Barba \cite{Mesnard:2016}, reproducibility is harder than you think. 
+Making a program to run on a different machine (than your own) is neither easy nor straightforward.
+First, you need to take care of listing precisely all the dependencies of your program.
+Depending on its complexity, it might be sufficient.
+But in some case this won't be enough.
+The reason is that you when you listed your the dependencies, you probably only listed explicit or direct dependencies.
+Problem is that these dependencies have in turn some dependencies.
+Those constitutes the hidden, indirect or implicit dependencies.
+And for these new implicit dependencies, you can have again new implicit dependecies. Scary.
+But things can be even more complex because you potentially have much more to take into account.
+For example the system your program is running on, the CPU type and architecture (32 bits, 64 bits), the endianness, to name just a few.
+Last, but not least, you may have set defaults somewhere, someday and you may have completely forgotten you did so. Unfortunately, this can also affect the way your program is ran.
+
+%\begin{itemize}
+%\itemsep 0em
+%\item Python 3
+%\item Python 3.6
+%\item {Python 3.6.1 (default, Mar 28 2017, 10:33:50) \\
+%      {\tt [GCC 4.2.1 Compatible Apple LLVM 8.1.0 (clang-802.0.38)] on darwin}}
+%\item {Python 3.6.1 (default, Mar 28 2017, 10:33:50) \\
+%      {\tt [GCC 4.2.1 Compatible Apple LLVM 8.1.0 (clang-802.0.38)] on darwin}\\
+%      {\tt configuration variables: -std=c99 -Wextra -Wno-unused-result -Wno-unused-parameter -Wno-missing-field-initializers ...}}
+%\end{itemize}
 
@@
 \begin{code}{Random walk (R$^3$)}
-# Random walk (R3: reproducible)
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
@@
 
-    # Actual results
-    random.seed(2)
-    print(walk(10))
+    # Random walk for 10 steps
+    seed = 1
+    random.seed(seed)
+    x = walk(10)
+
+    # Display & save results
+    print(x)
+    with open("results-R3-%d.txt" % seed, "w") as file:
+        file.write(str(x))
 \end{code}
@@
 
-There are some discrepancies between the numpy and the pure Python version due to subtle differences in the random generator.
+As explained by Peng et al. \cite{Peng:2006}, {\em the replication of important findings by multiple independent investigators is fundamental to the accumulation of scientific evidence.}\\
+
+In our case, the replication of the random walk brings some unexpected discrepancies between the original Python version and the replicated Numpy version.
+The reason is that there exist some subtle differences in the respective implementations of the RNG.
 As explained in the \href{https://docs.python.org/3.6/library/random.html}{Python documentation}, {\em Python uses the Mersenne Twister as the core generator. It produces 53-bit precision floats
@@
 The underlying implementation in C is both fast and threadsafe.}
-But the numpy random generator is also based on the same Mersenne Twister generator and the difference comes from the way seed is interpreted to initialize the generator.
-Fortunately, Python gives access to the internal state and we can {\em fix} it to make it match the behavior of the numpy generator (which is the default).
+Numpy RNG is also based on the Mersenne Twister generator but there are differences in the way seed is interpreted when initializing the generator.
+Fortunately, Both Numpy and Python offer access to the internal state and we can {\em fix} it to make the behavior of one RNG to match the other RNG.
+Here, we want to make the Numpy RNG to match the Python RNG behvior.
 
 \begin{code}{Random walk (R$^4$)}
-# Random walk (R4: replicable)
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
@@
 
-    # Actual results
-    print(walk(rng(seed=2), 10))
+    # Random walk for 10 steps
+    seed = 1
+    x = walk(rng(seed=2), 10)
+
+    # Display & save results
+    print(x)
+    with open("results-R4-%d.txt" % seed, "w") as file:
+        file.write(str(x))
 \end{code}
@@
 \begin{code}{Random walk (R$^5$)}
-# Random walk (R5: re-usable)
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
@@
 
+def test():
+    """ Unit tests """
+
+    return walk(rng(seed=1), 10) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+
 if __name__ == '__main__':
-    # Unit test
-    # assert walk(rng(seed=1), 10) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
-
     import argparse
@@
 
+    # Random walk for n steps
+    x = walk(rng(args.seed), args.n)
+
+    # Display & save results
     print("Seed:", args.seed)
     print("Number of steps:", args.n)
-    print("Result:",  walk(rng(args.seed), args.n))
+    print("Result:",  x)
+    with open("results-R5-%d.txt" % seed, "w") as file:
+        file.write("Version: R5")
+        file.write("Seed: %d" % args.seed)
+        file.write("Steps number: %d" % args.n)
+        file.write("Output: %s" % str(x))
 \end{code}

### 2017-05-01 main.tex
We're late

@@
           anchor=south west, yshift=-0.25pt, xshift=5pt,
-          font=\sc\footnotesize] at (O|-P) {\csname mdf@@codeheading\endcsname};},
+          font=\footnotesize] at (O|-P) {\csname mdf@@codeheading\endcsname};},
   firstextra={
@@
           anchor=south west, yshift=-0.5pt, xshift=5pt,
-          font=\sc\footnotesize] at (O|-P) {\csname mdf@@codeheading\endcsname};}
+          font=\footnotesize] at (O|-P) {\csname mdf@@codeheading\endcsname};}
 }
@@
 
-\title{Run Python, run!}
-% \title{Re-run, Repeat, Reproduce, Replicate, Re-use.}
+\title{Re-run, Repeat, Reproduce, Replicate, Re-use:\\Transforming Code into Scientific Contributions}
+% \title{Run Python, run!}
 % \title{The R Quintuplet (R$^5$)}
@@
 
-Scientific reproducibility and replicability issues have received increased attention in the past few years and computer science is no exception \cite{Collberg:2016}.
-One could think that computer science would mostly be shielded from those issues.
-But precisely because it is easy to believe that if a program runs once and gives expected results it will do so forever, 
-crucial steps to transform working code into meaningful scientific contributions are rarely undertaken. 
-In other words, computer science is plagued by replication problems precisely because it seems impervious to them.\\
+Replication is a cornerstone of science. 
+% Since science aims to discover general principles of how the world functions, 
+% * <nicolas.rougier@inria.fr> 2017-04-24T15:49:43.902Z:
+% 
+% > Since science aims to discover general principles of how the world functions, 
+% 
+% 
+% C'est un peu une déclaration à l'emporte pièce et je soupçconne qu'il y a une littérature énorme sur la définition de la science.
+Et à moins de vouloir s'aliéner toute l'astophysioque et la cosmologie, on pourrait au minimum étendre à l'univers.
+% 
+% ^ <fabien.benureau+overleaf@gmail.com> 2017-04-25T13:29:01.309Z:
+% 
+% Reopened by editing the source!
+% From [wikipedia](https://en.wikipedia.org/wiki/World): "In a philosophical context, the world is the whole of the physical Universe, or an ontological world."
+% 
+% ^ <nicolas.rougier@inria.fr> 2017-04-25T13:45:04.565Z:
+%
+% Mouais.
+%
+% ^.
+If an experimental result cannot be obtained more than once,
+% it might be because of a fluke, a bias, a bug, a false positive or even a hack.
+%it becomes, at best, nothing more than food for thought for future research. 
+it merely becomes, at best, an observation that
+may feed future research. 
+Replication issues have received increased attention in recent years,
+in the domain of medicine and psychology in particular.
+One could think that computer science would mostly be shielded from such issues.
+But precisely because it is easy to believe that
+if a program runs once and gives expected results it will do so forever, 
+crucial steps to transform working code into meaningful scientific contributions are rarely undertaken \cite{Collberg:2016}. 
+In other words, computer science is plagued by replication problems,
+in part, precisely because it seems impervious to them.
 
-In fact, a program can fail as a scientific object in many different ways for many different reasons.
-Borrowing the terms coined by Carole Goble \cite{Goble:2016}, for your program to survive for some noticeable time, you have to make sure it is re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$), replicable (R$^4$) and re-usable (R$^5$).\\
+In fact, a program can fail as a scientific contribution
+in many different ways for many different reasons.
+Borrowing the terms coined by Carole Goble \cite{Goble:2016},
+for your program to survive for some noticeable time,
+you have to make sure it is
+re-runnable (R$^1$),
+repeatable (R$^2$),
+reproducible (R$^3$),
+replicable (R$^4$)
+and re-usable (R$^5$).\\
 
-Let us illustrate this with a very simple example, a random walk written in Python:
+Let us illustrate this with a very simple example,
+a random walk written in Python:
 
-\begin{code}{Random walk (R$^0$)}
+\begin{code}{\textbf{\textsc{Listing 1:}} Random walk (R$^0$)}
 import random
@@
     step = random.choice([-1,+1])
-    x = x + step
+    x += step
     print x,
@@
 \begin{code}{Output}
-[-1  0 -1 -2 -1 -2 -3 -2 -1 -2] # with `steps` being [-1 +1 -1 -1 +1 -1 -1 +1 +1 -1]
+-1  0 -1 -2 -1 -2 -3 -2 -1 -2 # with the steps being -1, +1, -1, -1, +1, -1, -1, +1, +1, -1
 \end{code}
@@
 
-Have you ever try to re-run a program you wrote some years ago? Frustratingly, it is often surprisingly hard. 
-Part of the problem is that computer science is evolving at a fast pace and you cannot know how the system, the software and the libraries your program depends on will evolve.
-Since you wrote the code, you may have changed your computer or your operating system. The compiler, interpreter or set of libraries installed may have changed or been upgraded. 
-You may find yourself battling with arcane issues of library compatibility---thoroughly orthogonal to your immediate research goals---to execute again a code \emph{that worked perfectly before}. 
+Have you ever try to re-run a program you wrote some years ago?
+Frustratingly, it can often be surprisingly hard. 
+Part of the problem is that technology is evolving at a fast pace
+and you cannot  know in advance
+how the system, the software and the libraries your program depends on will evolve.
+Since you wrote the code,
+you may have changed your computer or your operating system.
+The compiler, interpreter or set of libraries installed may have changed or been upgraded. 
+You may find yourself battling with
+arcane issues of library compatibility---thoroughly orthogonal to your immediate research goals---to execute again a code \emph{that worked perfectly before}. 
 % "thoroughly orthogonal to your immediate research goals": possibly problematic
 
-To be clear, writing code that will run flawlessly months from now---without maintenance---is challenging.
-When considering years or decades, the best efforts can be stymied by the smallest changes, outside of the programmer's control.
-And modernizing a unmaintained ten-year-old code can reveal itself to be so complex and expensive that recreating the old execution environment is easier.
-For this to happen however, the dependencies in terms of systems, software and libraries must be made clear enough.
+To be clear, it is impossible to write future-proof code,
+as the best efforts can be stymied by the smallest change in one of the dependencies.
+On the other hand, modernizing an unmaintained ten-year-old code
+can reveal itself to be an difficult and expensive undertaking,
+with each change risking to affect the semantics of the program.
+In fact, recreating the old execution environment
+might be easier.
+For this to happen however,
+the dependencies in terms of systems, software and libraries
+must be made clear enough.
 
-A \emph{re-runnable} code is one that describes an execution environment in which it is executable and produce the expected results with enough details to be recreated.
-As shown by Colberg and Proebsting \cite{Collberg:2016}, this is far from being neither obvious nor easy.
+A \emph{re-runnable} code is one that describes
+an execution environment in which it is executable
+and produce the expected results with enough details to be recreated.
+As shown by Colberg and Proebsting \cite{Collberg:2016},
+this is far from being neither obvious nor easy.
 
-\begin{code}{Random walk (R$^1$)}
+% if using random.choices, won't be runnable with anything <3.6
+% if using itertools.accumulate, won't be runnable with anything <3.2
+% although perhaps not the best syntax, we might want to keep the
+% import mod
+% mod.function 
+% syntax for both random and intertools
+%
+% Problem with current code is that it is highly inefficient now (compared to R0)
+\begin{code}{\textbf{\textsc{Listing 2:}} Random walk (R$^1$)}
 # Tested with Python 3
 import random
-from itertools import accumulate
 
-steps = random.choices([-1,+1], k=10)
-x = list(accumulate(steps))
-print(x) 
+walk, total = [], 0
+for i in range(10):
+    step = random.choice([-1,+1])     
+    total += step
+    walk.append(total)
+
+print(walk)
 \end{code}
 
-In our case, the R$^0$ version of our tiny walker seems to imply that any version of Python would be fine. This not the case for two reasons.
-First, the R$^0$ version uses the print {\em instruction} that is available in Python 2 (this widly used, and installed by default in many operating systems) but has been deprecated in Python 3 (the current version, first released in 2008, almost a decade ago) in favor or a  print {\em function}.
+In our case, the R$^0$ version of our tiny walker seems to imply
+that any version of Python would be fine.
+This not the case for two reasons.
+First, the R$^0$ version uses the print {\em instruction}
+that is available in Python 2 (widely used, and installed by default in many operating systems)
+but has been deprecated in Python 3 (first released in 2008, almost a decade ago)
+in favor or a  print {\em function}.
 This means that R$^0$ cannot be run with Python 3.
-Second problem is the use of the {\tt xrange} operator that does not exist anymore in Python 3.
+Second problem is the use of the {\tt xrange} operator
+that does not exist anymore in Python 3.
 It has been replaced by the ubiquituous {\tt range} operator.
-Since Python 2 will soon be discontinued, it may be best to target Python 3 and write a  new version that takes advantage of Python 3.
-Last and most important, we need to specify all the dependencies of our new program.
+Since Python 2 is on its way out,
+it may be best to target Python 3, as we do in the R$^1$ version.
+Incidentally, it remains compatible with Python 2.
+
+But whatever the version of the dependencies chosen,
+the crucial step here is to document it.
 
@@
 
-After having taken care of specifying dependencies, you can legitimately hope to be able to run the program without too much trouble.
-However this does not guarantee you will get the same results at each run.
-For example, if you're using a (pseudo) random number generator (RNG), you have to take care of initializing the generator using a specific seed.
-If you don't do it, you won't be able to get twice the same output.
-But still, there are actually many more pitfalls ahead.
-For example, you can play with some parameters and find a specific set that produce interesting results that you save to a file.
-In the meantime, you continue playing with parameters in order get new results.
-But what if you want to come back to this one specific result you just saved? Did you save as well the set of parameters tied to this specific results? 
-This kind of situation happens quite regularly in the scientific literature \cite{Claerbout:2000} such that for some authors, it is virtually impossible to find the set of parameters that has been used to produce this or that figure.
-If you're note careful enough, this will happen to you as well.\\
+% * <fabien.benureau+overleaf@gmail.com> 2017-04-27T06:18:15.753Z:
+% 
+% - 
+% 
+% Ok, so I am not sure that the parameter talk should be here. It should be in reproducible. 
+% 
+% ^.
 
-Turning back to our random walk example, we can fix it quite easily by initializing the random number generator and reach the R² level.
+The code is running and producing the expected results. 
+The next step is to make sure that you can produce the same output over successive runs of your program, 
+should you want to. 
+In other words, the next step is to make your program deterministic, 
+producing {\em repeatable} output.
 
-\begin{code}{Random walk (R$^2$)}
-# Tested with Python 3.6
+Such repeatability is useful. 
+If a run of the program produces a particularly puzzling results, 
+it allows you to scrutinize any step of the execution of the program by re-running it again. 
+It is also the easiest way to prove that your program did indeed produce the results you published.
+
+Such repeatability is not always possible or easy \cite{Diethelm:2012}.
+But for sequential programs not depending on exotic hardware, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).
+
+For our program, that means setting the seed of the {\tt random}  module.
+We may also want to save the output of the program to a file, so that we can easily verify that consecutive runs do produce the same output; eyeballing differences is time-consuming (and therefore won't be done systematically) and unreliable.
+
+\begin{code}{\textbf{\textsc{Listing 3:}} Random walk (R$^2$)}
+# Tested with Python 3
 import random
-from itertools import accumulate
 
-# Random number generator initialization
-seed = 1
-random.seed(1)
+random.seed(0) # RNG initialization
 
-# Random walk for 10 steps
-steps = random.choices([-1,+1], k=10)
-x = list(accumulate(steps))
+walk, total = [], 0
+for i in range(10):
+    step = random.choice([-1,+1])     
+    total += step
+    walk.append(total)
 
-# Display & save results
-print(x)
-with open("results-R2-%d.txt" % seed, "w") as file:
+print(walk)
+with open("results.txt", "w") as file: # Saving output to disk
     file.write(str(x))
@@
 
+Setting seeds is not without danger. 
+Using 439 as a seed in the previous program would result in ten consecutive +1 steps, which---although a perfectly valid random walk---lend itself to gross misinterpretations of the overall dynamic of the algorithm.
+Verifying that the qualitative aspects of the results and the conclusion we make are not tied to a specific initialization of the pseudo-random generator is an integral part of any scientific undertaking in computational science;
+this is usually done by repeating the simulations multiple times with different seeds.
+
+Determinism is a tool that allow to study bugs and spurious or unexpected behavior. 
+Determinism is also critical for other to reproduce and study your results. 
+% talk about controversy about setting seeds
+
+% After having taken care of specifying dependencies,
+% you can legitimately hope to be able to
+% run the program without too much trouble.
+% However this does not guarantee
+% you will get the same results at each run.
+% For example, if you're using a (pseudo) random number generator (RNG),
+% you have to take care of initializing the generator using a specific seed.
+% If you don't do it, you won't be able to get twice the same output.
+% But still, there are actually many more pitfalls ahead.
+% For example, you can play with some parameters and
+% find a specific set that produce interesting results that you save to a file.
+% In the meantime, you continue playing with parameters in order get new results.
+% But what if you want to come back to this one specific result you just saved?
+% Did you save as well the set of parameters tied to this specific results? 
+% This kind of situation happens quite regularly in the scientific literature \cite{Claerbout:2000} 
+% such that in some case, it is virtually impossible
+% to find the set of parameters that has been used to produce this or that figure.
+% If you're note careful enough, this will happen to you as well.\\
+
 \clearpage
 \section*{Reproducible (R$^{\mathbf 3}$)}
-As nicely explained by Mesnard and Barba \cite{Mesnard:2016}, reproducibility is harder than you think. 
-Making a program to run on a different machine (than your own) is neither easy nor straightforward.
+
+The R$^2$ code seems fine enough, but it hides several problems that come to light when trying to {\em reproduce} results.
+In the strongest sense, a code is said to be {\em reproducible} if the same results it produced in the past can be re-obtained.
+
+As nicely explained by Mesnard and Barba \cite{Mesnard:2016}, 
+reproducibility is harder than you think. 
+Making a program to run on a machine different from your own
+is neither easy nor straightforward.
 First, you need to take care of listing precisely all the dependencies of your program.
@@
 But in some case this won't be enough.
-The reason is that you when you listed your the dependencies, you probably only listed explicit or direct dependencies.
+The reason is that you when you listed your the dependencies,
+you probably only listed explicit or direct dependencies.
 Problem is that these dependencies have in turn some dependencies.
 Those constitutes the hidden, indirect or implicit dependencies.
-And for these new implicit dependencies, you can have again new implicit dependecies. Scary.
-But things can be even more complex because you potentially have much more to take into account.
-For example the system your program is running on, the CPU type and architecture (32 bits, 64 bits), the endianness, to name just a few.
-Last, but not least, you may have set defaults somewhere, someday and you may have completely forgotten you did so. Unfortunately, this can also affect the way your program is ran.
+And for each of these implicit dependencies, 
+you can have of course new explicit and implicit dependencies.
+Scary!
+
+But things can be even more complex because
+you potentially have much more to take into account such as for example
+the system your program is running on, 
+the CPU type and architecture (e.g. 32 bits or 64 bits),
+the endianness,
+to name just a few.
+Last, but not least,
+you may have set defaults somewhere, someday
+and you may have completely forgotten you did so.
+Unfortunately, this can also affect the way your program is ran.
 
@@
 
-\begin{code}{Random walk (R$^3$)}
+\begin{code}{\textbf{\textsc{Listing 4:}} Random walk (R$^3$)}
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
@@
 
-\begin{code}{Random walk (R$^4$)}
+\begin{code}{\textbf{\textsc{Listing 5:}} Random walk (R$^4$)}
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
@@
 
-\begin{code}{Random walk (R$^5$)}
+Re-usable means your program can be used by people outside your lab,
+possibly using different environment, different parameters or different data.
+If people start using your program,
+they will most likely report bugs or malfunctions.
+And if you're lucky, they will even propose fixes.
+In other words, if your program is re-usable and is re-used, it is alive.
+As such and as for any other living being,
+it will evolve and mutate to adapt to its environment.
+This process will ensure long-term reproducibility
+to the extent people continue using and maintaining the program.
+
+\begin{code}{\textbf{\textsc{Listing 6:}} Random walk (R$^5$)}
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
@@
 
-    rng = random.Random()
+    rng = np.random.RandomState()
 	rng.seed(seed)
-    _, keys, _ = rng.getstate()
-    rng = np.random.RandomState()
-    state = rng.get_state()
-    rng.set_state((state[0], keys[:-1], state[2], state[3], state[4]))
     return rng

### 2017-05-09 main.tex
Reproducible Overhaul

@@
 crucial steps to transform working code into meaningful scientific contributions are rarely undertaken \cite{Collberg:2016}. 
-In other words, computer science is plagued by replication problems,
+Computer science is plagued by replication problems,
 in part, precisely because it seems impervious to them.
@@
 Borrowing the terms coined by Carole Goble \cite{Goble:2016},
-for your program to survive for some noticeable time,
+for your program to contribute to Science,
 you have to make sure it is
@@
 Since you wrote the code,
-you may have changed your computer or your operating system.
-The compiler, interpreter or set of libraries installed may have changed or been upgraded. 
+you may have reinstalled or upgraded  your operating system.
+The compiler, interpreter or set of libraries installed may have been replaced with newer versions. 
 You may find yourself battling with
@@
 as the best efforts can be stymied by the smallest change in one of the dependencies.
-On the other hand, modernizing an unmaintained ten-year-old code
-can reveal itself to be an difficult and expensive undertaking,
+At the same time, modernizing an unmaintained ten-year-old code
+can reveal itself to be an arduous and expensive undertaking,
 with each change risking to affect the semantics of the program.
-In fact, recreating the old execution environment
-might be easier.
+
+Rather trying to predict the future or painstakingly dusting off old code, 
+an often easier solution is to recreate the old execution environment.
 For this to happen however,
@@
 A \emph{re-runnable} code is one that describes
-an execution environment in which it is executable
-and produce the expected results with enough details to be recreated.
-As shown by Colberg and Proebsting \cite{Collberg:2016},
+an execution environment in which it is executable with enough details to be recreated.
+As shown by Collberg and Proebsting \cite{Collberg:2016},
 this is far from being neither obvious nor easy.
 
+% * <fabien.benureau+overleaf@gmail.com> 2017-05-08T18:09:47.037Z:
+% 
+% Wouldn't Collberg be better cited in R^2?
+% 
+% ^.
 % if using random.choices, won't be runnable with anything <3.6
@@
 that any version of Python would be fine.
-This not the case for two reasons.
-First, the R$^0$ version uses the print {\em instruction}
-that is available in Python 2 (widely used, and installed by default in many operating systems)
-but has been deprecated in Python 3 (first released in 2008, almost a decade ago)
+This not the case because it uses the print {\em instruction} and the {\tt xrange} operator, both specific to Python 2.
+The print {\em instruction}, available in Python 2 (widely used, installed by default in many operating systems), 
+has been deprecated in Python 3 (first released in 2008, almost a decade ago)
 in favor or a  print {\em function}.
-This means that R$^0$ cannot be run with Python 3.
-Second problem is the use of the {\tt xrange} operator
-that does not exist anymore in Python 3.
-It has been replaced by the ubiquituous {\tt range} operator.
-Since Python 2 is on its way out,
-it may be best to target Python 3, as we do in the R$^1$ version.
+The {\tt xrange} operator has been replaced by the {\tt range} operator in Python 3.
+In order to try to future-proof the code a bit, we might as well target Python 3, as is done in the R$^1$ version.
 Incidentally, it remains compatible with Python 2.
 
-But whatever the version of the dependencies chosen,
+But whatever the version chosen,
 the crucial step here is to document it.
@@
 it allows you to scrutinize any step of the execution of the program by re-running it again. 
-It is also the easiest way to prove that your program did indeed produce the results you published.
+It is also the easiest way to prove that the program did indeed produce the published results.
 
@@
 For our program, that means setting the seed of the {\tt random}  module.
-We may also want to save the output of the program to a file, so that we can easily verify that consecutive runs do produce the same output; eyeballing differences is time-consuming (and therefore won't be done systematically) and unreliable.
+We may also want to save the output of the program to a file, so that we can easily verify that consecutive runs do produce the same output; eyeballing differences is unreliable and  time-consuming (and therefore won't be done systematically).
 
@@
 with open("results.txt", "w") as file: # Saving output to disk
-    file.write(str(x))
+    file.write(str(walk))
 \end{code}
@@
 
-Determinism is a tool that allow to study bugs and spurious or unexpected behavior. 
-Determinism is also critical for other to reproduce and study your results. 
+%Determinism is a tool that allow to study bugs and spurious or unexpected behavior. 
+%Determinism is also critical for other to reproduce and study your results. 
 % talk about controversy about setting seeds
 
-% After having taken care of specifying dependencies,
-% you can legitimately hope to be able to
-% run the program without too much trouble.
-% However this does not guarantee
-% you will get the same results at each run.
-% For example, if you're using a (pseudo) random number generator (RNG),
-% you have to take care of initializing the generator using a specific seed.
-% If you don't do it, you won't be able to get twice the same output.
-% But still, there are actually many more pitfalls ahead.
 % For example, you can play with some parameters and
@@
 
-The R$^2$ code seems fine enough, but it hides several problems that come to light when trying to {\em reproduce} results.
-In the strongest sense, a code is said to be {\em reproducible} if the same results it produced in the past can be re-obtained.
+The R$^2$ code seems fine enough, 
+but it hides several problems that come to light when trying to {\em reproduce} results.
+In the strongest sense,
+a code is said to be {\em reproducible}
+if the same results it produced in the past can be re-obtained by someone else. As explained by Mesnard and Barba \cite{Mesnard:2016}, 
+reproducibility is harder than you think.
 
-As nicely explained by Mesnard and Barba \cite{Mesnard:2016}, 
-reproducibility is harder than you think. 
-Making a program to run on a machine different from your own
-is neither easy nor straightforward.
-First, you need to take care of listing precisely all the dependencies of your program.
-Depending on its complexity, it might be sufficient.
-But in some case this won't be enough.
-The reason is that you when you listed your the dependencies,
-you probably only listed explicit or direct dependencies.
-Problem is that these dependencies have in turn some dependencies.
-Those constitutes the hidden, indirect or implicit dependencies.
-And for each of these implicit dependencies, 
-you can have of course new explicit and implicit dependencies.
-Scary!
+For instance, our R$^2$ program will not produce the same results for everyone.
+It will, because it is repeatable, produce the same results over repeated executions.
+But it will not do so over different execution environments.
+The cause is found in a change that occurred in the pseudo-random number generator between CPython 3.2 and CPython 3.3.
 
-But things can be even more complex because
-you potentially have much more to take into account such as for example
-the system your program is running on, 
-the CPU type and architecture (e.g. 32 bits or 64 bits),
-the endianness,
-to name just a few.
-Last, but not least,
-you may have set defaults somewhere, someday
-and you may have completely forgotten you did so.
-Unfortunately, this can also affect the way your program is ran.
+Executed with CPython 3.2 or older, the code will produce the sequence 1, 2, 1, 2, 3, 4, 5, 6, 7, 6.
+But with CPython 3.3 or newer, it will produce 1, 2, 1, 0, 1, 0, 1, 0, -1, 0.
+Reproducibility impose to specify dependencies more precisely than re-runnability: the exact versions that were used must be documented.
+The platform should also be described as well: CPU architecture, operating system version, endianness, etc., as well as any defaults you may have set that might have an impact.
+Because of the increasing complexity of our computational stacks, retrieving and deciding what is pertinent might be difficult.
+A best effort must be made.
+
+Here, in addition to specifying the exact Python version and platform details, a short unit test is provided, that allows the code to verify its own reproducibility.
+Should this test fail, then there is little hope of reproducing the results.
+If it does not fail, problems may still be present.
+
+A second problem present in the R$^2$ program is that it does not store the parameters used along with the results.
+Therefore, should the code change after the production of the results, someone provided with the result file will not be able to know which seed was used to produce it.
+In R$^3$, the seed is stored in the name of the result file.
 
@@
 # Release under the BSD 2-clause license
-# Tested with Python 3.6 / macOS 10.12.4 / 64 bits architecture
+# Tested with CPython 3.6 / macOS 10.12.4 / 64 bits architecture
 import random
@@
 \end{code}
+
+Finally, a crucial condition to making a code reproducible is making it available.
+As shown in \cite{Collberg:2016}, code is often not available, or available upon request.
+The latter may seem sufficient, but changes in email address,  career or retirement can make the code just as unavailable.
+Code \emph{and input and result files} should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as Figshare or Zenodo.
+
+Providing results files is important: it allows other researchers to easily compare the results they obtains with the one used to generate the figure in the article.
+Comparing figures directly can be cumbersome and often only allows to conclude about a qualitative similarity.
+
+To recap, reproducibility imposes stronger conditions than re-runnability.
+Dependencies and platforms must be described as precisely and a specifically as possible.
+Unit tests are a good way to embed self-diagnostics of reproducibility in the code.
+Results must be accompanied with the parameters and inputs used to produce them.
+And the code, and the data files behind the figure must be made available.
 
@@
 In other words, if your program is re-usable and is re-used, it is alive.
-As such and as for any other living being,
+As such, and as for any other living being,
 it will evolve and mutate to adapt to its environment.

### 2017-05-20 main.tex
Corrupted document

@@
-\documentclass[a4paper,11pt]{article}
-
-% Wider margins (compared to default)
-\usepackage[margin=2.5cm]{geometry}
-
-% English support (typography and hyphenation)
-\usepackage[english]{babel}
-
-% Unicode encoding
-\usepackage[utf8]{inputenc}
-
-% Better default font (Libertine and Inconsolata)
-\usepackage[ttscale=.875]{libertine}
-\usepackage[scaled=0.96]{zi4}
-
-% Biber
-\usepackage[backend=biber, style=numeric, natbib=true]{biblatex}
-\bibliography{library.bib}
-
-% Graphics
-\usepackage{graphicx}
-
-% Code Listing
-\usepackage{listings}
-\usepackage[framemethod=tikz]{mdframed}
-\makeatletter
-\def\mdf@@codeheading{Code Listings}
-\define@key{mdf}{title}{\def\mdf@@codeheading{#1}}
-\mdfdefinestyle{lstlisting}{%
-  backgroundcolor=black!2.5,
-  innertopmargin=2pt,
-  middlelinewidth=0.75pt,
-  outerlinewidth=9pt,
-  outerlinecolor=white,
-  innerleftmargin=10pt,
-  innerrightmargin=10pt,
-  leftmargin=0pt,
-  rightmargin=0pt,
-  rightline=false,
-  leftline=false,
-  bottomline=false,
-  skipabove=\topskip,
-  skipbelow=\topskip,
-  roundcorner=0pt,
-  singleextra={
-    \node[text=black, % fill=white, draw,
-          anchor=south west, yshift=-0.25pt, xshift=5pt,
-          font=\footnotesize] at (O|-P) {\csname mdf@@codeheading\endcsname};},
-  firstextra={
-    \node[text=black, % fill=white, draw,
-          anchor=south west, yshift=-0.5pt, xshift=5pt,
-          font=\footnotesize] at (O|-P) {\csname mdf@@codeheading\endcsname};}
-}
-
-\lstset{ %
-  basicstyle=\small\tt\linespread{0.75},
-  language=Python,
-  commentstyle=\color{gray},
-  columns=fullflexible,
-}
-
-\lstnewenvironment{code}[2][]{%
-  \lstset{#1}%
-  \mdframed[style=lstlisting,title={#2}]%
-}{\endmdframed}
-
-%\usepackage{listings}
-%\surroundwithmdframed[
-%%  hidealllines=true,
-%  linewidth=0.25pt,
-%  linecolor=black!50,
-%  backgroundcolor=black!4,
-%  innerleftmargin=8pt,
-%  innertopmargin=3pt,
-%  innerbottommargin=3pt]{lstlisting}
-%\lstset{ %
-%  basicstyle=\small\tt\linespread{0.75},
-%  language=Python,
-%  commentstyle=\color{gray},
-%  columns=fullflexible,
-%}
-
-% Hyperref
-\usepackage{xcolor}
-\definecolor{blendedblue}{rgb}{0.2, 0.2, 0.6}
-\definecolor{blendedred}{rgb}{0.8, 0.2, 0.2}
-\usepackage[bookmarks=true,
-            breaklinks=true,
-            pdfborder={0 0 0},
-            citecolor=blendedblue,
-            colorlinks=true,
-            linkcolor=blendedblue,
-            urlcolor=blendedblue,
-            citecolor=blendedblue,
-            linktocpage=false,
-            hyperindex=true,
-            linkbordercolor=white]{hyperref}
-\usepackage{hyperref}
-\hypersetup{colorlinks=true}
-
-\title{Re-run, Repeat, Reproduce, Replicate, Re-use:\\Transforming Code into Scientific Contributions}
-% \title{Run Python, run!}
-% \title{The R Quintuplet (R$^5$)}
-
-\author{Nicolas P. Rougier and Fabien C. Y. Benureau}
-\date{\today}
-
-\begin{document}
-\maketitle
-
+-
 % -------------------------
@@
 This not the case because it uses the print {\em instruction} and the {\tt xrange} operator, both specific to Python 2.
-The print {\em instruction}, available in Python 2 (widely used, installed by default in many operating systems), 
+The print {\em instruction}, available in Python 2 (still widely used, will stop to be supported in 2020), 
 has been deprecated in Python 3 (first released in 2008, almost a decade ago)
@@
 
-As explained by Peng et al. \cite{Peng:2006}, {\em the replication of important findings by multiple independent investigators is fundamental to the accumulation of scientific evidence.}\\
+As explained by Peng et al.
+\cite{Peng:2006}, {\em the replication of important findings by multiple independent investigators is fundamental to the accumulation of scientific evidence}.
+Replicability is the assumption that any article that does not provide the code source makes: that the description it provides of the algorithms is sufficiently precise and complete to re-obtain the results it presents.
 

### 2017-05-20 main.tex
Fixed document?

--- 2017-05-09 main.tex
@@
 \maketitle
-
 % -------------------------

### 2017-05-29 main.tex
started re-usable section

@@
 % English support (typography and hyphenation)
-\usepackage[english]{babel}
+\usepackage[american]{babel}
+\usepackage{csquotes}
 
@@
 % Biber
-\usepackage[backend=biber, style=numeric, natbib=true]{biblatex}
+\usepackage[backend=biber,
+            style=apa,
+            maxcitenames=3,
+            maxbibnames=99,
+            apamaxprtauth=99,	
+            natbib=true]{biblatex}
+\DeclareLanguageMapping{american}{american-apa}
 \bibliography{library.bib}
@@
 
-\title{Re-run, Repeat, Reproduce, Replicate, Re-use:\\Transforming Code into Scientific Contributions}
+\title{Re-run, Repeat, Reproduce, Re-use, Replicate:\\Transforming Code into Scientific Contributions}
 % \title{Run Python, run!}
@@
 
-Replication is a cornerstone of science. 
+%Replication is a cornerstone of science. 
+Reproducibility is a cornerstone of science. 
 % Since science aims to discover general principles of how the world functions, 
@@
 if a program runs once and gives expected results it will do so forever, 
-crucial steps to transform working code into meaningful scientific contributions are rarely undertaken \cite{Collberg:2016}. 
+crucial steps to transform working code into meaningful scientific contributions are rarely undertaken \citep{Collberg:2016}. 
 Computer science is plagued by replication problems,
@@
 in many different ways for many different reasons.
-Borrowing the terms coined by Carole Goble \cite{Goble:2016},
+Borrowing the terms coined by \citeauthor{Goble:2016} \citep{Goble:2016},
 for your program to contribute to Science,
@@
 reproducible (R$^3$),
-replicable (R$^4$)
-and re-usable (R$^5$).\\
+re-usable (R$^4$)
+and replicable (R$^5$).\\
 
@@
 For this to happen however, 
-the dependencies in terms of systems, software and libraries
-must be made clear enough.
+the dependencies in terms of systems, software and libraries must be made clear enough.
 
-A \emph{re-runnable} code is one that describes
-an execution environment in which it is executable with enough details to be recreated.
-As shown by Collberg and Proebsting \cite{Collberg:2016},
-this is far from being neither obvious nor easy.
+A \emph{re-runnable} code is one that describes an execution environment in which it is executable with enough details to be recreated.
+As shown by \citep{Collberg:2016}, this is far from being neither obvious nor easy.
 
@@
 but it hides several problems that come to light when trying to {\em reproduce} results.
-In the strongest sense,
-a code is said to be {\em reproducible}
-if the same results it produced in the past can be re-obtained by someone else. As explained by Mesnard and Barba \cite{Mesnard:2016}, 
-reproducibility is harder than you think.
+In the strongest sense, a code is said to be {\em reproducible}
+if the same results it produced in the past can be re-obtained by someone else.
+As explained by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility is harder than you think.
 
@@
 Finally, a crucial condition to making a code reproducible is making it available.
-As shown in \cite{Collberg:2016}, code is often not available, or available upon request.
+As shown in \citep{Collberg:2016}, code is often not available, or available upon request.
 The latter may seem sufficient, but changes in email address,  career or retirement can make the code just as unavailable.
@@
 \clearpage
-\section*{Replicable (R$^{\mathbf 4}$)}
+\section*{Reusable (R$^{\mathbf 4}$)}
 
-As explained by Peng et al.
-\cite{Peng:2006}, {\em the replication of important findings by multiple independent investigators is fundamental to the accumulation of scientific evidence}.
-Replicability is the assumption that any article that does not provide the code source makes: that the description it provides of the algorithms is sufficiently precise and complete to re-obtain the results it presents.
+Making your program re-usable means it can be used by people outside your lab,
+using possibly different environment, different parameters or different data.
+If people start using your program, they will most likely report bugs or malfunctions they encountered while using your program.
+If you're lucky enough, they will even propose you bugfix, hence improving the overall quality and correctness of your software.
+In other words, if your program is re-usable and is re-used, it is alive.
+As such, and as for any other living being, it will evolve and mutate to adapt to its environment.
+This process will ensure long-term reproducibility to the extent people continue using and maintaining the program.\\
 
-In our case, the replication of the random walk brings some unexpected discrepancies between the original Python version and the replicated Numpy version.
-The reason is that there exist some subtle differences in the respective implementations of the RNG.
-As explained in the \href{https://docs.python.org/3.6/library/random.html}{Python documentation}, {\em Python uses the Mersenne Twister as the core generator. It produces 53-bit precision floats
-and has a period of 2**19937-1.
-The underlying implementation in C is both fast and threadsafe.}
-Numpy RNG is also based on the Mersenne Twister generator but there are differences in the way seed is interpreted when initializing the generator.
-Fortunately, Both Numpy and Python offer access to the internal state and we can {\em fix} it to make the behavior of one RNG to match the other RNG.
-Here, we want to make the Numpy RNG to match the Python RNG behvior.
+However, making your program to be re-usable by others requires a little work.
+If it is natural, when one writes a program for oneself, to make implicit choices that suit one needs, such implicit choices won't resist the will of a user community.
+For example, having magic numbers or ill-defined constants in your program might be cumbersome for others.
+Similarly, it might be acceptable to not have a proper documentation when you're the only user (actually it might not be acceptable because the you now and the you in six months are two different persons), but other users want to know precisely how to use this or that function and what it does exactly.\\
 
-\begin{code}{\textbf{\textsc{Listing 5:}} Random walk (R$^4$)}
-# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
-# Release under the BSD 2-clause license
-# Tested with Python 3.6 / Numpy 1.12.0 / macOS 10.12.4 / 64 bits architecture
-import random
-import numpy as np
-
-def walk(rng, n):
-    """ Random walk for n steps """
-
-    steps = 2*(rng.uniform(-1,+1,n) > 0) - 1
-    return steps.cumsum().tolist()
-
-def rng(seed):
-    """ Return a random number generator initialized with seed """
-
-    rng = random.Random()
-    rng.seed(seed)
-    _, keys, _ = rng.getstate()
-    rng = np.random.RandomState()
-    state = rng.get_state()
-    rng.set_state((state[0], keys[:-1], state[2], state[3], state[4]))
-    return rng
-
-if __name__ == '__main__':
-    # Unit test
-    assert walk(rng(seed=1), 10) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
-
-    # Random walk for 10 steps
-    seed = 1
-    x = walk(rng(seed=2), 10)
-
-    # Display & save results
-    print(x)
-    with open("results-R4-%d.txt" % seed, "w") as file:
-        file.write(str(x))
-\end{code}
-
-\clearpage
-\section*{Reusable (R$^{\mathbf 5}$)}
-
-Re-usable means your program can be used by people outside your lab,
-possibly using different environment, different parameters or different data.
-If people start using your program,
-they will most likely report bugs or malfunctions.
-And if you're lucky, they will even propose fixes.
-In other words, if your program is re-usable and is re-used, it is alive.
-As such, and as for any other living being,
-it will evolve and mutate to adapt to its environment.
-This process will ensure long-term reproducibility
-to the extent people continue using and maintaining the program.
+If we look back at our random walker, we can list all the implicit choices that have been made:
+\begin{itemize}
+\item Step number has been hard-coded (magic number 10)
+\item Step size has been fixed to 1
+\item Initial position has been set to 0
+\item No possibility of using a user-specified random seed
+\item Only one run can be started at once
+\item Walker walk along a single dimension
+\item ...
+\end{itemize}
+However, we won't try to implement all these features at once, but only giving the possibility for a user to implement them if needed.
 
@@
 \clearpage
+\section*{Replicable (R$^{\mathbf 5}$)}
+
+Having made a software re-usable by others should offer a decent protection against bugs.
+Unfortunately, this is not always the case and some recent cases have demonstrated the tremendous impact a bug can have in Science \citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}.
+This is why, as explained by Peng et al. \cite{Peng:2006}, {\em the replication of important findings by multiple independent investigators is fundamental to the accumulation of scientific evidence}.
+%Replicability is the assumption that any article that does not provide the code source makes: that the description it provides of the algorithms is sufficiently precise and complete to re-obtain the results it presents.
+
+In our case, the replication of the random walk brings some unexpected discrepancies between the original Python version and the replicated Numpy version.
+The reason is to be found in a differences in the respective implementations of the RNG between Python and Numpy.
+As explained in the \href{https://docs.python.org/3.6/library/random.html}{Python documentation}, {\em Python uses the Mersenne Twister as the core generator. It produces 53-bit precision floats
+and has a period of 2**19937-1.
+The underlying implementation in C is both fast and threadsafe.}
+Numpy RNG is also based on the Mersenne Twister generator but there are differences in the way seed is interpreted when initializing the generator.
+Fortunately, Both Numpy and Python offer access to the internal state and we can {\em fix} it to make the behavior of one RNG to match the other RNG.
+Here, we want to make the Numpy RNG to match the Python RNG behvior.
+
+\begin{code}{\textbf{\textsc{Listing 5:}} Random walk (R$^4$)}
+# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
+# Release under the BSD 2-clause license
+# Tested with Python 3.6 / Numpy 1.12.0 / macOS 10.12.4 / 64 bits architecture
+import random
+import numpy as np
+
+def walk(rng, n):
+    """ Random walk for n steps """
+
+    steps = 2*(rng.uniform(-1,+1,n) > 0) - 1
+    return steps.cumsum().tolist()
+
+def rng(seed):
+    """ Return a random number generator initialized with seed """
+
+    rng = random.Random()
+    rng.seed(seed)
+    _, keys, _ = rng.getstate()
+    rng = np.random.RandomState()
+    state = rng.get_state()
+    rng.set_state((state[0], keys[:-1], state[2], state[3], state[4]))
+    return rng
+
+if __name__ == '__main__':
+    # Unit test
+    assert walk(rng(seed=1), 10) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+
+    # Random walk for 10 steps
+    seed = 1
+    x = walk(rng(seed=2), 10)
+
+    # Display & save results
+    print(x)
+    with open("results-R4-%d.txt" % seed, "w") as file:
+        file.write(str(x))
+\end{code}
+
+<
+
+\clearpage
 \section*{Conclusion}

### 2017-05-29 main.tex
conclusion

@@
 In order to try to future-proof the code a bit, we might as well target Python 3, as is done in the R$^1$ version.
-Incidentally, it remains compatible with Python 2.
+Incidentally, it remains compatible with Python 2.\\
 
@@
 
-\clearpage
+% \clearpage
 \section*{Repeatable (R$^{\mathbf 2}$)}
@@
 The code is running and producing the expected results. 
-The next step is to make sure that you can produce the same output over successive runs of your program, 
-should you want to. 
-In other words, the next step is to make your program deterministic, 
-producing {\em repeatable} output.
-
+The next step is to make sure that you can produce the same output over successive runs of your program, should you want to. 
+In other words, the next step is to make your program deterministic,  producing {\em repeatable} output. 
 Such repeatability is useful. 
-If a run of the program produces a particularly puzzling results, 
-it allows you to scrutinize any step of the execution of the program by re-running it again. 
+If a run of the program produces a particularly puzzling results,  it allows you to scrutinize any step of the execution of the program by re-running it again. 
 It is also the easiest way to prove that the program did indeed produce the published results.
-
 Such repeatability is not always possible or easy \cite{Diethelm:2012}.
-But for sequential programs not depending on exotic hardware, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).
+But for sequential programs not depending on exotic hardware, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\
 
@@
 
-\clearpage
+% \clearpage
 \section*{Reproducible (R$^{\mathbf 3}$)}
 
-The R$^2$ code seems fine enough, 
-but it hides several problems that come to light when trying to {\em reproduce} results.
-In the strongest sense, a code is said to be {\em reproducible}
-if the same results it produced in the past can be re-obtained by someone else.
+The R$^2$ code seems fine enough, but it hides several problems that come to light when trying to {\em reproduce} results.
+In the strongest sense, a code is said to be {\em reproducible} if the same results it produced in the past can be re-obtained by someone else.
 As explained by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility is harder than you think. 
-
-For instance, our R$^2$ program will not produce the same results for everyone.
-It will, because it is repeatable, produce the same results over repeated executions.
-But it will not do so over different execution environments.
-The cause is found in a change that occurred in the pseudo-random number generator between CPython 3.2 and CPython 3.3.
-
-Executed with CPython 3.2 or older, the code will produce the sequence 1, 2, 1, 2, 3, 4, 5, 6, 7, 6.
-But with CPython 3.3 or newer, it will produce 1, 2, 1, 0, 1, 0, 1, 0, -1, 0.
+For instance, our R$^2$ program will not produce the same results for everyone. It will, because it is repeatable, produce the same results over repeated executions.
+But it will not do so over different execution environments. The cause is found in a change that occurred in the pseudo-random number generator between CPython 3.2 and CPython 3.3. 
+Executed with CPython 3.2 or older, the code will produce the sequence 1, 2, 1, 2, 3, 4, 5, 6, 7, 6. But with CPython 3.3 or newer, it will produce 1, 2, 1, 0, 1, 0, 1, 0, -1, 0.
 Reproducibility impose to specify dependencies more precisely than re-runnability: the exact versions that were used must be documented.
 The platform should also be described as well: CPU architecture, operating system version, endianness, etc., as well as any defaults you may have set that might have an impact.
-Because of the increasing complexity of our computational stacks, retrieving and deciding what is pertinent might be difficult.
-A best effort must be made.
+Because of the increasing complexity of our computational stacks, retrieving and deciding what is pertinent might be difficult. A best effort must be made.\\
 
@@
 If it does not fail, problems may still be present.
-
 A second problem present in the R$^2$ program is that it does not store the parameters used along with the results.
@@
 Code \emph{and input and result files} should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as Figshare or Zenodo.
-
 Providing results files is important: it allows other researchers to easily compare the results they obtains with the one used to generate the figure in the article.
-Comparing figures directly can be cumbersome and often only allows to conclude about a qualitative similarity.
+Comparing figures directly can be cumbersome and often only allows to conclude about a qualitative similarity.\\
 
@@
 
-\clearpage
-\section*{Reusable (R$^{\mathbf 4}$)}
+% \clearpage
+\section*{Re-usable (R$^{\mathbf 4}$)}
 
@@
 
-\begin{code}{\textbf{\textsc{Listing 6:}} Random walk (R$^5$)}
+\begin{code}{\textbf{\textsc{Listing 5:}} Random walk (R$^4$)}
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
@@
 
-\clearpage
+% \clearpage
 \section*{Replicable (R$^{\mathbf 5}$)}
@@
 
-\begin{code}{\textbf{\textsc{Listing 5:}} Random walk (R$^4$)}
+\begin{code}{\textbf{\textsc{Listing 6:}} Random walk (R$^5$)}
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
@@
 
-<
+% \clearpage
+\section*{Conclusion}
 
-\clearpage
-\section*{Conclusion}
+Using a very simple random walk example using Python, we've seen a variety of pitfalls of writing scientific code.
+The R$^0$ form, while easy to write and simple to understand is not a good scientific code while the full-fledge R$^{4}$ form is much preferred but requires a fair amount of work that might be unacceptable in some circumstances.
+$R^3$ form seems to be an acceptable compromise and should be accepted as the minimum scientific standard.
+This means this should be actually checked by reviewers and publishers when code is part of a work worth to be published.
+But it's hardly the case today.\\
+
+The choice of the Python programming language to illustrate our point was not arbitrary.
+In less than a decade, Python has become one the few programming language of Science (together with R and Matlab).
+It is now used by a huge and growing number of researchers from different domains, having various experience and background in computer science.
+The readability and ease of use of Python make it a premium choice for new comers.
+But this is precisely this apparent simplicity that might hinder correct implementation.
+Inexperienced people may think any working code is a scientific code, but, as we've explained, this is simply not true.
+
+% Writing reproducible software requires both knowledge and experience. For the former, there are plenty of resources online or onsite.
 

### 2017-05-31 main.tex
Update on Overleaf.

@@
+% !TeX program = pdflatex
 \documentclass[a4paper,11pt]{article}
@@
 \usepackage[scaled=0.96]{zi4}
+
+%\usepackage{mathspec}
+%\setmainfont{Minion Pro Cond}
+%\setmathsfont(Digits,Greek,Latin)[Numbers={Proportional}]{Minion Pro}
+%\setmathrm{Minion Pro}
+%\setsansfont{MyriadPro-Cond}
 
@@
 it merely becomes, at best, an observation that
-may feed future research. 
+may feed future research \cite{Mesirov:2010,osc:2015}. 
 Replication issues have received increased attention in recent years,
-in the domain of medicine and psychology in particular.
+in the domain of medicine and psychology in particular \cite{Iqbal:2016}.
 One could think that computer science would mostly be shielded from such issues.
@@
 if a program runs once and gives expected results it will do so forever, 
-crucial steps to transform working code into meaningful scientific contributions are rarely undertaken \citep{Collberg:2016}. 
+crucial steps to transform working code into meaningful scientific contributions are rarely undertaken \citep{Sandve:2013,Schwab:2000}. 
 Computer science is plagued by replication problems,
-in part, precisely because it seems impervious to them.
+in part, precisely because it seems impervious to them.\\
 
@@
 It is also the easiest way to prove that the program did indeed produce the published results.
-Such repeatability is not always possible or easy \cite{Diethelm:2012}.
+Such repeatability is not always possible or easy \citep{Diethelm:2012, Courtes:2015}.
 But for sequential programs not depending on exotic hardware, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\
@@
 Unfortunately, this is not always the case and some recent cases have demonstrated the tremendous impact a bug can have in Science \citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}.
-This is why, as explained by Peng et al. \cite{Peng:2006}, {\em the replication of important findings by multiple independent investigators is fundamental to the accumulation of scientific evidence}.
-%Replicability is the assumption that any article that does not provide the code source makes: that the description it provides of the algorithms is sufficiently precise and complete to re-obtain the results it presents.
+This is why, as explained by Peng et al.
+\cite{Peng:2006}, {\em the replication of important findings by multiple independent investigators is fundamental to the accumulation of scientific evidence}.\\
+% Replicability is the assumption that any article that does not provide the code source makes: that the description it provides of the algorithms is sufficiently precise and complete to re-obtain the results it presents.
 
@@
 Fortunately, Both Numpy and Python offer access to the internal state and we can {\em fix} it to make the behavior of one RNG to match the other RNG.
-Here, we want to make the Numpy RNG to match the Python RNG behvior.
+Here, we want to make the Numpy RNG to match the Python RNG behavior.\\
+
+Without such explicit replication effort, it would be very hard to discover this faulty behavior and this why the ReScience journal has been created.
+This journal targets computational research and encourages the explicit replication of already published research, promoting new and open-source implementations in order to ensure that the original research is reproducible.
 
@@
 
-Using a very simple random walk example using Python, we've seen a variety of pitfalls of writing scientific code.
-The R$^0$ form, while easy to write and simple to understand is not a good scientific code while the full-fledge R$^{4}$ form is much preferred but requires a fair amount of work that might be unacceptable in some circumstances.
+Through a very simple random walk example using Python, we've seen a variety of pitfalls of writing scientific code.
+The R$^0$ form, while easy to write and simple to understand is not a good scientific code while the full-fledge R$^{4}$ form is much preferred but requires a fair amount of work that might be too much of a burden in some circumstances.
 $R^3$ form seems to be an acceptable compromise and should be accepted as the minimum scientific standard.

### 2017-06-13 main.tex
Reworked Reproducible (R^3).

@@
 \documentclass[a4paper,11pt]{article}
+% \usepackage[utf8]{inputenc}
 
@@
 
-%Replication is a cornerstone of science. 
-Reproducibility is a cornerstone of science. 
-% Since science aims to discover general principles of how the world functions, 
-% * <nicolas.rougier@inria.fr> 2017-04-24T15:49:43.902Z:
-% 
-% > Since science aims to discover general principles of how the world functions, 
-% 
-% 
-% C'est un peu une déclaration à l'emporte pièce et je soupçconne qu'il y a une littérature énorme sur la définition de la science.
-Et à moins de vouloir s'aliéner toute l'astophysioque et la cosmologie, on pourrait au minimum étendre à l'univers.
-% 
-% ^ <fabien.benureau+overleaf@gmail.com> 2017-04-25T13:29:01.309Z:
-% 
-% Reopened by editing the source!
-% From [wikipedia](https://en.wikipedia.org/wiki/World): "In a philosophical context, the world is the whole of the physical Universe, or an ontological world."
-% 
-% ^ <nicolas.rougier@inria.fr> 2017-04-25T13:45:04.565Z:
-%
-% Mouais.
-%
-% ^.
-If an experimental result cannot be obtained more than once,
+Replicability is a cornerstone of science. 
+If an experimental result cannot be re-obtained by an independent party,
 % it might be because of a fluke, a bias, a bug, a false positive or even a hack.
@@
 it merely becomes, at best, an observation that
-may feed future research \cite{Mesirov:2010,osc:2015}. 
+may feed future research (\cite{Mesirov:2010,osc:2015}). 
 Replication issues have received increased attention in recent years,
-in the domain of medicine and psychology in particular \cite{Iqbal:2016}.
+in the domain of medicine and psychology in particular (\cite{Iqbal:2016}).
+
 One could think that computer science would mostly be shielded from such issues.
@@
 Computer science is plagued by replication problems,
-in part, precisely because it seems impervious to them.\\
+in part, precisely because it seems impervious to them.
 
@@
 re-usable (R$^4$)
-and replicable (R$^5$).\\
+and replicable (R$^5$).
 
@@
 To be clear, it is impossible to write future-proof code,
-as the best efforts can be stymied by the smallest change in one of the dependencies.
+and the best efforts can be stymied by the smallest change in one of the dependencies.
 At the same time, modernizing an unmaintained ten-year-old code
@@
 Rather trying to predict the future or painstakingly dusting off old code, 
-an often easier solution is to recreate the old execution environment.
+an often easier and less dangerous solution is to recreate the old execution environment.
 For this to happen however, 
@@
 A \emph{re-runnable} code is one that describes an execution environment in which it is executable with enough details to be recreated.
-As shown by \citep{Collberg:2016}, this is far from being neither obvious nor easy.
+As shown by \citep{Collberg:2016}, this is far from being either obvious or easy.
 
@@
 This not the case because it uses the print {\em instruction} and the {\tt xrange} operator, both specific to Python 2.
-The print {\em instruction}, available in Python 2 (still widely used, will stop to be supported in 2020), 
+The print {\em instruction}, available in Python 2 (a version still widely used; support is scheduled to stop in 2020), 
 has been deprecated in Python 3 (first released in 2008, almost a decade ago)
@@
 In order to try to future-proof the code a bit, we might as well target Python 3, as is done in the R$^1$ version.
-Incidentally, it remains compatible with Python 2.\\
+Incidentally, it remains compatible with Python 2.
 
-But whatever the version chosen,
+But whatever version is chosen,
 the crucial step here is to document it.
@@
 
-% * <fabien.benureau+overleaf@gmail.com> 2017-04-27T06:18:15.753Z:
-% 
-% - 
-% 
-% Ok, so I am not sure that the parameter talk should be here. It should be in reproducible. 
-% 
-% ^.
-
 The code is running and producing the expected results. 
-The next step is to make sure that you can produce the same output over successive runs of your program, should you want to. 
+The next step is to make sure that you can produce the same output over successive runs of your program. 
 In other words, the next step is to make your program deterministic,  producing {\em repeatable} output. 
 Such repeatability is useful. 
-If a run of the program produces a particularly puzzling results,  it allows you to scrutinize any step of the execution of the program by re-running it again. 
+If a run of the program produces a particularly puzzling results,  it allows you to scrutinize any step of the execution of the program by re-running it again with extraneous prints, or inside a debugger.
 It is also the easiest way to prove that the program did indeed produce the published results.
@@
 For our program, that means setting the seed of the {\tt random} module.
-We may also want to save the output of the program to a file, so that we can easily verify that consecutive runs do produce the same output; eyeballing differences is unreliable and  time-consuming (and therefore won't be done systematically).
+We may also want to save the output of the program to a file, so that we can easily verify that consecutive runs do produce the same output: eyeballing differences is unreliable and  time-consuming---and therefore won't be done systematically.
 
@@
 Setting seeds is not without danger. 
-Using 439 as a seed in the previous program would result in ten consecutive +1 steps, which---although a perfectly valid random walk---lend itself to gross misinterpretations of the overall dynamic of the algorithm.
+Using 439 as a seed in the previous program would result in ten consecutive +1 steps, which---although a perfectly valid random walk---lend itself to a gross misinterpretation of the overall dynamic of the algorithm.
 Verifying that the qualitative aspects of the results and the conclusion we make are not tied to a specific initialization of the pseudo-random generator is an integral part of any scientific undertaking in computational science;
@@
 
-% \clearpage
+%\clearpage
 \section*{Reproducible (R$^{\mathbf 3}$)}
 
+% Reproducible definition + R^2 not reproducible
 The R$^2$ code seems fine enough, but it hides several problems that come to light when trying to {\em reproduce} results.
-In the strongest sense, a code is said to be {\em reproducible} if the same results it produced in the past can be re-obtained by someone else.
-As explained by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility is harder than you think. 
-For instance, our R$^2$ program will not produce the same results for everyone. It will, because it is repeatable, produce the same results over repeated executions.
-But it will not do so over different execution environments. The cause is found in a change that occurred in the pseudo-random number generator between CPython 3.2 and CPython 3.3. 
-Executed with CPython 3.2 or older, the code will produce the sequence 1, 2, 1, 2, 3, 4, 5, 6, 7, 6. But with CPython 3.3 or newer, it will produce 1, 2, 1, 0, 1, 0, 1, 0, -1, 0.
-Reproducibility impose to specify dependencies more precisely than re-runnability: the exact versions that were used must be documented.
-The platform should also be described as well: CPU architecture, operating system version, endianness, etc., as well as any defaults you may have set that might have an impact.
-Because of the increasing complexity of our computational stacks, retrieving and deciding what is pertinent might be difficult. A best effort must be made.\\
+A result is said to be \emph{reproducible} if another researcher can take the original code and input data, execute it, and re-obtain the result (\cite{Peng:2006}). 
+As explained by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility is hard. 
+For instance, the R$^2$ program will not produce the same results all the time. It will, because it is repeatable, produce the same results over repeated executions.
+But it will not do so over different execution environments. The cause is found in a change that occurred in the pseudo-random number generator between Python 3.2 and Python 3.3. 
+Executed with Python 2.7 to 3.2, the code will produce the sequence 1, 2, 1, 2, 3, 4, 5, 6, 7, 6. But with Python 3.3 to 3.6, it will produce 1, 2, 1, 0, 1, 0, 1, 0, -1, 0.\\
 
-Here, in addition to specifying the exact Python version and platform details, a short unit test is provided, that allows the code to verify its own reproducibility.
+% Reproducible definition + R^2 not reproducible
+Because any dependency of your program---to the most basic one, the language itself---can change its behavior from one version to another, executability (R$^1$) and determinism (R$^2$) are necessary but not sufficient for reproducibility.
+The exact execution environment used to produce the results must also be specified---rather than the broadest set of environments where the code can be executed.
+In other words, assertions such as "the results were obtained with Python 3.6.1" are more valuable, in a scientific context, than "the program should work with Python 3.3.x and above".\\
+
+% Reproducibility is hard
+Because of the increasing complexity of our computational stacks, retrieving and deciding what is pertinent (CPU architecture? Operating system version? Endianness?) might be difficult.
+See Appendix \ref{appendix:executionenvs} for a discussion.\\
+
+% Reproducibility in practice.
+Having precisely described the execution environment, reproducibility is technically obtained. However, \emph{reproducibility in practice} requires more than that.
+
+% Saving Parameters
+The R$^2$ program uses a random seed but does not keep a trace of it except in the code.
+Therefore, should the code change after the production of the results, someone provided with the result file will not be able to know which seed was used, and would need to iterate through all possible random seeds, which is not possible in practice.
+R$^3$ saves the seed in the name of the result file.
+More generally, results should always come alongside their context, i.e.
+an exhaustive list of the parameters used and a description of the execution environment.\\
+
+% Data behind graphs
+Graphs in articles are powerful way to communicate complex information.
+They are imprecise and cumbersome, and sometimes plain unusable, however when it comes to checking that the reproduced data match the one published in the article.
+A good figure should always be accompanied by its underlying data to allow straightforward comparisons.\\
+
+% Unit test for R^3
+Another good practice is to make the code self-verifiable.
+In R$^3$, a short unit test is provided, that allows the code to verify its own reproducibility.
 Should this test fail, then there is little hope of reproducing the results.
-If it does not fail, problems may still be present.
-A second problem present in the R$^2$ program is that it does not store the parameters used along with the results.
-Therefore, should the code change after the production of the results, someone provided with the result file will not be able to know which seed was used to produce it.
-In R$^3$, the seed is stored in the name of the result file.
+Of course, passing the test is no definitive proof.\\
+
+% Reproducibility implies availability 
+It should be obvious by now but \emph{reproducibility implies availability} As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request.
+While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable.
+Code \emph{and input and result files} should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as Figshare or Zenodo\footnote{Online code repository such as GitHub \emph{are not} scientific repositories, and may disappear, change name, etc.
+at any moment.
+Direct links to them are not perpetual and should be avoided.}.\\
 
@@
 \begin{code}{\textbf{\textsc{Listing 4:}} Random walk (R$^3$)}
-# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
+# Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
 # Release under the BSD 2-clause license
@@
 
-Finally, a crucial condition to making a code reproducible is making it available.
-As shown in \citep{Collberg:2016}, code is often not available, or available upon request.
-The latter may seem sufficient, but changes in email address,  career or retirement can make the code just as unavailable.
-Code \emph{and input and result files} should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as Figshare or Zenodo.
-Providing results files is important: it allows other researchers to easily compare the results they obtains with the one used to generate the figure in the article.
-Comparing figures directly can be cumbersome and often only allows to conclude about a qualitative similarity.\\
+% Summary
+To recap, reproducibility implies re-runnability or repeatability and availability, yet  imposes additional conditions.
+Dependencies and platforms must be described as precisely and a specifically as possible.
+Parameters values and inputs should be an integral part of the result files.
+The data behind graphs must be published.
+Unit tests are a good way to embed self-diagnostics of reproducibility in the code.
+Reproducibility is hard, yet tremendously necessary.
 
-To recap, reproducibility imposes stronger conditions than re-runnability.
-Dependencies and platforms must be described as precisely and a specifically as possible.
-Unit tests are a good way to embed self-diagnostics of reproducibility in the code.
-Results must be accompanied with the parameters and inputs used to produce them.
-And the code, and the data files behind the figure must be made available.
-
-% \clearpage
+\clearpage
 \section*{Re-usable (R$^{\mathbf 4}$)}
@@
 If people start using your program, they will most likely report bugs or malfunctions they encountered while using your program.
-If you're lucky enough, they will even propose you bugfix, hence improving the overall quality and correctness of your software.
+If you're lucky enough, they will even propose you bugfixes, hence improving the overall quality and correctness of your software.
 In other words, if your program is re-usable and is re-used, it is alive.
@@
 
+\newpage 
+\appendix
+
+\section{Appendix: Recording Execution Environments}
+\label{appendix:executionenvs}
+Talk about provenance-tracking and its problems (endianess, "as well as any defaults you may have set that might have an impact.")
+
+Talk about saving VMs, Docker images and its problems.
+
 \end{document}

### 2017-06-27 main.tex
Update on Overleaf.
reworked re-usable section (WIP)

@@
 \documentclass[a4paper,11pt]{article}
-% \usepackage[utf8]{inputenc}
+\usepackage[utf8]{inputenc}
 
@@
 The R$^2$ code seems fine enough, but it hides several problems that come to light when trying to {\em reproduce} results.
-A result is said to be \emph{reproducible} if another researcher can take the original code and input data, execute it, and re-obtain the result (\cite{Peng:2006}). 
+A result is said to be \emph{reproducible} if another researcher can take the original code and input data, execute it, and re-obtain the result \parencite{Peng:2006}.
+As explained by \citeauthor{Donoho:2009} \parencite{Donoho:2009}, scientific practice must expect that {\em errors are ubiquitous}, and therefore be robust to them.
+Ensuring reproducibility is an important step toward this: it allows other to verify that your code does indeed produce the published results, and scrutinize the way that it does.\\
+
 As explained by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility is hard. 
@@
 % Reproducibility implies availability 
-It should be obvious by now but \emph{reproducibility implies availability} As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request.
+It should be obvious by now but \emph{reproducibility implies availability}.
+As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request.
 While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable.
@@
 \section*{Re-usable (R$^{\mathbf 4}$)}
+
+Making your program re-usable means it can be easily used, and modified, by you and other people, inside and outside your lab.
+Ensuring your program is re-usable is advantageous for a number of reasons.\\
+
+For you, first.
+Because the you now and the you in two years are two different persons.
+Details on how to use the code, its limitations, its quirks, may be present to your mind now, but will probably escape you in six months \parencite{Donoho:2009}.
+Here, comments and documentation can make a significant difference.\\
+
+Furthermore, a source code reflects the results of all the decisions that were made to create it, but not the reasons behind them.
+In science, where the method and its justification matters as much as the results, those reasons are precious knowledge.
+For instance, it is not sufficient to observe that a piece of code is doing correct predictions on a learning task, one is interested by the exact algorithm that is employed, and why.
+In that context, a comment on how a given parameter was chosen (optimization, experimental data, educated guess, …) is a valuable information.\\
+
+Re-usability of course directly benefits others researchers, from your team and outside of it.
+The easier it is to use your code, the lower is the threshold for other to study, modify and extend it.
+Scientists constantly face the constraint of time: if a model is available, documented, and can be installed, run and understood all in a few hours, it will be preferred over another that would require weeks to reach the same stage.\\
+
+Since scientists have a tendency to base their work on results they can verify, a reproducible and re-usable code offers a platform both verified and easy-to-use, fostering the development of derivative work by other researchers.
+Those derivative works contribute to the impact of your original contribution, and generate citations.
+
+% * <fabien.benureau+overleaf@gmail.com> 2017-06-16T09:22:07.537Z:
+% 
+% I'd like to refactor this section into:
+% 1. Why re-usability is important for you (you will be stranger to yourself in six month)
+% 2. Why it is important for others (1. they will be more likely to cite your work, 2. long-term reproducibility)
+% 3. Tips and general guidelines on how to do it.
+% 
+% ^.
 

### 2017-06-29 main.tex
Update on Overleaf.

@@
 
-Since scientists have a tendency to base their work on results they can verify, a reproducible and re-usable code offers a platform both verified and easy-to-use, fostering the development of derivative work by other researchers.
-Those derivative works contribute to the impact of your original contribution, and generate citations.
+Moreover, scientists prefer to base their work on results they can verify.
+A reproducible and re-usable code offers a platform both verified and easy-to-use, fostering the development of derivative works by other researchers.
+Those derivative works contribute to the impact of your original contribution, and generate citations.\\
+
+Having more people examining and using your code also means that potential errors have a higher chance to be caught.
+If people start using your program, they will most likely report bugs or malfunctions they encountered while using your program.
+If you're lucky enough, they might even propose you either bugfixes or improvements, hence improving the overall quality of your software.
+This process contributes to the long-term reproducibility to the extent people continue to use and maintain the program.
 
@@
 % ^.
-
-Making your program re-usable means it can be used by people outside your lab,
-using possibly different environment, different parameters or different data.
-If people start using your program, they will most likely report bugs or malfunctions they encountered while using your program.
-If you're lucky enough, they will even propose you bugfixes, hence improving the overall quality and correctness of your software.
-In other words, if your program is re-usable and is re-used, it is alive.
-As such, and as for any other living being, it will evolve and mutate to adapt to its environment.
-This process will ensure long-term reproducibility to the extent people continue using and maintaining the program.\\
 

### 2017-07-03 main.tex
Layout change

@@
 
-Replicability is a cornerstone of science.
-If an experimental result cannot be re-obtained by an independent party,
-% it might be because of a fluke, a bias, a bug, a false positive or even a hack.
-%it becomes, at best, nothing more than food for thought for future research. 
-it merely becomes, at best, an observation that
-may feed future research (\cite{Mesirov:2010,osc:2015}). 
-Replication issues have received increased attention in recent years,
-in the domain of medicine and psychology in particular (\cite{Iqbal:2016}).
-
-One could think that computer science would mostly be shielded from such issues.
-But precisely because it is easy to believe that
-if a program runs once and gives expected results it will do so forever, 
-crucial steps to transform working code into meaningful scientific contributions are rarely undertaken \citep{Sandve:2013,Schwab:2000}. 
-Computer science is plagued by replication problems,
-in part, precisely because it seems impervious to them.
-
-In fact, a program can fail as a scientific contribution
-in many different ways for many different reasons.
-Borrowing the terms coined by \citeauthor{Goble:2016} \citep{Goble:2016},
-for your program to contribute to Science,
-you have to make sure it is
-re-runnable (R$^1$),
-repeatable (R$^2$),
-reproducible (R$^3$),
-re-usable (R$^4$)
-and replicable (R$^5$).
+Replicability is a cornerstone of science.  If an experimental result cannot be
+re-obtained by an independent party, it merely becomes, at best, an observation
+that may feed future research (\cite{Mesirov:2010,osc:2015}).  Replication
+issues have received increased attention in recent years, in the domain of
+medicine and psychology in particular (\cite{Iqbal:2016}). One could think that
+computer science would mostly be shielded from such issues.  But precisely
+because it is easy to believe that if a program runs once and gives expected
+results it will do so forever, crucial steps to transform working code into
+meaningful scientific contributions are rarely undertaken
+\citep{Sandve:2013,Schwab:2000}.  Computer science is plagued by replication
+problems, in part, precisely because it seems impervious to them.
+In fact, a program can fail as a scientific contribution in many different ways
+for many different reasons.  Borrowing the terms coined by
+\citeauthor{Goble:2016} \citep{Goble:2016}, for your program to contribute to
+Science, you have to make sure it is re-runnable (R$^1$), repeatable (R$^2$),
+reproducible (R$^3$), re-usable (R$^4$) and replicable (R$^5$).
 
@@
 % -------------------------
-\clearpage
+% \clearpage
 \section*{Re-runnable (R$^{\mathbf 1}$)}
 
-Have you ever try to re-run a program you wrote some years ago?
-Frustratingly, it can often be surprisingly hard. 
-Part of the problem is that technology is evolving at a fast pace
-and you cannot  know in advance
-how the system, the software and the libraries your program depends on will evolve.
-Since you wrote the code,
-you may have reinstalled or upgraded  your operating system.
-The compiler, interpreter or set of libraries installed may have been replaced with newer versions. 
-You may find yourself battling with
-arcane issues of library compatibility---thoroughly orthogonal to your immediate research  goals---to execute again a code \emph{that worked perfectly before}. 
-% "thoroughly orthogonal to your immediate research goals": possibly problematic
+Have you ever try to re-run a program you wrote some years ago?  Frustratingly,
+it can often be surprisingly hard.  Part of the problem is that technology is
+evolving at a fast pace and you cannot know in advance how the system, the
+software and the libraries your program depends on will evolve.  Since you
+wrote the code, you may have reinstalled or upgraded your operating system.
+The compiler, interpreter or set of libraries installed may have been replaced
+with newer versions.  You may find yourself battling with arcane issues of
+library compatibility---thoroughly orthogonal to your immediate research
+goals---to execute again a code \emph{that worked perfectly before}.  To be
+clear, it is impossible to write future-proof code, and the best efforts can be
+stymied by the smallest change in one of the dependencies.  At the same time,
+modernizing an unmaintained ten-year-old code can reveal itself to be an
+arduous and expensive undertaking, with each change risking to affect the
+semantics of the program. Rather than trying to predict the future or painstakingly
+dusting off old code, an often easier and less dangerous solution is to
+recreate the old execution environment.  For this to happen however, the
+dependencies in terms of systems, software and libraries must be made clear
+enough.
 
-To be clear, it is impossible to write future-proof code,
-and the best efforts can be stymied by the smallest change in one of the dependencies.
-At the same time, modernizing an unmaintained ten-year-old code
-can reveal itself to be an arduous and expensive undertaking,
-with each change risking to affect the semantics of the program.
-
-Rather trying to predict the future or painstakingly dusting off old code, 
-an often easier and less dangerous solution is to recreate the old execution environment.
-For this to happen however, 
-the dependencies in terms of systems, software and libraries must be made clear enough.
-
-A \emph{re-runnable} code is one that describes an execution environment in which it is executable with enough details to be recreated.
-As shown by \citep{Collberg:2016}, this is far from being either obvious or easy.
+A \emph{re-runnable} code is one that describes an execution environment in
+which it is executable with enough details to be recreated.  As shown by
+\citep{Collberg:2016}, this is far from being either obvious or easy.
 
@@
 
-In our case, the R$^0$ version of our tiny walker seems to imply
-that any version of Python would be fine.
-This not the case because it uses the print {\em instruction} and the {\tt xrange} operator, both specific to Python 2.
-The print {\em instruction}, available in Python 2 (a version still widely used; support is scheduled to stop in 2020), 
-has been deprecated in Python 3 (first released in 2008, almost a decade ago)
-in favor or a  print {\em function}.
-The {\tt xrange} operator has been replaced by the {\tt range} operator in Python 3.
-In order to try to future-proof the code a bit, we might as well target Python 3, as is done in the R$^1$ version.
-Incidentally, it remains compatible with Python 2.
-
-But whatever version is chosen,
-the crucial step here is to document it.
+In our case, the R$^0$ version of our tiny walker seems to imply that any
+version of Python would be fine.  This not the case because it uses the print
+{\em instruction} and the {\tt xrange} operator, both specific to Python 2.
+The print {\em instruction}, available in Python 2 (a version still widely
+used; support is scheduled to stop in 2020), has been deprecated in Python 3
+(first released in 2008, almost a decade ago) in favor or a print {\em
+  function}.  The {\tt xrange} operator has been replaced by the {\tt range}
+operator in Python 3.  In order to try to future-proof the code a bit, we might
+as well target Python 3, as is done in the R$^1$ version.  Incidentally, it
+remains compatible with Python 2. But whatever version is chosen, the crucial
+step here is to document it.
 
@@
 
-The code is running and producing the expected results. 
-The next step is to make sure that you can produce the same output over successive runs of your program. 
-In other words, the next step is to make your program deterministic,  producing {\em repeatable} output. 
-Such repeatability is useful. 
-If a run of the program produces a particularly puzzling results,  it allows you to scrutinize any step of the execution of the program by re-running it again with extraneous prints, or inside a debugger.
-It is also the easiest way to prove that the program did indeed produce the published results.
-Such repeatability is not always possible or easy \citep{Diethelm:2012, Courtes:2015}.
-But for sequential programs not depending on exotic hardware, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\
-
-For our program, that means setting the seed of the {\tt random} module.
-We may also want to save the output of the program to a file, so that we can easily verify that consecutive runs do produce the same output: eyeballing differences is unreliable and  time-consuming---and therefore won't be done systematically.
+The code is running and producing the expected results.  The next step is to
+make sure that you can produce the same output over successive runs of your
+program.  In other words, the next step is to make your program deterministic,
+producing {\em repeatable} output.  Such repeatability is useful.  If a run of
+the program produces a particularly puzzling results, it allows you to
+scrutinize any step of the execution of the program by re-running it again with
+extraneous prints, or inside a debugger.  It is also the easiest way to prove
+that the program did indeed produce the published results.  Such repeatability
+is not always possible or easy \citep{Diethelm:2012, Courtes:2015}.  But for
+sequential programs not depending on exotic hardware, it often comes down to
+controlling the initialization of the pseudo-random number generators (RNG).
+For our program, that means setting the seed of the {\tt random} module. We may
+also want to save the output of the program to a file, so that we can easily
+verify that consecutive runs do produce the same output: eyeballing differences
+is unreliable and time-consuming---and therefore won't be done systematically.
 
@@
 
-Setting seeds is not without danger. 
-Using 439 as a seed in the previous program would result in ten consecutive +1 steps, which---although a perfectly valid random walk---lend itself to a gross misinterpretation of the overall dynamic of the algorithm.
-Verifying that the qualitative aspects of the results and the conclusion we make are not tied to a specific initialization of the pseudo-random generator is an integral part of any scientific undertaking in computational science;
-this is usually done by repeating the simulations multiple times with different seeds.
-
+Setting seeds is not without danger.  Using 439 as a seed in the previous
+program would result in ten consecutive +1 steps, which---although a perfectly
+valid random walk---lend itself to a gross misinterpretation of the overall
+dynamic of the algorithm.  Verifying that the qualitative aspects of the
+results and the conclusion we make are not tied to a specific initialization of
+the pseudo-random generator is an integral part of any scientific undertaking
+in computational science; this is usually done by repeating the simulations
+multiple times with different seeds.
 %Determinism is a tool that allow to study bugs and spurious or unexpected behavior. 
@@
 % talk about controversy about setting seeds
-
 % For example, you can play with some parameters and
@@
 % Reproducible definition + R^2 not reproducible
-The R$^2$ code seems fine enough, but it hides several problems that come to light when trying to {\em reproduce} results.
-A result is said to be \emph{reproducible} if another researcher can take the original code and input data, execute it, and re-obtain the result \parencite{Peng:2006}.
-As explained by \citeauthor{Donoho:2009} \parencite{Donoho:2009}, scientific practice must expect that {\em errors are ubiquitous}, and therefore be robust to them.
-Ensuring reproducibility is an important step toward this: it allows other to verify that your code does indeed produce the published results, and scrutinize the way that it does.\\
+The R$^2$ code seems fine enough, but it hides several problems that come to
+light when trying to {\em reproduce} results.  A result is said to be
+\emph{reproducible} if another researcher can take the original code and input
+data, execute it, and re-obtain the result \parencite{Peng:2006}. As explained
+by \citeauthor{Donoho:2009} \parencite{Donoho:2009}, scientific practice must
+expect that {\em errors are ubiquitous}, and therefore be robust to
+them. Ensuring reproducibility is an important step toward this: it allows
+other to verify that your code does indeed produce the published results, and
+scrutinize the way that it does.\\
 
-As explained by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility is hard. 
-For instance, the R$^2$ program will not produce the same results all the time. It will, because it is repeatable, produce the same results over repeated executions.
-But it will not do so over different execution environments. The cause is found in a change that occurred in the pseudo-random number generator between Python 3.2 and Python 3.3. 
-Executed with Python 2.7 to 3.2, the code will produce the sequence 1, 2, 1, 2, 3, 4, 5, 6, 7, 6. But with Python 3.3 to 3.6, it will produce 1, 2, 1, 0, 1, 0, 1, 0, -1, 0.\\
-
+As explained by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility
+is hard.  For instance, the R$^2$ program will not produce the same results all
+the time. It will, because it is repeatable, produce the same results over
+repeated executions.  But it will not do so over different execution
+environments. The cause is found in a change that occurred in the pseudo-random
+number generator between Python 3.2 and Python 3.3.  Executed with Python 2.7
+to 3.2, the code will produce the sequence 1, 2, 1, 2, 3, 4, 5, 6, 7, 6. But
+with Python 3.3 to 3.6, it will produce 1, 2, 1, 0, 1, 0, 1, 0, -1, 0.
 % Reproducible definition + R^2 not reproducible
-Because any dependency of your program---to the most basic one, the language itself---can change its behavior from one version to another, executability (R$^1$) and determinism (R$^2$) are necessary but not sufficient for reproducibility.
-The exact execution environment used to produce the results must also be specified---rather than the broadest set of environments where the code can be executed.
-In other words, assertions such as "the results were obtained with Python 3.6.1" are more valuable, in a scientific context, than "the program should work with Python 3.3.x and above".\\
-
+Because any dependency of your program---to the most basic one, the language
+itself---can change its behavior from one version to another, executability
+(R$^1$) and determinism (R$^2$) are necessary but not sufficient for
+reproducibility. The exact execution environment used to produce the results
+must also be specified---rather than the broadest set of environments where the
+code can be executed. In other words, assertions such as "the results were
+obtained with Python 3.6.1" are more valuable, in a scientific context, than
+"the program should work with Python 3.3.x and above".
 % Reproducibility is hard
-Because of the increasing complexity of our computational stacks, retrieving and deciding what is pertinent (CPU architecture? Operating system version? Endianness?) might be difficult.
-See Appendix \ref{appendix:executionenvs} for a discussion.\\
-
+Because of the increasing complexity of our computational stacks, retrieving
+and deciding what is pertinent (CPU architecture? Operating system version?
+Endianness?) might be difficult.  See Appendix \ref{appendix:executionenvs} for
+a discussion.
 % Reproducibility in practice.
 Having precisely described the execution environment, reproducibility is technically obtained. However, \emph{reproducibility in practice} requires more than that.
-
 % Saving Parameters
-The R$^2$ program uses a random seed but does not keep a trace of it except in the code.
-Therefore, should the code change after the production of the results, someone provided with the result file will not be able to know which seed was used, and would need to iterate through all possible random seeds, which is not possible in practice.
-R$^3$ saves the seed in the name of the result file.
-More generally, results should always come alongside their context, i.e.
-an exhaustive list of the parameters used and a description of the execution environment.\\
-
+The R$^2$ program uses a random seed but does not keep a trace of it except in
+the code. Therefore, should the code change after the production of the
+results, someone provided with the result file will not be able to know which
+seed was used, and would need to iterate through all possible random seeds,
+which is not possible in practice. R$^3$ saves the seed in the name of the
+result file. More generally, results should always come alongside their
+context, i.e. an exhaustive list of the parameters used and a description of
+the execution environment.
 % Data behind graphs
-Graphs in articles are powerful way to communicate complex information.
-They are imprecise and cumbersome, and sometimes plain unusable, however when it comes to checking that the reproduced data match the one published in the article.
-A good figure should always be accompanied by its underlying data to allow straightforward comparisons.\\
-
+Graphs in articles are powerful way to communicate complex information. They
+are imprecise and cumbersome, and sometimes plain unusable, however when it
+comes to checking that the reproduced data match the one published in the
+article. A good figure should always be accompanied by its underlying data to
+allow straightforward comparisons.
 % Unit test for R^3
-Another good practice is to make the code self-verifiable.
-In R$^3$, a short unit test is provided, that allows the code to verify its own reproducibility.
-Should this test fail, then there is little hope of reproducing the results.
-Of course, passing the test is no definitive proof.\\
+Another good practice is to make the code self-verifiable. In R$^3$, a short
+unit test is provided, that allows the code to verify its own
+reproducibility. Should this test fail, then there is little hope of
+reproducing the results. Of course, passing the test is no definitive proof.\\
 
 % Reproducibility implies availability 
-It should be obvious by now but \emph{reproducibility implies availability}.
-As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request.
-While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable.
-Code \emph{and input and result files} should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as Figshare or Zenodo\footnote{Online code repository such as GitHub \emph{are not} scientific repositories, and may disappear, change name, etc.
-at any moment.
-Direct links to them are not perpetual and should be avoided.}.\\
-
+It should be obvious by now but let us state it very explicitely:
+\emph{reproducibility implies availability}. As shown in \citep{Collberg:2016},
+code is often unavailable, or only available upon request. While the latter may
+seem sufficient, changes in email address, changes in career, retirement, a
+busy inbox or poor archiving practices can make a code just as
+unreachable. Code \emph{and input and result files} should be available with
+the published article, as supplementary data, or through a DOI link to a
+scientific repository such as Figshare or Zenodo\footnote{Online code
+  repository such as GitHub \emph{are not} scientific repositories, and may
+  disappear, change name, etc. at any moment. Direct links to them are not
+  perpetual and should be avoided.}.
 %\begin{itemize}
@@
 % Summary
-To recap, reproducibility implies re-runnability or repeatability and availability, yet  imposes additional conditions.
-Dependencies and platforms must be described as precisely and a specifically as possible.
-Parameters values and inputs should be an integral part of the result files.
-The data behind graphs must be published.
-Unit tests are a good way to embed self-diagnostics of reproducibility in the code.
-Reproducibility is hard, yet tremendously necessary.
+To recap, reproducibility implies re-runnability or repeatability and
+availability, yet imposes additional conditions. Dependencies and platforms
+must be described as precisely and a specifically as possible. Parameters
+values and inputs should be an integral part of the result files. The data
+behind graphs must be published. Unit tests are a good way to embed
+self-diagnostics of reproducibility in the code. Reproducibility is hard, yet
+tremendously necessary.
 
-\clearpage
+% \clearpage
 \section*{Re-usable (R$^{\mathbf 4}$)}
 
-Making your program re-usable means it can be easily used, and modified, by you and other people, inside and outside your lab.
-Ensuring your program is re-usable is advantageous for a number of reasons.\\
+Making your program re-usable means it can be easily used, and modified, by you
+and other people, inside and outside your lab. Ensuring your program is
+re-usable is advantageous for a number of reasons. For you, first. Because the
+you now and the you in two years (or two months in some cases) are two
+different persons. Details on how to use the code, its limitations, its quirks,
+may be present to your mind now, but will probably escape you in six months
+\parencite{Donoho:2009}. Here, comments and documentation can make a
+significant difference.  Furthermore, a source code reflects the results of all
+the decisions that were made to create it, but not the reasons behind them. In
+science, where the method and its justification matters as much as the results,
+those reasons are precious knowledge. For instance, it is not sufficient to
+observe that a piece of code is doing correct predictions on a learning task,
+one is interested by the exact algorithm that is employed, and why. In that
+context, a comment on how a given parameter was chosen (optimization,
+experimental data, educated guess, …) is a valuable information.
+%
+Re-usability of course directly benefits others researchers, from your team and
+outside of it. The easier it is to use your code, the lower is the threshold
+for other to study, modify and extend it. Scientists constantly face the
+constraint of time: if a model is available, documented, and can be installed,
+run and understood all in a few hours, it will be preferred over another that
+would require weeks to reach the same stage.  Moreover, scientists prefer to
+base their work on results they can verify. A reproducible and re-usable code
+offers a platform both verified and easy-to-use, fostering the development of
+derivative works by other researchers.\\
 
-For you, first.
-Because the you now and the you in two years are two different persons.
-Details on how to use the code, its limitations, its quirks, may be present to your mind now, but will probably escape you in six months \parencite{Donoho:2009}.
-Here, comments and documentation can make a significant difference.\\
-
-Furthermore, a source code reflects the results of all the decisions that were made to create it, but not the reasons behind them.
-In science, where the method and its justification matters as much as the results, those reasons are precious knowledge.
-For instance, it is not sufficient to observe that a piece of code is doing correct predictions on a learning task, one is interested by the exact algorithm that is employed, and why.
-In that context, a comment on how a given parameter was chosen (optimization, experimental data, educated guess, …) is a valuable information.\\
-
-Re-usability of course directly benefits others researchers, from your team and outside of it.
-The easier it is to use your code, the lower is the threshold for other to study, modify and extend it.
-Scientists constantly face the constraint of time: if a model is available, documented, and can be installed, run and understood all in a few hours, it will be preferred over another that would require weeks to reach the same stage.\\
-
-Moreover, scientists prefer to base their work on results they can verify.
-A reproducible and re-usable code offers a platform both verified and easy-to-use, fostering the development of derivative works by other researchers.
-Those derivative works contribute to the impact of your original contribution, and generate citations.\\
-
-Having more people examining and using your code also means that potential errors have a higher chance to be caught.
-If people start using your program, they will most likely report bugs or malfunctions they encountered while using your program.
-If you're lucky enough, they might even propose you either bugfixes or improvements, hence improving the overall quality of your software.
-This process contributes to the long-term reproducibility to the extent people continue to use and maintain the program.
-
-% * <fabien.benureau+overleaf@gmail.com> 2017-06-16T09:22:07.537Z:
+Those derivative works contribute to the impact of your original contribution,
+and generate citations. Having more people examining and using your code also
+means that potential errors have a higher chance to be caught. If people start
+using your program, they will most likely report bugs or malfunctions they
+encountered while using your program. If you're lucky enough, they might even
+propose you either bugfixes or improvements, hence improving the overall
+quality of your software. This process contributes to the long-term
+reproducibility to the extent people continue to use and maintain the program.
 %
-% I'd like to refactor this section into:
-% 1. Why re-usability is important for you (you will be stranger to yourself in six month)
-% 2. Why it is important for others (1. they will be more likely to cite your work, 2. long-term reproducibility)
-% 3. Tips and general guidelines on how to do it.
+However, making your program to be re-usable by others requires a little
+work. If it is natural, when one writes a program for oneself, to make implicit
+choices that suit one needs, such implicit choices won't resist the will of a
+user community. For example, having magic numbers or ill-defined constants in
+your program might be cumbersome for others. Similarly, it might be acceptable
+to not have a proper documentation when you're the only user (actually it might
+not be acceptable because the you now and the you in six months are two
+different persons), but other users want to know precisely how to use this or
+that function and what it does exactly.
 %
-% ^.
-
-However, making your program to be re-usable by others requires a little work.
-If it is natural, when one writes a program for oneself, to make implicit choices that suit one needs, such implicit choices won't resist the will of a user community.
-For example, having magic numbers or ill-defined constants in your program might be cumbersome for others.
-Similarly, it might be acceptable to not have a proper documentation when you're the only user (actually it might not be acceptable because the you now and the you in six months are two different persons), but other users want to know precisely how to use this or that function and what it does exactly.\\
-
-If we look back at our random walker, we can list all the implicit choices that have been made:
+If we look back at our random walker, we can list all the implicit choices that
+have been made:
 \begin{itemize}
@@
 \item Walker walk along a single dimension
-\item ...
+\item etc.
 \end{itemize}
@@
 
-Having made a software re-usable by others should offer a decent protection against bugs.
-Unfortunately, this is not always the case and some recent cases have demonstrated the tremendous impact a bug can have in Science \citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}.
-This is why, as explained by Peng et al.
-\cite{Peng:2006}, {\em the replication of important findings by multiple independent investigators is fundamental to the accumulation of scientific evidence}.\\
+Having made a software re-usable by others should offer a decent protection
+against bugs.  Unfortunately, this is not always the case and some recent cases
+have demonstrated the tremendous impact a bug can have in Science
+\citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}.  This is
+why, as explained by Peng et al. \cite{Peng:2006}, {\em the replication of
+  important findings by multiple independent investigators is fundamental to
+  the accumulation of scientific evidence}.\\
 % Replicability is the assumption that any article that does not provide the code source makes: that the description it provides of the algorithms is sufficiently precise and complete to re-obtain the results it presents.
 
-In our case, the replication of the random walk brings some unexpected discrepancies between the original Python version and the replicated Numpy version.
-The reason is to be found in a differences in the respective implementations of the RNG between Python and Numpy.
-As explained in the \href{https://docs.python.org/3.6/library/random.html}{Python documentation}, {\em Python uses the Mersenne Twister as the core generator. It produces 53-bit precision floats
-and has a period of 2**19937-1.
-The underlying implementation in C is both fast and threadsafe.}
-Numpy RNG is also based on the Mersenne Twister generator but there are differences in the way seed is interpreted when initializing the generator.
-Fortunately, Both Numpy and Python offer access to the internal state and we can {\em fix} it to make the behavior of one RNG to match the other RNG.
-Here, we want to make the Numpy RNG to match the Python RNG behavior.\\
-
-Without such explicit replication effort, it would be very hard to discover this faulty behavior and this why the ReScience journal has been created.
-This journal targets computational research and encourages the explicit replication of already published research, promoting new and open-source implementations in order to ensure that the original research is reproducible.
+In our case, the replication of the random walk brings some unexpected
+discrepancies between the original Python version and the replicated Numpy
+version.  The reason is to be found in a differences in the respective
+implementations of the RNG between Python and Numpy.  As explained in the
+\href{https://docs.python.org/3.6/library/random.html}{Python documentation},
+     {\em Python uses the Mersenne Twister as the core generator. It produces
+      53-bit precision floats and has a period of 2**19937-1.  The underlying
+       implementation in C is both fast and threadsafe.}  Numpy RNG is also
+     based on the Mersenne Twister generator but there are differences in the
+     way seed is interpreted when initializing the generator.  Fortunately,
+     Both Numpy and Python offer access to the internal state and we can {\em
+       fix} it to make the behavior of one RNG to match the other RNG.  Here,
+     we want to make the Numpy RNG to match the Python RNG behavior.
+%
+Without such explicit replication effort, it would be very hard to discover
+this faulty behavior and this why the ReScience journal has been created.  This
+journal targets computational research and encourages the explicit replication
+of already published research, promoting new and open-source implementations in
+order to ensure that the original research is reproducible.
 
@@
 \appendix
-
 \section{Appendix: Recording Execution Environments}
 \label{appendix:executionenvs}
-Talk about provenance-tracking and its problems (endianess, "as well as any defaults you may have set that might have an impact.")
+%% Talk about provenance-tracking and its problems (endianess, "as well as any defaults you may have set that might have an impact.")
 
-Talk about saving VMs, Docker images and its problems.
+%% Talk about saving VMs, Docker images and its problems.
 

### 2017-07-06 main.tex
Update on replicable (WIP)
I did not layout my lines correctly, I am afraid.

@@
 
-Having made a software re-usable by others should offer a decent protection
-against bugs.  Unfortunately, this is not always the case and some recent cases
-have demonstrated the tremendous impact a bug can have in Science
-\citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}.  This is
-why, as explained by Peng et al. \cite{Peng:2006}, {\em the replication of
-  important findings by multiple independent investigators is fundamental to
-  the accumulation of scientific evidence}.\\
-% Replicability is the assumption that any article that does not provide the code source makes: that the description it provides of the algorithms is sufficiently precise and complete to re-obtain the results it presents.
+Having made a software re-usable by others should offer a decent protection against bugs.
+Unfortunately, this is not always the case and some recent cases have demonstrated the tremendous impact a bug can have in Science \citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}.
+This is why, as explained by Peng et al.
+\cite{Peng:2006}, {\em the replication of important findings by multiple independent investigators is fundamental to the accumulation of scientific evidence}.\\
 
-In our case, the replication of the random walk brings some unexpected
-discrepancies between the original Python version and the replicated Numpy
-version.  The reason is to be found in a differences in the respective
-implementations of the RNG between Python and Numpy.  As explained in the
-\href{https://docs.python.org/3.6/library/random.html}{Python documentation},
-     {\em Python uses the Mersenne Twister as the core generator. It produces
-      53-bit precision floats and has a period of 2**19937-1.  The underlying
-       implementation in C is both fast and threadsafe.}  Numpy RNG is also
-     based on the Mersenne Twister generator but there are differences in the
-     way seed is interpreted when initializing the generator.  Fortunately,
-     Both Numpy and Python offer access to the internal state and we can {\em
-       fix} it to make the behavior of one RNG to match the other RNG.  Here,
-     we want to make the Numpy RNG to match the Python RNG behavior.
-%
-Without such explicit replication effort, it would be very hard to discover
-this faulty behavior and this why the ReScience journal has been created.  This
-journal targets computational research and encourages the explicit replication
-of already published research, promoting new and open-source implementations in
-order to ensure that the original research is reproducible.
+Replicability is the assumption that any article that does not provide the code source makes: that the description it provides of the algorithms is sufficiently precise and complete to re-obtain the results it presents.
+While every published article should strive for replicability, even the best efforts may fall short because an oversight, a typo or a difference between what is evident between the writer and the reader.
+In that case, a reproducible code is a crucial backup.
+Hence, replicability does not negate the necessity of reproducibility.
+
+In our case, the replication of the random walk brings some unexpected discrepancies between the original Python version and the replicated Numpy version.
+The reason is to be found in a differences in the respective implementations of the RNG between Python and Numpy.
+As explained in the \href{https://docs.python.org/3.6/library/random.html}{Python documentation}, {\em Python uses the Mersenne Twister as the core generator. It produces 53-bit precision floats
+and has a period of 2**19937-1.
+The underlying implementation in C is both fast and threadsafe.}
+Numpy RNG is also based on the Mersenne Twister generator but there are differences in the way seed is interpreted when initializing the generator.
+Fortunately, Both Numpy and Python offer access to the internal state and we can {\em fix} it to make the behavior of one RNG to match the other RNG.
+Here, we want to make the Numpy RNG to match the Python RNG behavior.
+
+In our case, a perfect quantitative match may seem unnecessary.
+But in many replication projects, quantitative comparisons are an objective tests that the behavior is reproduced.
+Statistical tests demand more work, and may fall short of detecting small differences in the algorithms, that may results in huge difference of behavior in different conditions.
+In all cases, eyeballing difference is graphs is always difficult.
+Therefore, one should always provide the data behind the graphs to allow such comparisons.
+
+Without such explicit replication effort, it would be hard to discover this faulty behavior and this why the ReScience journal has been created.
+This journal targets computational research and encourages the explicit replication of already published research, promoting new and open-source implementations in order to ensure that the original research is reproducible.
 

### 2017-07-07 draft.tex
Layout change

--- 2017-07-06 main.tex
@@
 \DeclareLanguageMapping{american}{american-apa}
-\bibliography{library.bib}
+\bibliography{draft.bib}
 
@@
 % \title{The R Quintuplet (R$^5$)}
-
-\author{Nicolas P. Rougier and Fabien C. Y. Benureau}
-\date{\today}
+% \author{Nicolas P. Rougier and Fabien C. Y. Benureau}
+\author{
+  \textbf{Nicolas P. Rougier}$^{1,2,3,*}$, \textbf{Fabien C. Y. Benureau}$^{1,2,3}$\\
+  \begin{minipage}{\textwidth}
+    \begin{center}
+      \vspace{2mm}
+      $^{1}$INRIA Bordeaux Sud-Ouest Talence, France $^{2}$ Institut des
+      Maladies Neurodégénératives, Université de Bordeaux, CNRS UMR 5293,
+      Bordeaux, France $^{3}$ LaBRI, Université de Bordeaux, Bordeaux INP, CNRS
+      UMR 5800, Talence, France\\
+      \vspace{2mm}
+      $^{*}$Corresponding author:
+      \href{mailto:Nicolas.Rougier@inria.fr}{Nicolas.Rougier@inria.fr}         
+    \end{center}
+  \end{minipage}
+}
+\date{}
 
@@
 \citep{Sandve:2013,Schwab:2000}. Computer science is plagued by replication
-problems, in part, precisely because it seems impervious to them.
-In fact, a program can fail as a scientific contribution in many different ways
-for many different reasons.  Borrowing the terms coined by
-\citeauthor{Goble:2016} \citep{Goble:2016}, for your program to contribute to
-Science, you have to make sure it is re-runnable (R$^1$), repeatable (R$^2$),
-reproducible (R$^3$), re-usable (R$^4$) and replicable (R$^5$).
+problems, in part, precisely because it seems impervious to them. In fact, a
+program can fail as a scientific contribution in many different ways for many
+different reasons. Borrowing the terms coined by \citeauthor{Goble:2016}
+\citep{Goble:2016}, for your program to contribute to Science, you have to make
+sure it is re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$),
+re-usable (R$^4$) and replicable (R$^5$).\\
 
-Let us illustrate this with a very simple example,
-a random walk written in Python:
+Let us illustrate this with a very simple example, a random walk written in Python:\\
 
+\noindent \begin{minipage}[c]{\linewidth}
 \begin{code}{\textbf{\textsc{Listing 1:}} Random walk (R$^0$)}
@@
 \end{code}
+\end{minipage}
+
 % That would be too easy if the program was commented...
@@
 % -------------------------
-% \clearpage
+\clearpage
 \section*{Re-runnable (R$^{\mathbf 1}$)}
 
-Have you ever try to re-run a program you wrote some years ago?  Frustratingly,
-it can often be surprisingly hard.  Part of the problem is that technology is
-evolving at a fast pace and you cannot know in advance how the system, the
-software and the libraries your program depends on will evolve.  Since you
-wrote the code, you may have reinstalled or upgraded your operating system.
-The compiler, interpreter or set of libraries installed may have been replaced
-with newer versions.  You may find yourself battling with arcane issues of
-library compatibility---thoroughly orthogonal to your immediate research
-goals---to execute again a code \emph{that worked perfectly before}.  To be
-clear, it is impossible to write future-proof code, and the best efforts can be
-stymied by the smallest change in one of the dependencies.  At the same time,
-modernizing an unmaintained ten-year-old code can reveal itself to be an
-arduous and expensive undertaking, with each change risking to affect the
-semantics of the program. Rather than trying to predict the future or painstakingly
-dusting off old code, an often easier and less dangerous solution is to
-recreate the old execution environment.  For this to happen however, the
-dependencies in terms of systems, software and libraries must be made clear
-enough.
+Have you ever try to re-run a program you wrote some years ago?
+Frustratingly, it can often be surprisingly hard.
+Part of the problem is that technology is evolving at a fast pace and you cannot  know in advance how the system, the software and the libraries your program depends on will evolve.
+Since you wrote the code, you may have reinstalled or upgraded  your operating system.
+The compiler, interpreter or set of libraries installed may have been replaced with newer versions.
+You may find yourself battling with arcane issues of library compatibility---thoroughly orthogonal to your immediate research  goals---to execute again a code \emph{that worked perfectly before}.
+To be clear, it is impossible to write future-proof code, and the best efforts can be stymied by the smallest change in one of the dependencies.
+At the same time, modernizing an unmaintained ten-year-old code can reveal itself to be an arduous and expensive undertaking, with each change risking to affect the semantics of the program.
+Rather trying to predict the future or painstakingly dusting off old code, an often easier and less dangerous solution is to recreate the old execution environment.
+For this to happen however, the dependencies in terms of systems, software and libraries must be made clear enough.\\
 
-A \emph{re-runnable} code is one that describes an execution environment in
-which it is executable with enough details to be recreated.  As shown by
-\citep{Collberg:2016}, this is far from being either obvious or easy.
+A \emph{re-runnable} code is one that describes an execution environment in which it is executable with enough details to be recreated.
+As shown by \citep{Collberg:2016}, this is far from being either obvious or easy.\\
 
-% * <fabien.benureau+overleaf@gmail.com> 2017-05-08T18:09:47.037Z:
-% 
-% Wouldn't Collberg be better cited in R^2?
-% 
-% ^.
-% if using random.choices, won't be runnable with anything <3.6
-% if using itertools.accumulate, won't be runnable with anything <3.2
-% although perhaps not the best syntax, we might want to keep the
-% import mod
-% mod.function 
-% syntax for both random and intertools
-%
-% Problem with current code is that it is highly inefficient now (compared to R0)
+\noindent \begin{minipage}[c]{\linewidth}
 \begin{code}{\textbf{\textsc{Listing 2:}} Random walk (R$^1$)}
@@
 \end{code}
+\end{minipage}\\
 
-In our case, the R$^0$ version of our tiny walker seems to imply that any
-version of Python would be fine.  This not the case because it uses the print
-{\em instruction} and the {\tt xrange} operator, both specific to Python 2.
-The print {\em instruction}, available in Python 2 (a version still widely
-used; support is scheduled to stop in 2020), has been deprecated in Python 3
-(first released in 2008, almost a decade ago) in favor or a print {\em
-  function}.  The {\tt xrange} operator has been replaced by the {\tt range}
-operator in Python 3.  In order to try to future-proof the code a bit, we might
-as well target Python 3, as is done in the R$^1$ version.  Incidentally, it
-remains compatible with Python 2. But whatever version is chosen, the crucial
-step here is to document it.
+In our case, the R$^0$ version of our tiny walker seems to imply that any version of Python would be fine.
+This not the case because it uses the print {\em instruction} and the {\tt xrange} operator, both specific to Python 2.
+The print {\em instruction}, available in Python 2 (a version still widely used; support is scheduled to stop in 2020), has been deprecated in Python 3 (first released in 2008, almost a decade ago) in favor or a  print {\em function}.
+The {\tt xrange} operator has been replaced by the {\tt range} operator in Python 3.
+In order to try to future-proof the code a bit, we might as well target Python 3, as is done in the R$^1$ version.
+Incidentally, it remains compatible with Python 2.
+But whatever version is chosen, the crucial step here is to document it.
 
@@
 
-The code is running and producing the expected results.  The next step is to
-make sure that you can produce the same output over successive runs of your
-program.  In other words, the next step is to make your program deterministic,
-producing {\em repeatable} output.  Such repeatability is useful.  If a run of
-the program produces a particularly puzzling results, it allows you to
-scrutinize any step of the execution of the program by re-running it again with
-extraneous prints, or inside a debugger.  It is also the easiest way to prove
-that the program did indeed produce the published results.  Such repeatability
-is not always possible or easy \citep{Diethelm:2012, Courtes:2015}.  But for
-sequential programs not depending on exotic hardware, it often comes down to
-controlling the initialization of the pseudo-random number generators (RNG).
-For our program, that means setting the seed of the {\tt random} module. We may
-also want to save the output of the program to a file, so that we can easily
-verify that consecutive runs do produce the same output: eyeballing differences
-is unreliable and time-consuming---and therefore won't be done systematically.
+The code is running and producing the expected results.
+The next step is to make sure that you can produce the same output over successive runs of your program.
+In other words, the next step is to make your program deterministic, producing {\em repeatable} output.
+Such repeatability is useful.
+If a run of the program produces a particularly puzzling results, it allows you to scrutinize any step of the execution of the program by re-running it again with extraneous prints, or inside a debugger.
+It is also the easiest way to prove that the program did indeed produce the published results.
+Such repeatability is not always possible or easy \citep{Diethelm:2012, Courtes:2015}.
+But for sequential programs not depending on exotic hardware, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\
 
+For our program, that means setting the seed of the {\tt random} module.
+We may also want to save the output of the program to a file, so that we can easily verify that consecutive runs do produce the same output: eyeballing differences is unreliable and  time-consuming---and therefore won't be done systematically.\\
+
+\noindent \begin{minipage}[c]{\linewidth}
 \begin{code}{\textbf{\textsc{Listing 3:}} Random walk (R$^2$)}
@@
 \end{code}
+\end{minipage}\\
 
-Setting seeds is not without danger.  Using 439 as a seed in the previous
-program would result in ten consecutive +1 steps, which---although a perfectly
-valid random walk---lend itself to a gross misinterpretation of the overall
-dynamic of the algorithm.  Verifying that the qualitative aspects of the
-results and the conclusion we make are not tied to a specific initialization of
-the pseudo-random generator is an integral part of any scientific undertaking
-in computational science; this is usually done by repeating the simulations
-multiple times with different seeds.
+Setting seeds is not without danger.
+Using 439 as a seed in the previous program would result in ten consecutive +1 steps, which---although a perfectly valid random walk---lend itself to a gross misinterpretation of the overall dynamic of the algorithm.
+Verifying that the qualitative aspects of the results and the conclusion we make are not tied to a specific initialization of the pseudo-random generator is an integral part of any scientific undertaking in computational science;
+this is usually done by repeating the simulations multiple times with different seeds.
+
 %Determinism is a tool that allow to study bugs and spurious or unexpected behavior. 
@@
 % talk about controversy about setting seeds
+
 % For example, you can play with some parameters and
@@
 % Reproducible definition + R^2 not reproducible
-The R$^2$ code seems fine enough, but it hides several problems that come to
-light when trying to {\em reproduce} results.  A result is said to be
-\emph{reproducible} if another researcher can take the original code and input
-data, execute it, and re-obtain the result \parencite{Peng:2006}. As explained
-by \citeauthor{Donoho:2009} \parencite{Donoho:2009}, scientific practice must
-expect that {\em errors are ubiquitous}, and therefore be robust to
-them. Ensuring reproducibility is an important step toward this: it allows
-other to verify that your code does indeed produce the published results, and
-scrutinize the way that it does.\\
+The R$^2$ code seems fine enough, but it hides several problems that come to light when trying to {\em reproduce} results.
+A result is said to be \emph{reproducible} if another researcher can take the original code and input data, execute it, and re-obtain the result \parencite{Peng:2006}.
+As explained by \citeauthor{Donoho:2009} \parencite{Donoho:2009}, scientific practice must expect that {\em errors are ubiquitous}, and therefore be robust to them.
+Ensuring reproducibility is an important step toward this: it allows other to verify that your code does indeed produce the published results, and scrutinize the way that it does.
+As explained by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility is hard.
+For instance, the R$^2$ program will not produce the same results all the time.
+It will, because it is repeatable, produce the same results over repeated executions.
+But it will not do so over different execution environments.
+The cause is to be found in a change that occurred in the pseudo-random number generator between Python 3.2 and Python 3.3.
+Executed with Python 2.7 to 3.2, the code will produce the sequence 1, 2, 1, 2, 3, 4, 5, 6, 7, 6.
+But with Python 3.3 to 3.6, it will produce 1, 2, 1, 0, 1, 0, 1, 0, -1, 0.\\
 
-As explained by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility
-is hard.  For instance, the R$^2$ program will not produce the same results all
-the time. It will, because it is repeatable, produce the same results over
-repeated executions.  But it will not do so over different execution
-environments. The cause is found in a change that occurred in the pseudo-random
-number generator between Python 3.2 and Python 3.3.  Executed with Python 2.7
-to 3.2, the code will produce the sequence 1, 2, 1, 2, 3, 4, 5, 6, 7, 6. But
-with Python 3.3 to 3.6, it will produce 1, 2, 1, 0, 1, 0, 1, 0, -1, 0.
-% Reproducible definition + R^2 not reproducible
-Because any dependency of your program---to the most basic one, the language
-itself---can change its behavior from one version to another, executability
-(R$^1$) and determinism (R$^2$) are necessary but not sufficient for
-reproducibility. The exact execution environment used to produce the results
-must also be specified---rather than the broadest set of environments where the
-code can be executed. In other words, assertions such as "the results were
-obtained with Python 3.6.1" are more valuable, in a scientific context, than
-"the program should work with Python 3.3.x and above".
-% Reproducibility is hard
-Because of the increasing complexity of our computational stacks, retrieving
-and deciding what is pertinent (CPU architecture? Operating system version?
-Endianness?) might be difficult.  See Appendix \ref{appendix:executionenvs} for
-a discussion.
-% Reproducibility in practice.
-Having precisely described the execution environment, reproducibility is technically obtained. However, \emph{reproducibility in practice} requires more than that.
-% Saving Parameters
-The R$^2$ program uses a random seed but does not keep a trace of it except in
-the code. Therefore, should the code change after the production of the
-results, someone provided with the result file will not be able to know which
-seed was used, and would need to iterate through all possible random seeds,
-which is not possible in practice. R$^3$ saves the seed in the name of the
-result file. More generally, results should always come alongside their
-context, i.e. an exhaustive list of the parameters used and a description of
-the execution environment.
+Because any dependency of your program---to the most basic one, the language itself---can change its behavior from one version to another, executability (R$^1$) and determinism (R$^2$) are necessary but not sufficient for reproducibility.
+The exact execution environment used to produce the results must also be specified---rather than the broadest set of environments where the code can be executed.
+In other words, assertions such as "the results were obtained with Python 3.6.1" are more valuable, in a scientific context, than "the program should work with Python 3.3.x and above".\\
+
+Furthermore, with the increasing complexity of our computational stacks, retrieving and deciding what is pertinent (CPU architecture? Operating system version? Endianness?) might be difficult.
+Having precisely described the execution environment, reproducibility is technically obtained.\\
+
+However, \emph{reproducibility in practice} requires more than that.
+%
+The R$^2$ program uses a random seed but does not keep a trace of it except in the code.
+Therefore, should the code change after the production of the results, someone provided with the result file will not be able to know which seed was used, and would need to iterate through all possible random seeds, which is not possible in practice.
+R$^3$ saves the seed in the name of the result file.
+More generally, results should always come alongside their context, i.e.
+an exhaustive list of the parameters used and a description of the execution environment.
+%
 % Data behind graphs
-Graphs in articles are powerful way to communicate complex information. They
-are imprecise and cumbersome, and sometimes plain unusable, however when it
-comes to checking that the reproduced data match the one published in the
-article. A good figure should always be accompanied by its underlying data to
-allow straightforward comparisons.
+Graphs in articles are powerful way to communicate complex information.
+They are imprecise and cumbersome, and sometimes plain unusable, however when it comes to checking that the reproduced data match the one published in the article.
+A good figure should always be accompanied by its underlying data to allow straightforward comparisons.
+%
 % Unit test for R^3
-Another good practice is to make the code self-verifiable. In R$^3$, a short
-unit test is provided, that allows the code to verify its own
-reproducibility. Should this test fail, then there is little hope of
-reproducing the results. Of course, passing the test is no definitive proof.\\
+Another good practice is to make the code self-verifiable.
+In R$^3$, a short unit test is provided, that allows the code to verify its own reproducibility.
+Should this test fail, then there is little hope of reproducing the results.
+Of course, passing the test is no definitive proof.\\
 
 % Reproducibility implies availability 
-It should be obvious by now but let us state it very explicitely:
-\emph{reproducibility implies availability}. As shown in \citep{Collberg:2016},
-code is often unavailable, or only available upon request. While the latter may
-seem sufficient, changes in email address, changes in career, retirement, a
-busy inbox or poor archiving practices can make a code just as
-unreachable. Code \emph{and input and result files} should be available with
-the published article, as supplementary data, or through a DOI link to a
-scientific repository such as Figshare or Zenodo\footnote{Online code
-  repository such as GitHub \emph{are not} scientific repositories, and may
-  disappear, change name, etc. at any moment. Direct links to them are not
-  perpetual and should be avoided.}.
-%\begin{itemize}
-%\itemsep 0em
-%\item Python 3
-%\item Python 3.6
-%\item {Python 3.6.1 (default, Mar 28 2017, 10:33:50) \\
-%      {\tt [GCC 4.2.1 Compatible Apple LLVM 8.1.0 (clang-802.0.38)] on darwin}}
-%\item {Python 3.6.1 (default, Mar 28 2017, 10:33:50) \\
-%      {\tt [GCC 4.2.1 Compatible Apple LLVM 8.1.0 (clang-802.0.38)] on darwin}\\
-%      {\tt configuration variables: -std=c99 -Wextra -Wno-unused-result -Wno-unused-parameter -Wno-missing-field-initializers ...}}
-%\end{itemize}
+It should be obvious by now that \emph{reproducibility implies availability}.
+As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request.
+While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable.
+Code \emph{and input and result files} should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as Figshare or Zenodo\footnote{Online code repository such as GitHub \emph{are not} scientific repositories, and may disappear, change name, etc.
+at any moment.
+Direct links to them are not perpetual and should be avoided.}.\\
 
-% Adding license for diffusion
-% Lorena Barba example (Science blog)
-
+\noindent \begin{minipage}[c]{\linewidth}
 \begin{code}{\textbf{\textsc{Listing 4:}} Random walk (R$^3$)}
@@
 \end{code}
+\end{minipage}\\
 
 % Summary
-To recap, reproducibility implies re-runnability or repeatability and
-availability, yet imposes additional conditions. Dependencies and platforms
-must be described as precisely and a specifically as possible. Parameters
-values and inputs should be an integral part of the result files. The data
-behind graphs must be published. Unit tests are a good way to embed
-self-diagnostics of reproducibility in the code. Reproducibility is hard, yet
-tremendously necessary.
+To recap, reproducibility implies re-runnability or repeatability and availability, yet  imposes additional conditions.
+Dependencies and platforms must be described as precisely and a specifically as possible.
+Parameters values and inputs should be an integral part of the result files.
+The data behind graphs must be published.
+Unit tests are a good way to embed self-diagnostics of reproducibility in the code.
+Reproducibility is hard, yet tremendously necessary.
 
-% \clearpage
+%\clearpage
 \section*{Re-usable (R$^{\mathbf 4}$)}
 
-Making your program re-usable means it can be easily used, and modified, by you
-and other people, inside and outside your lab. Ensuring your program is
-re-usable is advantageous for a number of reasons. For you, first. Because the
-you now and the you in two years (or two months in some cases) are two
-different persons. Details on how to use the code, its limitations, its quirks,
-may be present to your mind now, but will probably escape you in six months
-\parencite{Donoho:2009}. Here, comments and documentation can make a
-significant difference.  Furthermore, a source code reflects the results of all
-the decisions that were made to create it, but not the reasons behind them. In
-science, where the method and its justification matters as much as the results,
-those reasons are precious knowledge. For instance, it is not sufficient to
-observe that a piece of code is doing correct predictions on a learning task,
-one is interested by the exact algorithm that is employed, and why. In that
-context, a comment on how a given parameter was chosen (optimization,
-experimental data, educated guess, …) is a valuable information.
+Making your program re-usable means it can be easily used, and modified, by you and other people, inside and outside your lab.
+Ensuring your program is re-usable is advantageous for a number of reasons.
 %
-Re-usability of course directly benefits others researchers, from your team and
-outside of it. The easier it is to use your code, the lower is the threshold
-for other to study, modify and extend it. Scientists constantly face the
-constraint of time: if a model is available, documented, and can be installed,
-run and understood all in a few hours, it will be preferred over another that
-would require weeks to reach the same stage.  Moreover, scientists prefer to
-base their work on results they can verify. A reproducible and re-usable code
-offers a platform both verified and easy-to-use, fostering the development of
-derivative works by other researchers.\\
+For you, first.
+Because the you now and the you in two years are two different persons.
+Details on how to use the code, its limitations, its quirks, may be present to your mind now, but will probably escape you in six months \parencite{Donoho:2009}.
+Here, comments and documentation can make a significant difference.
+%
+Furthermore, a source code reflects the results of all the decisions that were made to create it, but not the reasons behind them.
+In science, where the method and its justification matters as much as the results, those reasons are precious knowledge.
+For instance, it is not sufficient to observe that a piece of code is doing correct predictions on a learning task, one is interested by the exact algorithm that is employed, and why.
+In that context, a comment on how a given parameter was chosen (optimization, experimental data, educated guess, …) is a valuable information.
+%
+Re-usability of course directly benefits others researchers, from your team and outside of it.
+The easier it is to use your code, the lower is the threshold for other to study, modify and extend it.
+Scientists constantly face the constraint of time: if a model is available, documented, and can be installed, run and understood all in a few hours, it will be preferred over another that would require weeks to reach the same stage.
+%
+Moreover, scientists prefer to base their work on results they can verify.
+A reproducible and re-usable code offers a platform both verified and easy-to-use, fostering the development of derivative works by other researchers.
+Those derivative works contribute to the impact of your original contribution, and generate citations.
+%
+Having more people examining and using your code also means that potential errors have a higher chance to be caught.
+If people start using your program, they will most likely report bugs or malfunctions they encountered while using your program.
+If you're lucky enough, they might even propose you either bugfixes or improvements, hence improving the overall quality of your software.
+This process contributes to the long-term reproducibility to the extent people continue to use and maintain the program.\\
 
-Those derivative works contribute to the impact of your original contribution,
-and generate citations. Having more people examining and using your code also
-means that potential errors have a higher chance to be caught. If people start
-using your program, they will most likely report bugs or malfunctions they
-encountered while using your program. If you're lucky enough, they might even
-propose you either bugfixes or improvements, hence improving the overall
-quality of your software. This process contributes to the long-term
-reproducibility to the extent people continue to use and maintain the program.
-%
-However, making your program to be re-usable by others requires a little
-work. If it is natural, when one writes a program for oneself, to make implicit
-choices that suit one needs, such implicit choices won't resist the will of a
-user community. For example, having magic numbers or ill-defined constants in
-your program might be cumbersome for others. Similarly, it might be acceptable
-to not have a proper documentation when you're the only user (actually it might
-not be acceptable because the you now and the you in six months are two
-different persons), but other users want to know precisely how to use this or
-that function and what it does exactly.
-%
-If we look back at our random walker, we can list all the implicit choices that
-have been made:
+However, making your program to be re-usable by others requires a little work.
+If it is natural, when one writes a program for oneself, to make implicit choices that suit one needs, such implicit choices won't resist the will of a user community.
+For example, having magic numbers or ill-defined constants in your program might be cumbersome for others.
+Similarly, it might be acceptable to not have a proper documentation when you're the only user (actually it might not be acceptable because the you now and the you in six months are two different persons), but other users want to know precisely how to use this or that function and what it does exactly.
+
+If we look back at our random walker, we can list all the implicit choices that have been made:
 \begin{itemize}
@@
 \item Walker walk along a single dimension
-\item etc.
+\item ...
 \end{itemize}
-However, we won't try to implement all these features at once, but only giving the possibility for a user to implement them if needed.
+Here, we won't try to implement all these features at once, but only giving the possibility for a user to implement them if needed.\\
 
+\noindent \begin{minipage}[c]{\linewidth}
 \begin{code}{\textbf{\textsc{Listing 5:}} Random walk (R$^4$)}
@@
     print("Result:",  x)
-    with open("results-R5-%d.txt" % seed, "w") as file:
-        file.write("Version: R5")
+    with open("results-R4-%d.txt" % seed, "w") as file:
+        file.write("Version: R4")
         file.write("Seed: %d" % args.seed)
@@
 \end{code}
+\end{minipage}\\
 
@@
 
-Having made a software re-usable by others should offer a decent protection against bugs.
-Unfortunately, this is not always the case and some recent cases have demonstrated the tremendous impact a bug can have in Science \citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}.
-This is why, as explained by Peng et al.
-\cite{Peng:2006}, {\em the replication of important findings by multiple independent investigators is fundamental to the accumulation of scientific evidence}.\\
+Having made a software re-usable by others should offer a decent protection
+against bugs. Unfortunately, this is not always the case and some recent cases
+have dramatically illustrated the tremendous impact a bug can have in Science
+\citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}. This is
+why, as explained by Peng et al. \cite{Peng:2006}, {\em the replication of
+  important findings by multiple independent investigators is fundamental to
+  the accumulation of scientific evidence}. Replicability is the implicit
+assumption that any article that does not provide the code source makes: that
+the description it provides of the algorithms is sufficiently precise and
+complete to re-obtain the results it presents. While every published article
+should strive for replicability, even the best efforts may fall short because
+an oversight, a typo or a difference between what is evident between the writer
+and the reader. In that case, a reproducible code is a crucial backup. Hence,
+replicability does not negate the necessity of reproducibility. To illustrate
+this, let us consider what could be the textual description of your random
+walker (as it would be written in an article describing it):
+%
+\begin{quotation}
+{\em The model has been implemented using the standard Python random generator initialized with the seed 0.
+At each iteration, a uniform number between -1 (excluded) and +1 (excluded) was drawn and the sign of the result was used for generating a positive or negative step.}
+\end{quotation}
+%
+If we were to replicate this random walk, for example using Numpy which is a
+numerical Python library, from this description, we would be confronted with some
+unexpected discrepancies between the original Python version and the replicated
+version. The reason is to be found in a subtle difference in the respective
+implementations of the RNG between Python and Numpy. As explained in the
+\href{https://docs.python.org/3.6/library/random.html}{Python documentation},
+{\em Python uses the Mersenne Twister as the core generator. It produces
+53-bit precision floats and has a period of 2**19937-1.} Numpy RNG is
+also based on the Mersenne Twister generator but there are differences in the
+way seed is interpreted when initializing the generator. Fortunately, Both
+Numpy and Python offer access to the internal state and we can {\em fix} it to
+make the behavior of one RNG to match the other RNG. Here, we want to make the
+Numpy RNG to match the Python RNG behavior.But there are also more sublte
+problems with this description. If we look more closely at it, we can realize
+that nothing is said about the specific case of {\tt 0} when generating a step.
+Do whe have to consider {\tt 0} to be a positive or a negative step? Without
+further information and without the original code, it is up to the reader to
+decide. If this would not make a huge difference for small simulations, it
+may nonetheless compromise statistics of the results for long runs.\\
 
-Replicability is the assumption that any article that does not provide the code source makes: that the description it provides of the algorithms is sufficiently precise and complete to re-obtain the results it presents.
-While every published article should strive for replicability, even the best efforts may fall short because an oversight, a typo or a difference between what is evident between the writer and the reader.
-In that case, a reproducible code is a crucial backup.
-Hence, replicability does not negate the necessity of reproducibility.
+Such perfect quantitative match may seem unnecessary in most cases, yet, in
+many replication projects, quantitative comparisons are an objective tests that
+the behavior is reproduced. Statistical tests demand more work, and may fall
+short of detecting small differences in the algorithms, that may results in
+huge difference of behavior in different conditions. In all cases, eyeballing
+difference is graphs is always difficult. Therefore, one should always provide
+the data behind the graphs to allow such comparisons. Without such explicit
+replication effort, it would be hard to discover this faulty behavior and this
+is one of the reason why the \href{https://rescience.github.io}{ReScience
+  journal} has been created. This journal {\em targets computational research and
+encourages the explicit replication of already published research, promoting
+new and open-source implementations in order to ensure that the original
+research is reproducible}.
 
-In our case, the replication of the random walk brings some unexpected discrepancies between the original Python version and the replicated Numpy version.
-The reason is to be found in a differences in the respective implementations of the RNG between Python and Numpy.
-As explained in the \href{https://docs.python.org/3.6/library/random.html}{Python documentation}, {\em Python uses the Mersenne Twister as the core generator. It produces 53-bit precision floats
-and has a period of 2**19937-1.
-The underlying implementation in C is both fast and threadsafe.}
-Numpy RNG is also based on the Mersenne Twister generator but there are differences in the way seed is interpreted when initializing the generator.
-Fortunately, Both Numpy and Python offer access to the internal state and we can {\em fix} it to make the behavior of one RNG to match the other RNG.
-Here, we want to make the Numpy RNG to match the Python RNG behavior.
-
-In our case, a perfect quantitative match may seem unnecessary.
-But in many replication projects, quantitative comparisons are an objective tests that the behavior is reproduced.
-Statistical tests demand more work, and may fall short of detecting small differences in the algorithms, that may results in huge difference of behavior in different conditions.
-In all cases, eyeballing difference is graphs is always difficult.
-Therefore, one should always provide the data behind the graphs to allow such comparisons.
-
-Without such explicit replication effort, it would be hard to discover this faulty behavior and this why the ReScience journal has been created.
-This journal targets computational research and encourages the explicit replication of already published research, promoting new and open-source implementations in order to ensure that the original research is reproducible.
-
+\noindent \begin{minipage}[c]{\linewidth}
 \begin{code}{\textbf{\textsc{Listing 6:}} Random walk (R$^5$)}
@@
 
+    # Note: We consider 0 to be a positive step
     steps = 2*(rng.uniform(-1,+1,n) > 0) - 1
@@
     print(x)
-    with open("results-R4-%d.txt" % seed, "w") as file:
+    with open("results-R5-%d.txt" % seed, "w") as file:
         file.write(str(x))
 \end{code}
+\end{minipage}
 
@@
 The R$^0$ form, while easy to write and simple to understand is not a good scientific code while the full-fledge R$^{4}$ form is much preferred but requires a fair amount of work that might be too much of a burden in some circumstances.
-$R^3$ form seems to be an acceptable compromise and should be accepted as the minimum scientific standard.
+$R^3$ form seems to be an acceptable compromise and should be accepted as the minimum scientific standard \citep{Wilson:2017}.
 This means this should be actually checked by reviewers and publishers when code is part of a work worth to be published.
-But it's hardly the case today.\\
-
+But it's hardly the case today.
+%
 The choice of the Python programming language to illustrate our point was not arbitrary.
@@
 
-\newpage 
-\appendix
-\section{Appendix: Recording Execution Environments}
-\label{appendix:executionenvs}
-%% Talk about provenance-tracking and its problems (endianess, "as well as any defaults you may have set that might have an impact.")
+% \newpage 
+% \appendix
 
-%% Talk about saving VMs, Docker images and its problems.
+% \section{Appendix: Recording Execution Environments}
+% \label{appendix:executionenvs}
+% Talk about provenance-tracking and its problems (endianess, "as well as any defaults you may have set that might have an impact.")
+% Talk about saving VMs, Docker images and its problems.
 

### 2017-07-17 draft.tex
Update Introduction

@@
 re-obtained by an independent party, it merely becomes, at best, an observation
-that may feed future research (\cite{Mesirov:2010,osc:2015}). Replication
-issues have received increased attention in recent years, in the domain of
-medicine and psychology in particular (\cite{Iqbal:2016}). One could think that
-computer science would mostly be shielded from such issues. But precisely
-because it is easy to believe that if a program runs once and gives expected
+that may inspire future research (\cite{Mesirov:2010,osc:2015}). Replication
+issues have received increased attention in recent years, with a particular focus on medicine and psychology (\cite{Iqbal:2016}).
+One could think that
+computer science would mostly be shielded from such issues, since a computer program
+intrinsically describes precisely what it does and is easy to disseminate to other researchers.
+
+But precisely because it is easy to believe that if a program runs once and gives the expected
 results it will do so forever, crucial steps to transform working code into
@@
 
-Let us illustrate this with a very simple example, a random walk written in Python:\\
+Let us illustrate this with a small example, a random walk written in Python:\\
 

### 2017-07-17 draft.tex
Update Re-runnable

@@
 At the same time, modernizing an unmaintained ten-year-old code can reveal itself to be an arduous and expensive undertaking, with each change risking to affect the semantics of the program.
-Rather trying to predict the future or painstakingly dusting off old code, an often easier and less dangerous solution is to recreate the old execution environment.
+Rather trying to predict the future or painstakingly dusting off old code, an often easier and less dangerous solution is to recreate the old execution environment\footnote{To be clear, and although virtual machines are often a great help, this is not always possible.
+It is, however, \emph{always} more difficult when the original execution environment is not documented.}.
 For this to happen however, the dependencies in terms of systems, software and libraries must be made clear enough.\\
 
-A \emph{re-runnable} code is one that describes an execution environment in which it is executable with enough details to be recreated.
+A \emph{re-runnable} code is one that can---surprisingly---be run again, i.e.
+more than the one time that was needed to produce the results.
+It is important to notice that the re-runnability of a code is not an intrinsic property.
+Rather, it depends on the context, and as mentionned above, re-runnability is increasingly difficult as the code ages.
+Therefore, to be re-runnable in other researchers' computers and through time, a re-runnable code should describes an execution environment in which it is executable with enough details to be recreated.
 As shown by \citep{Collberg:2016}, this is far from being either obvious or easy.\\
@@
 This not the case because it uses the print {\em instruction} and the {\tt xrange} operator, both specific to Python 2.
-The print {\em instruction}, available in Python 2 (a version still widely used; support is scheduled to stop in 2020), has been deprecated in Python 3 (first released in 2008, almost a decade ago) in favor or a  print {\em function}.
-The {\tt xrange} operator has been replaced by the {\tt range} operator in Python 3.
+The print {\em instruction}, available in Python 2 (a version still widely used; support is scheduled to stop in 2020), has been deprecated in Python 3 (first released in 2008, almost a decade ago) in favor or a  print {\em function}, while the {\tt xrange} operator has been replaced by the {\tt range} operator in Python 3.
 In order to try to future-proof the code a bit, we might as well target Python 3, as is done in the R$^1$ version.
 Incidentally, it remains compatible with Python 2.
-But whatever version is chosen, the crucial step here is to document it.
+But whichever version is chosen, the crucial step here is to document it.
 

### 2017-07-19 draft.tex
Update Repeatable
Update reproducible
* Update code with provenance function

@@
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\textbf{\textsc{Listing 3:}} Random walk (R$^2$)}
+\begin{code}{\textbf{\textsc{Listing 3:}} Re-runnable, repeatable random walk (R$^2$)}
 # Tested with Python 3
@@
 
-%Determinism is a tool that allow to study bugs and spurious or unexpected behavior. 
-%Determinism is also critical for other to reproduce and study your results. 
-% talk about controversy about setting seeds
-
 % For example, you can play with some parameters and
@@
 The R$^2$ code seems fine enough, but it hides several problems that come to light when trying to {\em reproduce} results.
-A result is said to be \emph{reproducible} if another researcher can take the original code and input data, execute it, and re-obtain the result \parencite{Peng:2006}.
+A result is said to be \emph{reproducible} if another researcher can take the original code and input data, execute it, and re-obtain the same result \parencite{Peng:2006}.
 As explained by \citeauthor{Donoho:2009} \parencite{Donoho:2009}, scientific practice must expect that {\em errors are ubiquitous}, and therefore be robust to them.
-Ensuring reproducibility is an important step toward this: it allows other to verify that your code does indeed produce the published results, and scrutinize the way that it does.
-As explained by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility is hard.
+Ensuring reproducibility is an important step toward this: it allows other to verify that your code does indeed produce the published results, and to scrutinize the procedure it uses to do it.
+As demonstrated by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility is hard.\\
+
 For instance, the R$^2$ program will not produce the same results all the time.
 It will, because it is repeatable, produce the same results over repeated executions.
-But it will not do so over different execution environments.
+But it will not necessarily do so over different execution environments.
 The cause is to be found in a change that occurred in the pseudo-random number generator between Python 3.2 and Python 3.3.
@@
 Because any dependency of your program---to the most basic one, the language itself---can change its behavior from one version to another, executability (R$^1$) and determinism (R$^2$) are necessary but not sufficient for reproducibility.
-The exact execution environment used to produce the results must also be specified---rather than the broadest set of environments where the code can be executed.
-In other words, assertions such as "the results were obtained with Python 3.6.1" are more valuable, in a scientific context, than "the program should work with Python 3.3.x and above".\\
+The exact execution environment used to produce the results must also be specified---rather than the broadest set of environments where the code can be effectively run.
+In other words, assertions such as "the results were obtained with Python 3.6.1" are more valuable, in a scientific context, than "the program should work with Python 3.x and above".\\
 
-Furthermore, with the increasing complexity of our computational stacks, retrieving and deciding what is pertinent (CPU architecture? Operating system version? Endianness?) might be difficult.
-Having precisely described the execution environment, reproducibility is technically obtained.\\
+With the increasing complexity of our computational stacks, retrieving and deciding what is pertinent (CPU architecture?
+Operating system version?
+Endianness?) can be difficult, however.
+A best effort must be made.\\
 
-However, \emph{reproducibility in practice} requires more than that.
-%
+Having precisely described the execution environment, reproducibility is \emph{technically} obtained. However, \emph{reproducibility in practice} requires more than that.\\
+
 The R$^2$ program uses a random seed but does not keep a trace of it except in the code.
-Therefore, should the code change after the production of the results, someone provided with the result file will not be able to know which seed was used, and would need to iterate through all possible random seeds, which is not possible in practice.
-R$^3$ saves the seed in the name of the result file.
-More generally, results should always come alongside their context, i.e.
-an exhaustive list of the parameters used and a description of the execution environment.
-%
+Therefore, should the code change after the production of the results, someone provided with the result file will not be able to know which seed was used, and would need to iterate through all possible random seeds, an impossible task in practice.
+More generally, results files should always come alongside their context, i.e.
+an exhaustive list of the parameters used and a description of the execution environment, as the R$^3$ code does.\\
+
+%FIXME: add complete scripts to graphs.
+
 % Data behind graphs
 Graphs in articles are powerful way to communicate complex information.
-They are imprecise and cumbersome, and sometimes plain unusable, however when it comes to checking that the reproduced data match the one published in the article.
-A good figure should always be accompanied by its underlying data to allow straightforward comparisons.
+However, they are imprecise and cumbersome, and sometimes plain unusable, when it comes to checking that the reproduced data match the one published in the article.
+A good figure should always be accompanied by its underlying data to allow straightforward numeric comparisons.
 %
@@
 at any moment.
-Direct links to them are not perpetual and should be avoided.}.\\
+Direct links to them are not perpetual, and, when used, should always be completed by a link to a scientific archive.}.\\
 
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\textbf{\textsc{Listing 4:}} Random walk (R$^3$)}
+\begin{code}{\textbf{\textsc{Listing 4:}} Re-runnable, repeatable, reproducible random walk (R$^3$)}
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
 # Release under the BSD 2-clause license
-# Tested with CPython 3.6 / macOS 10.12.4 / 64 bits architecture
+# Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
+import platform, datetime
 import random
-from itertools import accumulate
 
-def walk(n):
-    """ Random walk for n steps """
+def provenance(): # Get info about the execution environment
+    return {'python'   : {'implementation': platform.python_implementation(),
+                          'version'       : platform.python_version_tuple(),
+                          'compiler'      : platform.python_compiler(),
+                          'branch'        : platform.python_branch(),
+                          'revision'      : platform.python_revision()},
+            'platform' : platform.platform(),
+            'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC
 
-    steps = [-1 if random.uniform(-1,+1) < 0 else +1 for i in range(n)]
-    return list(accumulate(steps))
+def walk():
+    promenade, total = [], 0
+    for i in range(10):
+        step = +1 if random.uniform(-1,+1) > 0 else -1
+        total += step
+        promenade.append(total)
+    return promenade
 
-if __name__ == '__main__':
 # Unit test
 random.seed(1)
-    assert walk(10) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+assert walk() == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
 
@@
 random.seed(seed)
-    x = walk(10)
+w = walk()
 
 # Display & save results
-    print(x)
-    with open("results-R3-%d.txt" % seed, "w") as file:
-        file.write(str(x))
+print(w)
+results = {'data': w, 'seed': seed, 'provenance': provenance()}
+with open("results-R3.txt", "w") as fd:
+    fd.write(str(results))
 \end{code}
@@
 % Summary
-To recap, reproducibility implies re-runnability or repeatability and availability, yet  imposes additional conditions.
+To recap, reproducibility implies re-runnability and repeatability and availability, yet imposes additional conditions.
 Dependencies and platforms must be described as precisely and a specifically as possible.
 Parameters values and inputs should be an integral part of the result files.
-The data behind graphs must be published.
+The data behind the graphs must be published.
 Unit tests are a good way to embed self-diagnostics of reproducibility in the code.

### 2017-07-19 draft.tex
Update Reusable
* renamed all instance of `re-use` into `reuse`

@@
 
-\title{Re-run, Repeat, Reproduce, Re-use, Replicate:\\Transforming Code into Scientific Contributions}
+\title{Re-run, Repeat, Reproduce, Reuse, Replicate:\\Transforming Code into Scientific Contributions}
 % \title{Run Python, run!}
@@
 sure it is re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$),
-re-usable (R$^4$) and replicable (R$^5$).\\
+reusable (R$^4$) and replicable (R$^5$).\\
 
@@
 
-Have you ever try to re-run a program you wrote some years ago?
+Have you ever tried to re-run a program you wrote some years ago?
 Frustratingly, it can often be surprisingly hard.
@@
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\textbf{\textsc{Listing 2:}} Random walk (R$^1$)}
+\begin{code}{\textbf{\textsc{Listing 2:}} Re-runnable random walk (R$^1$)}
 # Tested with Python 3
@@
 %\clearpage
-\section*{Re-usable (R$^{\mathbf 4}$)}
+\section*{Reusable (R$^{\mathbf 4}$)}
 
-Making your program re-usable means it can be easily used, and modified, by you and other people, inside and outside your lab.
-Ensuring your program is re-usable is advantageous for a number of reasons.
-%
+Making your program reusable means it can be easily used, and modified, by you and other people, inside and outside your lab.
+Ensuring your program is reusable is advantageous for a number of reasons.\\
+
 For you, first.
@@
 Here, comments and documentation can make a significant difference.
-%
-Furthermore, a source code reflects the results of all the decisions that were made to create it, but not the reasons behind them.
+In fact, a source code reflects the results of all the decisions that were made to create it, but not the reasons behind them.
 In science, where the method and its justification matters as much as the results, those reasons are precious knowledge.
-For instance, it is not sufficient to observe that a piece of code is doing correct predictions on a learning task, one is interested by the exact algorithm that is employed, and why.
-In that context, a comment on how a given parameter was chosen (optimization, experimental data, educated guess, …) is a valuable information.
-%
-Re-usability of course directly benefits others researchers, from your team and outside of it.
+In that context, a comment on how a given parameter was chosen (optimization, experimental data, educated guess), why a library was chosen over another (conceptual or technical?) is valuable information.\\
+
+Reusability of course directly benefits others researchers from your team and outside of it.
 The easier it is to use your code, the lower is the threshold for other to study, modify and extend it.
-Scientists constantly face the constraint of time: if a model is available, documented, and can be installed, run and understood all in a few hours, it will be preferred over another that would require weeks to reach the same stage.
-%
+Scientists constantly face the constraint of time: if a model is available, documented, and can be installed, run and understood all in a few hours, it will be preferred over another that would require weeks to reach the same stage.\\
+
 Moreover, scientists prefer to base their work on results they can verify.
-A reproducible and re-usable code offers a platform both verified and easy-to-use, fostering the development of derivative works by other researchers.
-Those derivative works contribute to the impact of your original contribution, and generate citations.
-%
+A reproducible and reusable code offers a platform both validated and easy-to-use, fostering the development of derivative works by other researchers.
+Those derivative works contribute to the impact of your original contribution.\\
+
 Having more people examining and using your code also means that potential errors have a higher chance to be caught.
@@
 
-However, making your program to be re-usable by others requires a little work.
+Despites all this, reusability is often overlooked, and it is not hard to see why.
+Scientists are rarely trained in software engineering, and reusability can represent an expensive endeavour if undertaken as an afterthought, for little tangible short-term benefits, for a codebase that might, after all, see only a single use.
+And, in fact, reusability is not as indispensable a requirement as the others one presented here.\\
+
+Yet, some simple measures can tremendously increase reusability, and at the same time strenghten reproducibility or re-runnability.\\
+
+Avoid hardcoded or magic numbers.
+Magic numbers are numbers present directly in the source code, that do not have a name and therefore can be difficult to interpret semantically.
+Hardcoded values are variables that cannot be changed through a function argument or a parameter configuration file.
+To be modified, they involve editing the code, which is cumbersome and error-prone.
+In the R$^3$ code, the seed and the number of steps are respectively hardcoded and magic.\\
+
+%FIXME add ref
+Similarily, code behavior should not have to be changed by commenting/uncommenting code.
+Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitely set through parameters accessible to the end-user.
+This improves reproducibility in two-ways: it allows those conditions to be explicitely tracked in the result files since they contains the parameters used.
+It also allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
+With a documention explaining which script or configuration file correspond to which experiment, reproducing the different figures becomes straightforward.\\
+
+% def walk():
+%     promenade, total = [], 0
+%     # for i in range(20): # Figure 2
+%     for i in range(10): # Figure 1
+%         step = +1 if random.uniform(-1,+1) > 0 else -1
+%         total += step
+%         promenade.append(total)
+%     return promenade
+
+Documentation, in fact, is one of the most potent tools for reusability.
+a proper documentation on how to install and run the software often makes the difference for other reasearchers whether they manage to use the software or not.\\
+
+A comment describing what each function does, however evident, can also avoid hours of head-scratching.
+Great code may need few comments.
+Scientists, however, rarely qualify as senior developers.
+Code quality, in science, is often mediocre.
+Bad, complicated code should be rewritten until is simple enough to explain itself.
+But realistically, this is hardly ever going to be done: there is simply not enough incentive for it.
+There, a comment that explains the intention and scientific reasons behind a block of code can tremendously useful.\\
+
+% However, making your program reusable requires a little work.
 If it is natural, when one writes a program for oneself, to make implicit choices that suit one needs, such implicit choices won't resist the will of a user community.
@@
 Similarly, it might be acceptable to not have a proper documentation when you're the only user (actually it might not be acceptable because the you now and the you in six months are two different persons), but other users want to know precisely how to use this or that function and what it does exactly.
-
-If we look back at our random walker, we can list all the implicit choices that have been made:
-\begin{itemize}
-\item Step number has been hard-coded (magic number 10)
-\item Step size has been fixed to 1
-\item Initial position has been set to 0
-\item No possibility of using a user-specified random seed
-\item Only one run can be started at once
-\item Walker walk along a single dimension
-\item ...
-\end{itemize}
-Here, we won't try to implement all these features at once, but only giving the possibility for a user to implement them if needed.\\
+%FIXME: reword
+%
+% If we look back at our random walker, we can list all the implicit choices that have been made:
+% \begin{itemize}
+% \item Step number has been hard-coded (magic number 10)
+% \item Step size has been fixed to 1
+% \item Initial position has been set to 0
+% \item No possibility of using a user-specified random seed
+% \item Only one run can be started at once
+% \item Walker walk along a single dimension
+% \item ...
+% \end{itemize}
+% Here, we won't try to implement all these features at once, but only giving the possibility for a user to implement them if needed.\\
 
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\textbf{\textsc{Listing 5:}} Random walk (R$^4$)}
+\begin{code}{\textbf{\textsc{Listing 5:}} Re-runnable, repeatable, reproducible, reusable random walk (R$^4$)}
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
@@
 
-Having made a software re-usable by others should offer a decent protection
+Having made a software reusable by others should offer a decent protection
 against bugs. Unfortunately, this is not always the case and some recent cases
@@
 should strive for replicability, even the best efforts may fall short because
-an oversight, a typo or a difference between what is evident between the writer
-and the reader. In that case, a reproducible code is a crucial backup. Hence,
+an oversight, a typo or a difference between what is evident for the writer
+and for the reader. In that case, a reproducible code is a crucial backup. Hence,
 replicability does not negate the necessity of reproducibility. To illustrate
@@
 \begin{quotation}
-{\em The model has been implemented using the standard Python random generator initialized with the seed 0.
+{\em The model has been implemented using the Mersene Twister generator initialized with the seed 0.
 At each iteration, a uniform number between -1 (excluded) and +1 (excluded) was drawn and the sign of the result was used for generating a positive or negative step.}
@@
 make the behavior of one RNG to match the other RNG. Here, we want to make the
-Numpy RNG to match the Python RNG behavior.But there are also more sublte
+Numpy RNG to match the Python RNG behavior. But there are also more sublte
 problems with this description. If we look more closely at it, we can realize
@@
 short of detecting small differences in the algorithms, that may results in
-huge difference of behavior in different conditions. In all cases, eyeballing
-difference is graphs is always difficult. Therefore, one should always provide
-the data behind the graphs to allow such comparisons. Without such explicit
+huge difference of behavior in different conditions. Without such explicit
 replication effort, it would be hard to discover this faulty behavior and this
@@
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\textbf{\textsc{Listing 6:}} Random walk (R$^5$)}
+\begin{code}{\textbf{\textsc{Listing 6:}} Replicated random walk (R$^5$)}
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau

### 2017-07-22 codes/r3.py
Add r3, r4 codes

@@ -1,0 +1,37 @@
+# Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
+# Release under the BSD 2-clause license
+# Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
+import platform, datetime
+import random
+
+def provenance(): # Get info about the execution environment
+    return {'python'   : {'implementation': platform.python_implementation(),
+                          'version'       : platform.python_version_tuple(),
+                          'compiler'      : platform.python_compiler(),
+                          'branch'        : platform.python_branch(),
+                          'revision'      : platform.python_revision()},
+            'platform' : platform.platform(),
+            'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC
+
+def walk():
+    promenade, total = [], 0
+    for i in range(10):
+        step = +1 if random.uniform(-1,+1) > 0 else -1
+        total += step
+        promenade.append(total)
+    return promenade
+
+# Unit test
+random.seed(1)
+assert walk() == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+
+# Random walk for 10 steps
+seed = 1
+random.seed(seed)
+w = walk()
+
+# Display & save results
+print(w)
+results = {'data': w, 'seed': seed, 'provenance': provenance()}
+with open("results-R3.txt", "w") as fd:
+    fd.write(str(results))

### 2017-07-22 codes/r4.py

@@ -1,0 +1,42 @@
+# Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
+# Release under the BSD 2-clause license
+# Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
+import platform, datetime, json
+import random
+
+def provenance():
+    """Retrieve data about the execution environment."""
+    return {'python'   : {'implementation': platform.python_implementation(),
+                          'version'       : platform.python_version_tuple(),
+                          'compiler'      : platform.python_compiler(),
+                          'branch'        : platform.python_branch(),
+                          'revision'      : platform.python_revision()},
+            'platform' : platform.platform(),
+            'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC
+
+def walk(n, seed):
+    """Return a `n` steps random walk.
+
+    `seed` is used to initialized the Mersene Twister random number generator."""
+    random.seed(seed)
+    rndwalk, total = [], 0
+    for i in range(10):
+        step = +1 if random.uniform(-1,+1) > 0 else -1
+        total += step
+        rndwalk.append(total)
+    return rndwalk
+
+# Unit test
+assert walk(10, 1) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+
+def generate_results(n, seed, filename='results-R4'):
+    """Generate results and save them to disk as a JSON file"""
+    w = walk(n, seed)
+
+    results = {'data': w, 'n': n, 'seed': seed, 'provenance': provenance()}
+    with open(filename + '.json', 'w') as fd:
+        json.dump(results, fd)
+    return results
+
+if __name__ == '__main__':
+    print(generate_results(10, 1)['data'])

### 2017-07-24 draft.tex
Update article
* Replicable and conclusion have been extensively ammended.
* Other sections have been lightly ammended.
* Swapped authors.

@@
 \author{
-  \textbf{Nicolas P. Rougier}$^{1,2,3,*}$, \textbf{Fabien C. Y. Benureau}$^{1,2,3}$\\
-  \begin{minipage}{\textwidth}
+  \textbf{Fabien C. Y. Benureau}$^{1,2,3,*}$,
+  \textbf{Nicolas P. Rougier}$^{1,2,3}$\\ \begin{minipage}{\textwidth}
     \begin{center}
@@
       $^{*}$Corresponding author:
-      \href{mailto:Nicolas.Rougier@inria.fr}{Nicolas.Rougier@inria.fr}
+      \href{mailto:fabien.benureau@gmail.com}{fabien.benureau@gmail.com}
     \end{center}
@@
 that may inspire future research (\cite{Mesirov:2010,osc:2015}). Replication
-issues have received increased attention in recent years, with a particular focus on medicine and psychology (\cite{Iqbal:2016}).
+issues have received increased attention in recent years, with a particular mediatic focus on medicine and psychology (\cite{Iqbal:2016}).
 One could think that
 computer science would mostly be shielded from such issues, since a computer program
-intrinsically describes precisely what it does and is easy to disseminate to other researchers.
+intrinsically describes precisely what it does and is easy to transmit perfectly to other researchers.
 
@@
 \citep{Sandve:2013,Schwab:2000}. Computer science is plagued by replication
-problems, in part, precisely because it seems impervious to them. In fact, a
+problems, in part, precisely because it seems impervious to them. A
 program can fail as a scientific contribution in many different ways for many
 different reasons. Borrowing the terms coined by \citeauthor{Goble:2016}
-\citep{Goble:2016}, for your program to contribute to Science, you have to make
+\citep{Goble:2016}, for your program to contribute to science, you have to make
 sure it is re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$),
@@
 
-Executed, this program would display
+Executed, this program would display:
 \begin{code}{Output}
@@
 To be clear, it is impossible to write future-proof code, and the best efforts can be stymied by the smallest change in one of the dependencies.
-At the same time, modernizing an unmaintained ten-year-old code can reveal itself to be an arduous and expensive undertaking, with each change risking to affect the semantics of the program.
-Rather trying to predict the future or painstakingly dusting off old code, an often easier and less dangerous solution is to recreate the old execution environment\footnote{To be clear, and although virtual machines are often a great help, this is not always possible.
+At the same time, modernizing an unmaintained ten-year-old code can reveal itself to be an arduous and expensive undertaking, and dangerous as that, since each change risks to affect the semantics of the program.
+Rather trying to predict the future or painstakingly dusting off old code, an often more straightforward and less dangerous solution is to recreate the old execution environment\footnote{To be clear, and although virtual machines are often a great help here, this is not always possible.
 It is, however, \emph{always} more difficult when the original execution environment is not documented.}.
@@
 
-A \emph{re-runnable} code is one that can---surprisingly---be run again, i.e.
+A \emph{re-runnable} code is one that can be run again when needed, i.e.
 more than the one time that was needed to produce the results.
 It is important to notice that the re-runnability of a code is not an intrinsic property.
-Rather, it depends on the context, and as mentionned above, re-runnability is increasingly difficult as the code ages.
-Therefore, to be re-runnable in other researchers' computers and through time, a re-runnable code should describes an execution environment in which it is executable with enough details to be recreated.
+Rather, it depends on the context, and re-runnability becomes increasingly difficult as the code ages.
+Therefore, to be re-runnable in other researchers' computers and through time, a re-runnable code should describes---with enough details to be recreated---an execution environment in which it is executable.
 As shown by \citep{Collberg:2016}, this is far from being either obvious or easy.\\
@@
 Such repeatability is useful.
-If a run of the program produces a particularly puzzling results, it allows you to scrutinize any step of the execution of the program by re-running it again with extraneous prints, or inside a debugger.
+If a run of the program produces a particularly puzzling result, it allows you to scrutinize any step of the execution of the program by re-running it again with extraneous prints, or inside a debugger.
 It is also the easiest way to prove that the program did indeed produce the published results.
 Such repeatability is not always possible or easy \citep{Diethelm:2012, Courtes:2015}.
-But for sequential programs not depending on exotic hardware, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\
+But for sequential programs not depending on analog inputs, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\
 
 For our program, that means setting the seed of the {\tt random} module.
-We may also want to save the output of the program to a file, so that we can easily verify that consecutive runs do produce the same output: eyeballing differences is unreliable and  time-consuming---and therefore won't be done systematically.\\
+We may also want to save the output of the program to a file, so that we can easily verify that consecutive runs do produce the same output: eyeballing differences is unreliable and  time-consuming, and therefore won't be done systematically.\\
 
@@
 print(walk)
-with open("results.txt", "w") as file: # Saving output to disk
+with open("results-R2.txt", "w") as file: # Saving output to disk
     file.write(str(walk))
@@
 
-Setting seeds is not without danger.
-Using 439 as a seed in the previous program would result in ten consecutive +1 steps, which---although a perfectly valid random walk---lend itself to a gross misinterpretation of the overall dynamic of the algorithm.
+Setting seeds is should be done carefully.
+Using 439 as a seed in the previous program would result in ten consecutive +1 steps\footnote{With CPython 3.3-3.6.
+See the next section for details.}, which---although a perfectly valid random walk---lend itself to a gross misinterpretation of the overall dynamic of the algorithm.
 Verifying that the qualitative aspects of the results and the conclusion we make are not tied to a specific initialization of the pseudo-random generator is an integral part of any scientific undertaking in computational science;
 this is usually done by repeating the simulations multiple times with different seeds.
+%FIXME: is `simulations` pertinent here?
 
@@
 As explained by \citeauthor{Donoho:2009} \parencite{Donoho:2009}, scientific practice must expect that {\em errors are ubiquitous}, and therefore be robust to them.
-Ensuring reproducibility is an important step toward this: it allows other to verify that your code does indeed produce the published results, and to scrutinize the procedure it uses to do it.
+Ensuring reproducibility is an important step toward this: it allows other to verify that your code does indeed produce the published results, and to scrutinize the procedure it uses to produce them.
 As demonstrated by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility is hard.\\
@@
 The exact execution environment used to produce the results must also be specified---rather than the broadest set of environments where the code can be effectively run.
-In other words, assertions such as "the results were obtained with Python 3.6.1" are more valuable, in a scientific context, than "the program should work with Python 3.x and above".\\
+In other words, assertions such as "the results were obtained with CPython 3.6.1" are more valuable, in a scientific context, than "the program should work with Python 3.x and above".\\
 
+%FIXME: this is not a good-enough paragraph
 With the increasing complexity of our computational stacks, retrieving and deciding what is pertinent (CPU architecture?
@@
 
-%FIXME: add complete scripts to graphs.
+% scripts to generates figures
+The code should also contains all the scripts that were used to produce the figures.
+While it may be considered extraneous, code for plotting figures may contains non-trivial statistics code, and is actually just as vulnerable to bugs.
+With a code that is re-runnable, repeatable and contains such scripts, a user can quickly check that published figures can be reproduced.\\
 
-% Data behind graphs
-Graphs in articles are powerful way to communicate complex information.
-However, they are imprecise and cumbersome, and sometimes plain unusable, when it comes to checking that the reproduced data match the one published in the article.
-A good figure should always be accompanied by its underlying data to allow straightforward numeric comparisons.
-%
-% Unit test for R^3
+% data behind graphs
+However, figures by themselves are imprecise and cumbersome, and sometimes plain unusable, when it comes to checking that the reproduced data match the one published in the article.
+A published figure should always be accompanied by its underlying data to allow straightforward numeric comparisons.\\
+
+% unit test for R^3
 Another good practice is to make the code self-verifiable.
@@
 
-% Reproducibility implies availability
+% reproducibility implies availability
 It should be obvious by now that \emph{reproducibility implies availability}.
@@
 While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable.
-Code \emph{and input and result files} should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as Figshare or Zenodo\footnote{Online code repository such as GitHub \emph{are not} scientific repositories, and may disappear, change name, etc.
-at any moment.
-Direct links to them are not perpetual, and, when used, should always be completed by a link to a scientific archive.}.\\
+Code \emph{and} input \emph{and} result files \emph{and} figure data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as Figshare or Zenodo\footnote{Online code repository such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment.
+Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.}.\\
 
@@
 Parameters values and inputs should be an integral part of the result files.
-The data behind the graphs must be published.
+The data and scripts behind the graphs must be published.
 Unit tests are a good way to embed self-diagnostics of reproducibility in the code.
@@
 Here, comments and documentation can make a significant difference.
-In fact, a source code reflects the results of all the decisions that were made to create it, but not the reasons behind them.
-In science, where the method and its justification matters as much as the results, those reasons are precious knowledge.
-In that context, a comment on how a given parameter was chosen (optimization, experimental data, educated guess), why a library was chosen over another (conceptual or technical?) is valuable information.\\
+In fact, a source code reflects the results of all the decisions that were made to create it, but not the reasons behind the decisions.
+In science, where the method and its justification matter as much as the results, those reasons are precious knowledge.
+In that context, a comment on how a given parameter was chosen (optimization, experimental data, educated guess), why a library was chosen over another (conceptual or technical reasons?) is valuable information.\\
 
 Reusability of course directly benefits others researchers from your team and outside of it.
-The easier it is to use your code, the lower is the threshold for other to study, modify and extend it.
-Scientists constantly face the constraint of time: if a model is available, documented, and can be installed, run and understood all in a few hours, it will be preferred over another that would require weeks to reach the same stage.\\
-
+The easier it is to use your code, the lower the threshold is for other to study, modify and extend it.
+Scientists constantly face the constraint of time: if a model is available, documented, and can be installed, run and understood all in a few hours, it will be preferred over another that would require weeks to reach the same stage.
 Moreover, scientists prefer to base their work on results they can verify.
@@
 Having more people examining and using your code also means that potential errors have a higher chance to be caught.
-If people start using your program, they will most likely report bugs or malfunctions they encountered while using your program.
-If you're lucky enough, they might even propose you either bugfixes or improvements, hence improving the overall quality of your software.
+If people start using your program, they will most likely report bugs or malfunctions they encounter.
+If you're lucky enough, they might even propose either bugfixes or improvements, hence improving the overall quality of your software.
 This process contributes to the long-term reproducibility to the extent people continue to use and maintain the program.\\
@@
 
-Yet, some simple measures can tremendously increase reusability, and at the same time strenghten reproducibility or re-runnability.\\
+Yet, some simple measures can tremendously increase reusability, and at the same time strenghten reproducibility and re-runnability.\\
 
@@
 
-%FIXME add ref
-Similarily, code behavior should not have to be changed by commenting/uncommenting code.
+Similarily, code behavior should be changed by commenting/uncommenting code \citep{Wilson:2017}.
 Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitely set through parameters accessible to the end-user.
-This improves reproducibility in two-ways: it allows those conditions to be explicitely tracked in the result files since they contains the parameters used.
+This improves reproducibility in two ways: it allows those conditions to be explicitely tracked in the result files since they contain the parameters used.
 It also allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
 With a documention explaining which script or configuration file correspond to which experiment, reproducing the different figures becomes straightforward.\\
-
-% def walk():
-%     promenade, total = [], 0
-%     # for i in range(20): # Figure 2
-%     for i in range(10): # Figure 1
-%         step = +1 if random.uniform(-1,+1) > 0 else -1
-%         total += step
-%         promenade.append(total)
-%     return promenade
+%FIXME repeating what was said in reproducible.
 
 Documentation, in fact, is one of the most potent tools for reusability.
-a proper documentation on how to install and run the software often makes the difference for other reasearchers whether they manage to use the software or not.\\
+a proper documentation on how to install and run the software often makes the difference for other reasearchers whether they manage to use the code or not.\\
 
-A comment describing what each function does, however evident, can also avoid hours of head-scratching.
+Furthermore, A comment describing what each function does, however evident, can avoid hours of head-scratching.
 Great code may need few comments.
-Scientists, however, rarely qualify as senior developers.
-Code quality, in science, is often mediocre.
-Bad, complicated code should be rewritten until is simple enough to explain itself.
+Scientists, however, are not always brilliant developers.
+Code quality, in science, averages on mediocre.
+Of course, bad, complicated code should be rewritten until is simple enough to explain itself.
 But realistically, this is hardly ever going to be done: there is simply not enough incentive for it.
@@
 \begin{code}{\textbf{\textsc{Listing 5:}} Re-runnable, repeatable, reproducible, reusable random walk (R$^4$)}
-# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
+  # Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
   # Release under the BSD 2-clause license
-# Tested with Python 3.6 / Numpy 1.12.0 / macOS 10.12.4 / 64 bits architecture
+  # Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
+  import platform, datetime, json
   import random
-import numpy as np
 
-def walk(rng, n):
-    """ Random walk for n steps """
+  def provenance():
+      """Retrieve data about the execution environment."""
+      return {'python'   : {'implementation': platform.python_implementation(),
+                            'version'       : platform.python_version_tuple(),
+                            'compiler'      : platform.python_compiler(),
+                            'branch'        : platform.python_branch(),
+                            'revision'      : platform.python_revision()},
+              'platform' : platform.platform(),
+              'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC
 
-    steps = 2*(rng.uniform(-1,+1,n) > 0) - 1
-    return steps.cumsum().tolist()
+  def walk(n, seed):
+      """Return a `n` steps random walk.
 
-def rng(seed):
-    """ Return a random number generator initialized with seed """
+      `seed` is used to initialized the Mersene Twister random number generator."""
+      random.seed(seed)
+      rndwalk, total = [], 0
+      for i in range(10):
+          step = +1 if random.uniform(-1,+1) > 0 else -1
+          total += step
+          rndwalk.append(total)
+      return rndwalk
 
-    rng = np.random.RandomState()
-    rng.seed(seed)
-    return rng
+  # Unit test
+  assert walk(10, 1) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
 
-def test():
-    """ Unit tests """
+  def generate_results(n, seed, filename='results-R4'):
+      """Generate results and save them to disk as a JSON file"""
+      w = walk(n, seed)
 
-    return walk(rng(seed=1), 10) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+      results = {'data': w, 'n': n, 'seed': seed, 'provenance': provenance()}
+      with open(filename + '.json', 'w') as fd:
+          json.dump(results, fd)
+      return results
 
   if __name__ == '__main__':
-    import argparse
-    parser = argparse.ArgumentParser("Random walk")
-    parser.add_argument('--seed', type=int, default=1,
-                        help='Seed for random number generator ')
-    parser.add_argument('n', type=int, default=10,
-                        help='Number of step(s) to walk')
-    args = parser.parse_args()
-
-    # Random walk for n steps
-    x = walk(rng(args.seed), args.n)
-
-    # Display & save results
-    print("Seed:", args.seed)
-    print("Number of steps:", args.n)
-    print("Result:",  x)
-    with open("results-R4-%d.txt" % seed, "w") as file:
-        file.write("Version: R4")
-        file.write("Seed: %d" % args.seed)
-        file.write("Steps number: %d" % args.n)
-        file.write("Output: %s" % str(x))
+      print(generate_results(10, 1)['data'])
 \end{code}
 \end{minipage}\\
+
+Re-usability is not an absolute requirement for scientific code. But it has many benefit, and a few simple measure can foster it considerably.
 
@@
 
-Having made a software reusable by others should offer a decent protection
-against bugs. Unfortunately, this is not always the case and some recent cases
-have dramatically illustrated the tremendous impact a bug can have in Science
+% reusable is not foolproof
+Having made a software reusable offers an additional way to find bugs, especially if your scientific contribution is popular. Unfortunately, this is not always effective, and some recent cases
+have dramatically illustrated the tremendous impact a bug can have in science
 \citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}. This is
@@
   important findings by multiple independent investigators is fundamental to
-  the accumulation of scientific evidence}. Replicability is the implicit
+  the accumulation of scientific evidence}.\\
+
+% reproducible, a definition
+\emph{Replicability} is the implicit
 assumption that any article that does not provide the code source makes: that
@@
 complete to re-obtain the results it presents. While every published article
-should strive for replicability, even the best efforts may fall short because
+should strive for replicability, it is seldom obtained. In fact, absent an explicit effort to make an algorithmic description replicable, there is little probability that it will be.\\
+
+% conceptual description != replicable description
+This is because most papers strive to communicate the main ideas behind their contribution is terms as simple and as clear as possible, so that the reader may be able to easily understand them and the results that are presented to him.
+Trying to ensure replicability in the main text adds a myriad of esoteric details that are not conceptually significant, and clutter the explanations.
+Therefore, unless the writer dedicates an addendum or a section of the supplementary information for technical details specifically aimed at replicability, the information will not be there because there are incentives not to do so.\\
+
+% errors may always remain: reproducibility does not make reproducibility useless.
+But even when those details are present, the best efforts may fall short because
 an oversight, a typo or a difference between what is evident for the writer
-and for the reader. In that case, a reproducible code is a crucial backup. Hence,
+and for the reader \citep{Mesnard:2016}. In that case, a reproducible code is a crucial backup. Hence,
 replicability does not negate the necessity of reproducibility. To illustrate
-this, let us consider what could be the textual description of your random
+this, let us consider what could be the textual description of the random
 walker (as it would be written in an article describing it):
@@
 \begin{quotation}
-{\em The model has been implemented using the Mersene Twister generator initialized with the seed 0.
-At each iteration, a uniform number between -1 (excluded) and +1 (excluded) was drawn and the sign of the result was used for generating a positive or negative step.}
+{\em The model uses the Mersene Twister generator initialized with the seed 0.
+At each iteration, a uniform number between -1 (included) and +1 (excluded) is drawn and the sign of the result is used for generating a positive or negative step.}
 \end{quotation}
 %
-If we were to replicate this random walk, for example using Numpy which is a
-numerical Python library, from this description, we would be confronted with some
-unexpected discrepancies between the original Python version and the replicated
-version. The reason is to be found in a subtle difference in the respective
-implementations of the RNG between Python and Numpy. As explained in the
+This description, while somewhat precise, forgoes---as it is common---the initialization of the variables (here the starting value of the walk ({\tt 0})), and the technical details about which implementation of the RNG is used.\\
+
+% NumPy and Python: differences in the Mersene Twister
+It may look innocuous. After all, the
 \href{https://docs.python.org/3.6/library/random.html}{Python documentation},
-{\em Python uses the Mersenne Twister as the core generator. It produces
-53-bit precision floats and has a period of 2**19937-1.} Numpy RNG is
-also based on the Mersenne Twister generator but there are differences in the
-way seed is interpreted when initializing the generator. Fortunately, Both
-Numpy and Python offer access to the internal state and we can {\em fix} it to
-make the behavior of one RNG to match the other RNG. Here, we want to make the
-Numpy RNG to match the Python RNG behavior. But there are also more sublte
-problems with this description. If we look more closely at it, we can realize
+states that {\em Python uses the Mersenne Twister as the core generator. It produces
+53-bit precision floats and has a period of 2**19937-1.} Someone trying to replicate the work however might choose to use the RNG from the \href{http://www.numpy.org/}{NumPy library}.
+The NumPy library is extensively used in the science community, and it provides an implementation of the Mersene Twister generator too.
+Unfortunately, the way the seed is interpreted by the two implementations is different, yielding different random sequences.\\
+
+%FIXME: I feel like this paragraph is openning me to attack in a way that is probably unecessary.
+Striving, as we do here, for a perfect quantitative match may seem unnecessary.
+Yet, in replication projects, quantitative comparisons are a powerful way to verify that the behavior has been reproduced: they eliminate such a quantity of possible errors that they can act, in practice, as \emph{good enough} test of replication.
+They are also particularly helpful to track exactly where the code of a tentative replication fails to reproduce the results that were published.\\
+
+Here we are be able to replicate exactly the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details (the use of the {\tt random} module of the standart Python library of CPython 3.6.1), or to the code itself.\\
+
+But there are still more sublte problems with the description given above.
+If we look more closely at it, we can realize
 that nothing is said about the specific case of {\tt 0} when generating a step.
@@
 further information and without the original code, it is up to the reader to
-decide. If this would not make a huge difference for small simulations, it
-may nonetheless compromise statistics of the results for long runs.\\
+decide.
+Likewise, the description is ambiguous regarding the first element of the walk.
+Is the initialization value included (it was not in our codes so far)?
+This slight difference changes the statistics, especially of short runs.\\
 
-Such perfect quantitative match may seem unnecessary in most cases, yet, in
-many replication projects, quantitative comparisons are an objective tests that
-the behavior is reproduced. Statistical tests demand more work, and may fall
-short of detecting small differences in the algorithms, that may results in
-huge difference of behavior in different conditions. Without such explicit
-replication effort, it would be hard to discover this faulty behavior and this
-is one of the reason why the \href{https://rescience.github.io}{ReScience
+There is no better way to ferret out the ambiguities, big and small, of an article than to replicate it. This is one of the reason why the \href{https://rescience.github.io}{ReScience
   journal} has been created. This journal {\em targets computational research and
@@
 new and open-source implementations in order to ensure that the original
-research is reproducible}.
+research is reproducible}.\\
 
@@
 def walk(rng, n):
-    """ Random walk for n steps """
-
+    """Random walk for n steps."""
     # Note: We consider 0 to be a positive step
@@
 def rng(seed):
-    """ Return a random number generator initialized with seed """
-
+    """Return a random number generator initialized with seed."""
     rng = random.Random()
@@
     seed = 1
-    x = walk(rng(seed=2), 10)
+    x = walk(rng(seed=seed), 10)
 
@@
 
-Through a very simple random walk example using Python, we've seen a variety of pitfalls of writing scientific code.
-The R$^0$ form, while easy to write and simple to understand is not a good scientific code while the full-fledge R$^{4}$ form is much preferred but requires a fair amount of work that might be too much of a burden in some circumstances.
-$R^3$ form seems to be an acceptable compromise and should be accepted as the minimum scientific standard \citep{Wilson:2017}.
+%FIXME: can still be improved
+Through a small random walk example using Python, we've seen a variety of pitfalls that are easy to remain unaware of while writing scientific code.
+The R$^0$ form, while easy to write and simple to understand is not a good scientific code while the full-fledge R$^4$ form is much preferred but requires a fair amount of work that might be too much of a burden.
+R$^3$ form seems to be an acceptable compromise and should be accepted as the minimum scientific standard \citep{Wilson:2017}.
 This means this should be actually checked by reviewers and publishers when code is part of a work worth to be published.
-But it's hardly the case today.
-%
-The choice of the Python programming language to illustrate our point was not arbitrary.
+But it's hardly the case today.\\
+
+% F: I am not sure that this is really that important a point for the conclusion.
+% The choice of the Python programming language to illustrate our point was not arbitrary.
 In less than a decade, Python has become one the few programming language of Science (together with R and Matlab).
@@
 
+Compared to psychology or biology, the replication issues of computational works have reasonable and efficient solutions.
+But making sure that these solutions are adopted will not be solved by articles such as this one.
+Just like in psychology and biology, we have to modify the incentive of the researchers, by adopting exigences, enforced domain-wide, of what constitute an acceptable scientific computational work.
+
 % Writing reproducible software requires both knowledge and experience. For the former, there are plenty of resources online or onsite.
-
-% underscore that compared to psy/bio/etc. the replication issues of CS are easily (reasonably) managed: good solutions that (mostly) works exist right now.
 

### 2017-07-24 draft.tex
Added codes for all versions

@@
 # Release under the BSD 2-clause license
-# Tested with Python 3.6 / Numpy 1.12.0 / macOS 10.12.4 / 64 bits architecture
+# Tested with Python 3.6.1 / Numpy 1.12.0 / macOS 10.12.4 / 64 bits architecture
 import random

### 2017-07-24 codes/r0.py

@@ -1,0 +1,8 @@
+Random walk (R$^0$)}
+import random
+
+x = 0
+for i in xrange(10):
+    step = random.choice([-1,+1])
+    x += step
+    print x,

### 2017-07-24 codes/r1.py

@@ -1,0 +1,10 @@
+# Tested with Python 3
+import random
+
+walk, total = [], 0
+for i in range(10):
+    step = random.choice([-1,+1])
+    total += step
+    walk.append(total)
+
+print(walk)

### 2017-07-24 codes/r2.py

@@ -1,0 +1,14 @@
+# Tested with Python 3
+import random
+
+random.seed(0) # RNG initialization
+
+walk, total = [], 0
+for i in range(10):
+    step = random.choice([-1,+1])
+    total += step
+    walk.append(total)
+
+print(walk)
+with open("results-R2.txt", "w") as file: # Saving output to disk
+    file.write(str(walk))

### 2017-07-24 codes/r5.py

@@ -1,0 +1,34 @@
+# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
+# Release under the BSD 2-clause license
+# Tested with Python 3.6.1 / Numpy 1.12.0 / macOS 10.12.4 / 64 bits architecture
+import random
+import numpy as np
+
+def walk(rng, n):
+    """Random walk for n steps."""
+    # Note: We consider 0 to be a positive step
+    steps = 2*(rng.uniform(-1,+1,n) > 0) - 1
+    return steps.cumsum().tolist()
+
+def rng(seed):
+    """Return a random number generator initialized with seed."""
+    rng = random.Random()
+    rng.seed(seed)
+    _, keys, _ = rng.getstate()
+    rng = np.random.RandomState()
+    state = rng.get_state()
+    rng.set_state((state[0], keys[:-1], state[2], state[3], state[4]))
+    return rng
+
+if __name__ == '__main__':
+    # Unit test
+    assert walk(rng(seed=1), 10) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+
+    # Random walk for 10 steps
+    seed = 1
+    x = walk(rng(seed=seed), 10)
+
+    # Display & save results
+    print(x)
+    with open("results-R5-%d.txt" % seed, "w") as file:
+        file.write(str(x))

### 2017-07-24 draft.tex
Modified slightly the conclusion

@@
 that may inspire future research (\cite{Mesirov:2010,osc:2015}). Replication
-issues have received increased attention in recent years, with a particular mediatic focus on medicine and psychology (\cite{Iqbal:2016}).
-One could think that
-computer science would mostly be shielded from such issues, since a computer program
-intrinsically describes precisely what it does and is easy to transmit perfectly to other researchers.
+issues have received increased attention in recent years, with a particular
+mediatic focus on medicine and psychology (\cite{Iqbal:2016}).  One could think
+that computer science would mostly be shielded from such issues, since a
+computer program intrinsically describes precisely what it does and is easy to
+transmit perfectly to other researchers.
 
-But precisely because it is easy to believe that if a program runs once and gives the expected
-results it will do so forever, crucial steps to transform working code into
-meaningful scientific contributions are rarely undertaken
+But precisely because it is easy to believe that if a program runs once and
+gives the expected results it will do so forever, crucial steps to transform
+working code into meaningful scientific contributions are rarely undertaken
 \citep{Sandve:2013,Schwab:2000}. Computer science is plagued by replication
-problems, in part, precisely because it seems impervious to them. A
-program can fail as a scientific contribution in many different ways for many
-different reasons. Borrowing the terms coined by \citeauthor{Goble:2016}
+problems, in part, precisely because it seems impervious to them. A program can
+fail as a scientific contribution in many different ways for many different
+reasons. Borrowing the terms coined by \citeauthor{Goble:2016}
 \citep{Goble:2016}, for your program to contribute to science, you have to make
 sure it is re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$),
-reusable (R$^4$) and replicable (R$^5$).\\
-
-Let us illustrate this with a small example, a random walk written in Python:\\
+reusable (R$^4$) and replicable (R$^5$). Let us illustrate this with a small
+example, a random walk written in Python:\\
 
@@
 def walk():
-    promenade, total = [], 0
+ea    promenade, total = [], 0
     for i in range(10):
@@
 %FIXME: can still be improved
-Through a small random walk example using Python, we've seen a variety of pitfalls that are easy to remain unaware of while writing scientific code.
-The R$^0$ form, while easy to write and simple to understand is not a good scientific code while the full-fledge R$^4$ form is much preferred but requires a fair amount of work that might be too much of a burden.
-R$^3$ form seems to be an acceptable compromise and should be accepted as the minimum scientific standard \citep{Wilson:2017}.
-This means this should be actually checked by reviewers and publishers when code is part of a work worth to be published.
-But it's hardly the case today.\\
+Through a small random walk example using Python, we've seen a variety of
+pitfalls that are easy to remain unaware of while writing scientific code. The
+R$^0$ form, while easy to write and simple to understand is not a good
+scientific code. Adding a single comment, we transformed this code into the
+R$^1$ version (rerunnable) that partially solved the very common dependencies
+problem and will allow us to re-run the program two months or two years from
+now. In the meantime, the R$^2$ version (repeatable) guarantee us (to some
+extent) we can faithfully repeat our experiment from one run to the other.  Of
+course, the full-fledge R$^4$ (reusable) form is much preferred but it requires
+a fair amount of work that might be too much of a burden. R$^3$ (reproducible)
+form seems to be an acceptable compromise and should be accepted as the minimum
+scientific standard \citep{Wilson:2017}. This means this should be actually
+checked by reviewers and publishers when code is part of a work worth to be
+published. But it's hardly the case today.\\
 

### 2017-07-25 draft.tex
Fix spelling

@@
 issues have received increased attention in recent years, with a particular
-mediatic focus on medicine and psychology (\cite{Iqbal:2016}).  One could think
+media focus on medicine and psychology (\cite{Iqbal:2016}).  One could think
 that computer science would mostly be shielded from such issues, since a
@@
 Rather, it depends on the context, and re-runnability becomes increasingly difficult as the code ages.
-Therefore, to be re-runnable in other researchers' computers and through time, a re-runnable code should describes---with enough details to be recreated---an execution environment in which it is executable.
+Therefore, to be re-runnable in other researchers' computers and through time, a re-runnable code should describe---with enough details to be recreated---an execution environment in which it is executable.
 As shown by \citep{Collberg:2016}, this is far from being either obvious or easy.\\
@@
 See the next section for details.}, which---although a perfectly valid random walk---lend itself to a gross misinterpretation of the overall dynamic of the algorithm.
-Verifying that the qualitative aspects of the results and the conclusion we make are not tied to a specific initialization of the pseudo-random generator is an integral part of any scientific undertaking in computational science;
+Verifying that the qualitative aspects of the results and the conclusions that are made are not tied to a specific initialization of the pseudo-random generator is an integral part of any scientific undertaking in computational science;
 this is usually done by repeating the simulations multiple times with different seeds.
@@
 % scripts to generates figures
-The code should also contains all the scripts that were used to produce the figures.
-While it may be considered extraneous, code for plotting figures may contains non-trivial statistics code, and is actually just as vulnerable to bugs.
+The code should also contain all the scripts that were used to produce the figures.
+While it may be considered extraneous, code for plotting figures may contain non-trivial statistics code, and is actually just as vulnerable to bugs.
 With a code that is re-runnable, repeatable and contains such scripts, a user can quickly check that published figures can be reproduced.\\
@@
 While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable.
-Code \emph{and} input \emph{and} result files \emph{and} figure data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as Figshare or Zenodo\footnote{Online code repository such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment.
+Code \emph{and} input \emph{and} result files \emph{and} figure data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as Figshare or Zenodo\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment.
 Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.}.\\
@@
 
-Reusability of course directly benefits others researchers from your team and outside of it.
+Reusability of course directly benefits other researchers from your team and outside of it.
 The easier it is to use your code, the lower the threshold is for other to study, modify and extend it.
@@
 
-Despites all this, reusability is often overlooked, and it is not hard to see why.
+Despite all this, reusability is often overlooked, and it is not hard to see why.
 Scientists are rarely trained in software engineering, and reusability can represent an expensive endeavour if undertaken as an afterthought, for little tangible short-term benefits, for a codebase that might, after all, see only a single use.
-And, in fact, reusability is not as indispensable a requirement as the others one presented here.\\
+And, in fact, reusability is not as indispensable a requirement as the other ones presented here.\\
 
-Yet, some simple measures can tremendously increase reusability, and at the same time strenghten reproducibility and re-runnability.\\
+Yet, some simple measures can tremendously increase reusability, and at the same time strengthen reproducibility and re-runnability.\\
 
@@
 
-Similarily, code behavior should be changed by commenting/uncommenting code \citep{Wilson:2017}.
-Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitely set through parameters accessible to the end-user.
-This improves reproducibility in two ways: it allows those conditions to be explicitely tracked in the result files since they contain the parameters used.
+Similarly, code behavior should be changed by commenting/uncommenting code \citep{Wilson:2017}.
+Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user.
+This improves reproducibility in two ways: it allows those conditions to be explicitly tracked in the result files since they contain the parameters used.
 It also allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
@@
 Documentation, in fact, is one of the most potent tools for reusability.
-a proper documentation on how to install and run the software often makes the difference for other reasearchers whether they manage to use the code or not.\\
+A proper documentation on how to install and run the software often makes the difference for other researchers whether they manage to use the code or not.\\
 
-Furthermore, A comment describing what each function does, however evident, can avoid hours of head-scratching.
+Furthermore, a comment describing what each function does, however evident, can avoid hours of head-scratching.
 Great code may need few comments.
@@
 
-Re-usability is not an absolute requirement for scientific code. But it has many benefit, and a few simple measure can foster it considerably.
+Re-usability is not an absolute requirement for scientific code. But it has many benefits, and a few simple measures can foster it considerably.
 
@@
 
-Here we are be able to replicate exactly the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details (the use of the {\tt random} module of the standart Python library of CPython 3.6.1), or to the code itself.\\
+Here we are able to replicate exactly the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details (the use of the {\tt random} module of the standard Python library of CPython 3.6.1), or to the code itself.\\
 
-But there are still more sublte problems with the description given above.
+But there are still more subtle problems with the description given above.
 If we look more closely at it, we can realize
 that nothing is said about the specific case of {\tt 0} when generating a step.
-Do whe have to consider {\tt 0} to be a positive or a negative step? Without
+Do we have to consider {\tt 0} to be a positive or a negative step? Without
 further information and without the original code, it is up to the reader to
@@
 
-There is no better way to ferret out the ambiguities, big and small, of an article than to replicate it. This is one of the reason why the \href{https://rescience.github.io}{ReScience
+There is no better way to ferret out the ambiguities, big and small, of an article than to replicate it. This is one of the reasons why the \href{https://rescience.github.io}{ReScience
   journal} has been created. This journal {\em targets computational research and
@@
 scientific code. Adding a single comment, we transformed this code into the
-R$^1$ version (rerunnable) that partially solved the very common dependencies
+R$^1$ version (rerunnable) that partially solved the common dependencies
 problem and will allow us to re-run the program two months or two years from
-now. In the meantime, the R$^2$ version (repeatable) guarantee us (to some
+now. In the meantime, the R$^2$ version (repeatable) guarantees us (to some
 extent) we can faithfully repeat our experiment from one run to the other.  Of
@@
 scientific standard \citep{Wilson:2017}. This means this should be actually
-checked by reviewers and publishers when code is part of a work worth to be
+checked by reviewers and publishers when code is part of a work worth being
 published. But it's hardly the case today.\\
@@
 But making sure that these solutions are adopted will not be solved by articles such as this one.
-Just like in psychology and biology, we have to modify the incentive of the researchers, by adopting exigences, enforced domain-wide, of what constitute an acceptable scientific computational work.
+Just like in psychology and biology, we have to modify the incentive of the researchers, by adopting exigences, enforced domain-wide, of what constitutes an acceptable scientific computational work.
 

### 2017-08-08 draft.tex
Added ReScience arXiv reference
Minor tweaks in R0-R2

@@
 that computer science would mostly be shielded from such issues, since a
-computer program intrinsically describes precisely what it does and is easy to
-transmit perfectly to other researchers.
+computer program intrinsically describes precisely what it does and is easily
+transferred to other researchers.
 
@@
 reasons. Borrowing the terms coined by \citeauthor{Goble:2016}
-\citep{Goble:2016}, for your program to contribute to science, you have to make
-sure it is re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$),
+\citep{Goble:2016}, for a program to contribute to science, it should be
+re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$),
 reusable (R$^4$) and replicable (R$^5$). Let us illustrate this with a small
@@
 \end{minipage}
-
-% That would be too easy if the program was commented...
-% # randomly choose ten instances of either 1 or -1.
-% # compute the cumulative sum (the sum of all preceeding elements)
 
@@
 To be clear, it is impossible to write future-proof code, and the best efforts can be stymied by the smallest change in one of the dependencies.
-At the same time, modernizing an unmaintained ten-year-old code can reveal itself to be an arduous and expensive undertaking, and dangerous as that, since each change risks to affect the semantics of the program.
-Rather trying to predict the future or painstakingly dusting off old code, an often more straightforward and less dangerous solution is to recreate the old execution environment\footnote{To be clear, and although virtual machines are often a great help here, this is not always possible.
+At the same time, modernizing an unmaintained ten-year-old code can reveal itself to be an arduous and expensive undertaking, and precarious as that, since each change risks to affect the semantics of the program.
+Rather trying to predict the future or painstakingly dusting off old code, an often more straightforward and less risky solution is to recreate the old execution environment\footnote{To be clear, and although virtual machines are often a great help here, this is not always possible.
 It is, however, \emph{always} more difficult when the original execution environment is not documented.}.
@@
 
-A \emph{re-runnable} code is one that can be run again when needed, i.e.
-more than the one time that was needed to produce the results.
+A \emph{re-runnable} code is one that can be run again when needed, and in particular more than the one time that was needed to produce the results.
 It is important to notice that the re-runnability of a code is not an intrinsic property.
@@
 As explained by \citeauthor{Donoho:2009} \parencite{Donoho:2009}, scientific practice must expect that {\em errors are ubiquitous}, and therefore be robust to them.
-Ensuring reproducibility is an important step toward this: it allows other to verify that your code does indeed produce the published results, and to scrutinize the procedure it uses to produce them.
+Ensuring reproducibility is an fundamental step toward this: it provides other researchers the means to verify that your code does indeed produce the published results, and to scrutinize the procedures it uses to produce them.
 As demonstrated by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility is hard.\\
@@
 
-There is no better way to ferret out the ambiguities, big and small, of an article than to replicate it. This is one of the reasons why the \href{https://rescience.github.io}{ReScience
-  journal} has been created. This journal {\em targets computational research and
-encourages the explicit replication of already published research, promoting
-new and open-source implementations in order to ensure that the original
-research is reproducible}.\\
+There is no better way to ferret out the ambiguities, big and small, of an
+article than to replicate it. This is one of the reasons why the ReScience
+journal \cite{Rougier:2017} has been created. This journal {\em targets
+  computational research and encourages the explicit replication of already
+  published research, promoting new and open-source implementations in order to
+  ensure that the original research is reproducible}.\\
 

### 2017-08-08 codes/r0.py

@@ -1,4 +1,3 @@
-Random walk (R$^0$)}
 import random
 
 x = 0

### 2017-08-09 draft.tex
rearranged reproducible, added git tracking, other minor edits

@@
 \begin{code}{Output}
--1  0 -1 -2 -1 -2 -3 -2 -1 -2 # with the steps being -1, +1, -1, -1, +1, -1, -1, +1, +1, -1
+-1 0 -1 -2 -1 -2 -1 -2 -3 -4 # with the steps being -1, +1, -1, -1, +1, -1, +1, -1, -1,  -1
 \end{code}
@@
 
-Because any dependency of your program---to the most basic one, the language itself---can change its behavior from one version to another, executability (R$^1$) and determinism (R$^2$) are necessary but not sufficient for reproducibility.
+% precise execution environment recording
+Because any dependency of a program---to the most basic one, the language itself---can change its behavior from one version to another, executability (R$^1$) and determinism (R$^2$) are necessary but not sufficient for reproducibility.
 The exact execution environment used to produce the results must also be specified---rather than the broadest set of environments where the code can be effectively run.
-In other words, assertions such as "the results were obtained with CPython 3.6.1" are more valuable, in a scientific context, than "the program should work with Python 3.x and above".\\
+In other words, assertions such as "the results were obtained with CPython 3.6.1" are more valuable, in a scientific context, than "the program should work with Python 3.x and above".
+With the increasing complexity of computational stacks, retrieving and deciding what is pertinent (CPU architecture?
+operating system version?
+endianness?) might be non-trivial.
+A good rule of thumb is to include more information than necessary rather than not enough.\\
 
-%FIXME: this is not a good-enough paragraph
-With the increasing complexity of our computational stacks, retrieving and deciding what is pertinent (CPU architecture?
-Operating system version?
-Endianness?) can be difficult, however.
-A best effort must be made.\\
+% if code changes, parameters are lost
+Recording the execution environment is not enough though.
+The R$^2$ program uses a random seed but does not keep a trace of it except in the code.
+Therefore, should the code change after the production of the results, someone provided with the result file will not be able to know which seed was used, and would need to iterate through all possible random seeds, an impossible task in practice\footnote{Here, with $2^{10}$ possibilities for a 10-step random walk, a seed matching the results could obviouly be found.
+For instance, seed 1123581321 matches the results of R$^0$ with Python 2.7.
+This is not possible for a 100-step walk.}.
 
-Having precisely described the execution environment, reproducibility is \emph{technically} obtained. However, \emph{reproducibility in practice} requires more than that.\\
+% 1. parameters and 2. code version with results
+From this observation, we conclude two things.
+First, results files should always come alongside their context, i.e.
+an exhaustive list of the parameters used as well as a precise description of the execution environment, as the R$^3$ code does.
+Second, the version of the code itself should be recorded as part of the provenance data.
+It is common for results to have been generated by different versions of the code.
+Ideally, all results should originate from the same (and last) version of the code.
+But for long or expensive computations, this may not be feasible.
+In that case, the result files should contains the version of the code that was used to produce it.
+This information can be obtained from the version control software.
+This also allow, if some errors are found and corrected after some results have been obtained, to identify which ones should be recomputed.
+In R$^3$, the code records the current git hash, and wether the repository holds uncommited changes when the computation starts.\\
 
-The R$^2$ program uses a random seed but does not keep a trace of it except in the code.
-Therefore, should the code change after the production of the results, someone provided with the result file will not be able to know which seed was used, and would need to iterate through all possible random seeds, an impossible task in practice.
-More generally, results files should always come alongside their context, i.e.
-an exhaustive list of the parameters used and a description of the execution environment, as the R$^3$ code does.\\
-
-% scripts to generates figures
+% scripts to generates figures & data behind graphs
 The code should also contain all the scripts that were used to produce the figures.
-While it may be considered extraneous, code for plotting figures may contain non-trivial statistics code, and is actually just as vulnerable to bugs.
-With a code that is re-runnable, repeatable and contains such scripts, a user can quickly check that published figures can be reproduced.\\
-
-% data behind graphs
+Plotting code may seem mundane, but it is as vulnerable as any other piece of the code to bugs and errors.
+With a code that is re-runnable, repeatable and contains such scripts, a user can quickly check that published figures can be reproduced.
 However, figures by themselves are imprecise and cumbersome, and sometimes plain unusable, when it comes to checking that the reproduced data match the one published in the article.
@@
 import random
+import git
+
+def git_info(): # Retrieve the code version
+    repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
+    return {'hash': repo.head.object.hexsha,
+            'dirty': repo.is_dirty()}
 
@@
             'platform' : platform.platform(),
+            'git_info' : git_info(),
             'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC
@@
 def walk():
-ea    promenade, total = [], 0
+    promenade, total = [], 0
     for i in range(10):
@@
 Dependencies and platforms must be described as precisely and a specifically as possible.
-Parameters values and inputs should be an integral part of the result files.
+Parameters values and inputs should accompany the result files.
 The data and scripts behind the graphs must be published.
@@
 Here, comments and documentation can make a significant difference.
-In fact, a source code reflects the results of all the decisions that were made to create it, but not the reasons behind the decisions.
+In fact, a source code reflects the results of all the decisions that were made to create it, but not the reasons behind those decisions.
 In science, where the method and its justification matter as much as the results, those reasons are precious knowledge.
@@
 Scientists constantly face the constraint of time: if a model is available, documented, and can be installed, run and understood all in a few hours, it will be preferred over another that would require weeks to reach the same stage.
-Moreover, scientists prefer to base their work on results they can verify.
-A reproducible and reusable code offers a platform both validated and easy-to-use, fostering the development of derivative works by other researchers.
+A reproducible and reusable code offers a platform both \emph{verifiable} and easy-to-use, fostering the development of derivative works by other researchers on solid fundations.
 Those derivative works contribute to the impact of your original contribution.\\
@@
 Scientists are rarely trained in software engineering, and reusability can represent an expensive endeavour if undertaken as an afterthought, for little tangible short-term benefits, for a codebase that might, after all, see only a single use.
-And, in fact, reusability is not as indispensable a requirement as the other ones presented here.\\
-
+And, in fact, reusability is not as indispensable a requirement as the other ones presented here.
 Yet, some simple measures can tremendously increase reusability, and at the same time strengthen reproducibility and re-runnability.\\
@@
 
-Similarly, code behavior should be changed by commenting/uncommenting code \citep{Wilson:2017}.
+Similarly, code behavior should not be changed by commenting/uncommenting code \citep{Wilson:2017}.
 Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user.
-This improves reproducibility in two ways: it allows those conditions to be explicitly tracked in the result files since they contain the parameters used.
-It also allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
+This improves reproducibility in two ways: it allows those conditions to be explicitly tracked in the result files since they contain the parameters used, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
 With a documention explaining which script or configuration file correspond to which experiment, reproducing the different figures becomes straightforward.\\
-%FIXME repeating what was said in reproducible.
 
@@
 import random
+import git
+
+def git_info():
+    """Retrieve the current git hash, and if the repository has uncommited changes."""
+    repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
+    return {'hash': repo.head.object.hexsha,
+            'dirty': repo.is_dirty()}
 
 def provenance():
-      """Retrieve data about the execution environment."""
+    """Get info about the execution environment and the code."""
     return {'python'   : {'implementation': platform.python_implementation(),
@@
     """Return a `n` steps random walk.
-
-      `seed` is used to initialized the Mersene Twister random number generator."""
+    `seed` is used to initialize the Mersene Twister random number generator."""
     random.seed(seed)
@@
 
-  # Unit test
-  assert walk(10, 1) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
-
 def generate_results(n, seed, filename='results-R4'):
@@
     w = walk(n, seed)
-
     results = {'data': w, 'n': n, 'seed': seed, 'provenance': provenance()}
@@
 if __name__ == '__main__':
+    # Unit test checking reproducibility
+    assert walk(10, 1) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+
     print(generate_results(10, 1)['data'])
@@
 % reusable is not foolproof
-Having made a software reusable offers an additional way to find bugs, especially if your scientific contribution is popular. Unfortunately, this is not always effective, and some recent cases
-have dramatically illustrated the tremendous impact a bug can have in science
+Having made a software reusable offers an additional way to find bugs, especially if your scientific contribution is popular.
+Unfortunately, this is not always effective, and some recent cases have dramatically illustrated the tremendous impact a bug can have in science
 \citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}. This is
@@
 %
-This description, while somewhat precise, forgoes---as it is common---the initialization of the variables (here the starting value of the walk ({\tt 0})), and the technical details about which implementation of the RNG is used.\\
+This description, while somewhat precise, forgoes---as it is common---the initialization of the variables (here the starting value of the walk: {\tt 0}), and the technical details about which implementation of the RNG is used.\\
 
@@
 Striving, as we do here, for a perfect quantitative match may seem unnecessary.
-Yet, in replication projects, quantitative comparisons are a powerful way to verify that the behavior has been reproduced: they eliminate such a quantity of possible errors that they can act, in practice, as \emph{good enough} test of replication.
-They are also particularly helpful to track exactly where the code of a tentative replication fails to reproduce the results that were published.\\
+Yet, in replication projects, quantitative comparisons are an effective way to verify that the behavior has been reproduced.
+Moreover, they are particularly helpful to track exactly where the code of a tentative replication fails to reproduce the published results.\\
 
@@
 Is the initialization value included (it was not in our codes so far)?
-This slight difference changes the statistics, especially of short runs.\\
+This slight difference affects statistics, especially for short runs.\\
 
@@
 article than to replicate it. This is one of the reasons why the ReScience
-journal \cite{Rougier:2017} has been created. This journal {\em targets
+journal \citep{Rougier:2017} has been created. This journal {\em targets
   computational research and encourages the explicit replication of already
@@
 %FIXME: can still be improved
-Through a small random walk example using Python, we've seen a variety of
-pitfalls that are easy to remain unaware of while writing scientific code. The
+Through a small random walk example implemented in Python, we've seen that writing good scientific code is not trivial. The
 R$^0$ form, while easy to write and simple to understand is not a good
@@
 But making sure that these solutions are adopted will not be solved by articles such as this one.
-Just like in psychology and biology, we have to modify the incentive of the researchers, by adopting exigences, enforced domain-wide, of what constitutes an acceptable scientific computational work.
+Just like in psychology and biology, we have to modify the incentive of the researchers, by adopting exigences, enforced domain-wide, on what constitutes an acceptable scientific computational work.
 

### 2017-08-09 codes/r2.py

@@ -1,11 +1,12 @@
 # Tested with Python 3
 import random
 
-random.seed(0) # RNG initialization
+random.seed(1123581321) # RNG initialization
 
 walk, total = [], 0
 for i in range(10):
     step = random.choice([-1,+1])
+    print(step)
     total += step
     walk.append(total)
 

### 2017-08-09 codes/r3.py

@@ -3,6 +3,12 @@
 # Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
 import platform, datetime
 import random
+import git
+
+def git_info(): # Retrieve the code version
+    repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
+    return {'hash': repo.head.object.hexsha,
+            'dirty': repo.is_dirty()}
 
 def provenance(): # Get info about the execution environment
     return {'python'   : {'implementation': platform.python_implementation(),
@@ -11,6 +17,7 @@ def provenance(): # Get info about the execution environment
                           'branch'        : platform.python_branch(),
                           'revision'      : platform.python_revision()},
             'platform' : platform.platform(),
+            'git_info' : git_info(),
             'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC
 
 def walk():

### 2017-08-09 codes/r4.py

@@ -3,21 +3,28 @@
 # Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
 import platform, datetime, json
 import random
+import git
+
+def git_info():
+    """Retrieve the current git hash, and if the repository has uncommited changes."""
+    repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
+    return {'hash': repo.head.object.hexsha,
+            'dirty': repo.is_dirty()}
 
 def provenance():
-    """Retrieve data about the execution environment."""
+    """Get info about the execution environment and the code."""
     return {'python'   : {'implementation': platform.python_implementation(),
                           'version'       : platform.python_version_tuple(),
                           'compiler'      : platform.python_compiler(),
                           'branch'        : platform.python_branch(),
                           'revision'      : platform.python_revision()},
             'platform' : platform.platform(),
+            'git_info' : git_info(),
             'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC
 
 def walk(n, seed):
     """Return a `n` steps random walk.
-
-    `seed` is used to initialized the Mersene Twister random number generator."""
+    `seed` is used to initialize the Mersene Twister random number generator."""
     random.seed(seed)
     rndwalk, total = [], 0
     for i in range(10):
@@ -26,17 +33,16 @@ def walk(n, seed):
         rndwalk.append(total)
     return rndwalk
 
-# Unit test
-assert walk(10, 1) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
-
 def generate_results(n, seed, filename='results-R4'):
     """Generate results and save them to disk as a JSON file"""
     w = walk(n, seed)
-
     results = {'data': w, 'n': n, 'seed': seed, 'provenance': provenance()}
     with open(filename + '.json', 'w') as fd:
         json.dump(results, fd)
     return results
 
 if __name__ == '__main__':
+    # Unit test checking reproducibility
+    assert walk(10, 1) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+
     print(generate_results(10, 1)['data'])

### 2017-08-09 codes/r2.py
Restore r2 code

@@ -1,12 +1,11 @@
 # Tested with Python 3
 import random
 
-random.seed(1123581321) # RNG initialization
+random.seed(0) # RNG initialization
 
 walk, total = [], 0
 for i in range(10):
     step = random.choice([-1,+1])
-    print(step)
     total += step
     walk.append(total)
 

### 2017-08-17 draft.tex
Update codes, fix spelling

@@
 The R$^2$ program uses a random seed but does not keep a trace of it except in the code.
-Therefore, should the code change after the production of the results, someone provided with the result file will not be able to know which seed was used, and would need to iterate through all possible random seeds, an impossible task in practice\footnote{Here, with $2^{10}$ possibilities for a 10-step random walk, a seed matching the results could obviouly be found.
+Therefore, should the code change after the production of the results, someone provided with the result file will not be able to know which seed was used, and would need to iterate through all possible random seeds, an impossible task in practice\footnote{Here, with $2^{10}$ possibilities for a 10-step random walk, a seed matching the results could obviously be found.
 For instance, seed 1123581321 matches the results of R$^0$ with Python 2.7.
@@
 But for long or expensive computations, this may not be feasible.
-In that case, the result files should contains the version of the code that was used to produce it.
+In that case, the result files should contain the version of the code that was used to produce it.
 This information can be obtained from the version control software.
-This also allow, if some errors are found and corrected after some results have been obtained, to identify which ones should be recomputed.
-In R$^3$, the code records the current git hash, and wether the repository holds uncommited changes when the computation starts.\\
+This also allows, if some errors are found and corrected after some results have been obtained, to identify which ones should be recomputed.
+In R$^3$, the code records the current git hash, and whether the repository holds uncommited changes when the computation starts.\\
 
@@
 import random
+import pip
 import git
@@
     repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
-    return {'hash': repo.head.object.hexsha,
-            'dirty': repo.is_dirty()}
+    return {'hash': repo.head.object.hexsha, 'dirty': repo.is_dirty(),
+            'git_version': str(pip.git.Git().get_git_version())}
 
@@
             'platform' : platform.platform(),
+            'packages'  : list(pip.commands.freeze.freeze()), # list of installed packages
             'git_info' : git_info(),
@@
 Scientists constantly face the constraint of time: if a model is available, documented, and can be installed, run and understood all in a few hours, it will be preferred over another that would require weeks to reach the same stage.
-A reproducible and reusable code offers a platform both \emph{verifiable} and easy-to-use, fostering the development of derivative works by other researchers on solid fundations.
+A reproducible and reusable code offers a platform both \emph{verifiable} and easy-to-use, fostering the development of derivative works by other researchers on solid foundations.
 Those derivative works contribute to the impact of your original contribution.\\
@@
 If people start using your program, they will most likely report bugs or malfunctions they encounter.
-If you're lucky enough, they might even propose either bugfixes or improvements, hence improving the overall quality of your software.
+If you're lucky enough, they might even propose either bug fixes or improvements, hence improving the overall quality of your software.
 This process contributes to the long-term reproducibility to the extent people continue to use and maintain the program.\\

### 2017-08-17 codes/r3.py

@@ -3,12 +3,13 @@
 # Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
 import platform, datetime
 import random
+import pip
 import git
 
 def git_info(): # Retrieve the code version
     repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
-    return {'hash': repo.head.object.hexsha,
-            'dirty': repo.is_dirty()}
+    return {'hash': repo.head.object.hexsha, 'dirty': repo.is_dirty(),
+            'git_version': str(pip.git.Git().get_git_version())}
 
 def provenance(): # Get info about the execution environment
     return {'python'   : {'implementation': platform.python_implementation(),
@@ -17,6 +18,7 @@ def provenance(): # Get info about the execution environment
                           'branch'        : platform.python_branch(),
                           'revision'      : platform.python_revision()},
             'platform' : platform.platform(),
+            'packages'  : list(pip.commands.freeze.freeze()), # list of installed packages
             'git_info' : git_info(),
             'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC
 

### 2017-08-17 codes/r4.py

@@ -3,13 +3,14 @@
 # Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
 import platform, datetime, json
 import random
+import pip
 import git
 
 def git_info():
     """Retrieve the current git hash, and if the repository has uncommited changes."""
     repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
-    return {'hash': repo.head.object.hexsha,
-            'dirty': repo.is_dirty()}
+    return {'hash': repo.head.object.hexsha, 'dirty': repo.is_dirty(),
+            'git_version': str(pip.git.Git().get_git_version())}
 
 def provenance():
     """Get info about the execution environment and the code."""
@@ -19,6 +20,7 @@ def provenance():
                           'branch'        : platform.python_branch(),
                           'revision'      : platform.python_revision()},
             'platform' : platform.platform(),
+            'packages'  : list(pip.commands.freeze.freeze()), # list of installed packages
             'git_info' : git_info(),
             'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC
 

### 2017-08-19 draft.tex
Last pass, improving language (WIP).

@@
 that may inspire future research (\cite{Mesirov:2010,osc:2015}). Replication
-issues have received increased attention in recent years, with a particular
-media focus on medicine and psychology (\cite{Iqbal:2016}).  One could think
+issues have received increased attention in recent years, with a particular focus on medicine and psychology (\cite{Iqbal:2016}).  One could think
 that computer science would mostly be shielded from such issues, since a
 computer program intrinsically describes precisely what it does and is easily
-transferred to other researchers.
+disseminated to other researchers without alteration.
 
@@
 To be clear, it is impossible to write future-proof code, and the best efforts can be stymied by the smallest change in one of the dependencies.
-At the same time, modernizing an unmaintained ten-year-old code can reveal itself to be an arduous and expensive undertaking, and precarious as that, since each change risks to affect the semantics of the program.
+At the same time, modernizing an unmaintained ten-year-old code can reveal itself to be an arduous and expensive undertaking, and precarious as that, since each change risks affecting the semantics of the program.
 Rather trying to predict the future or painstakingly dusting off old code, an often more straightforward and less risky solution is to recreate the old execution environment\footnote{To be clear, and although virtual machines are often a great help here, this is not always possible.
@@
 In our case, the R$^0$ version of our tiny walker seems to imply that any version of Python would be fine.
-This not the case because it uses the print {\em instruction} and the {\tt xrange} operator, both specific to Python 2.
+This not the case: it uses the print {\em instruction} and the {\tt xrange} operator, both specific to Python 2.
 The print {\em instruction}, available in Python 2 (a version still widely used; support is scheduled to stop in 2020), has been deprecated in Python 3 (first released in 2008, almost a decade ago) in favor or a  print {\em function}, while the {\tt xrange} operator has been replaced by the {\tt range} operator in Python 3.
@@
 As explained by \citeauthor{Donoho:2009} \parencite{Donoho:2009}, scientific practice must expect that {\em errors are ubiquitous}, and therefore be robust to them.
-Ensuring reproducibility is an fundamental step toward this: it provides other researchers the means to verify that your code does indeed produce the published results, and to scrutinize the procedures it uses to produce them.
+Ensuring reproducibility is an fundamental step toward this: it provides other researchers the means to verify that your code does indeed produce the published results, and to scrutinize the procedures it used to produce them.
 As demonstrated by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility is hard.\\
@@
 Executed with Python 2.7 to 3.2, the code will produce the sequence 1, 2, 1, 2, 3, 4, 5, 6, 7, 6.
-But with Python 3.3 to 3.6, it will produce 1, 2, 1, 0, 1, 0, 1, 0, -1, 0.\\
+But with Python 3.3 to 3.6, it will produce 1, 2, 1, 0, 1, 0, 1, 0, -1, 0.
+With future version of the language, it may change still.\\
 
@@
 The exact execution environment used to produce the results must also be specified---rather than the broadest set of environments where the code can be effectively run.
-In other words, assertions such as "the results were obtained with CPython 3.6.1" are more valuable, in a scientific context, than "the program should work with Python 3.x and above".
+In other words, assertions such as "the results were obtained with CPython 3.6.1" are more valuable, in a scientific context, than "the program works with Python 3.x and above".
 With the increasing complexity of computational stacks, retrieving and deciding what is pertinent (CPU architecture?
@@
 endianness?) might be non-trivial.
-A good rule of thumb is to include more information than necessary rather than not enough.\\
+A good rule of thumb is to include more information than necessary rather than not enough, and some rather than none.\\
 
@@
 The R$^2$ program uses a random seed but does not keep a trace of it except in the code.
-Therefore, should the code change after the production of the results, someone provided with the result file will not be able to know which seed was used, and would need to iterate through all possible random seeds, an impossible task in practice\footnote{Here, with $2^{10}$ possibilities for a 10-step random walk, a seed matching the results could obviously be found.
+Should the code change after the production of the results, someone provided with the last version of the code will not be able to know which seed was used to produce the results, and would need to iterate through all possible random seeds, an impossible task in practice\footnote{Here, with $2^{10}$ possibilities for a 10-step random walk, a seed matching the results could obviously be found.
 For instance, seed 1123581321 matches the results of R$^0$ with Python 2.7.
-This is not possible for a 100-step walk.}.
+This is not possible for a 100-step walk.}.\\
 
 % 1. parameters and 2. code version with results
-From this observation, we conclude two things.
-First, results files should always come alongside their context, i.e.
+This is why results files should come alongside their context, i.e.
 an exhaustive list of the parameters used as well as a precise description of the execution environment, as the R$^3$ code does.
-Second, the version of the code itself should be recorded as part of the provenance data.
-It is common for results to have been generated by different versions of the code.
+The code itself is part of that context: the version of the code must be recorded as part of the provenance data.
+It is common for different results or different figures to have been generated by different versions of the code.
 Ideally, all results should originate from the same (and last) version of the code.
@@
 Should this test fail, then there is little hope of reproducing the results.
-Of course, passing the test is no definitive proof.\\
+Of course, passing the test does not guarantee anything.\\
 

### 2017-08-20 draft.tex
Improving the replicable section. Still needs some love.

@@
 This also allows, if some errors are found and corrected after some results have been obtained, to identify which ones should be recomputed.
-In R$^3$, the code records the current git hash, and whether the repository holds uncommited changes when the computation starts.\\
+In R$^3$, the code records the last git commit hash, and whether the repository holds uncommited changes when the computation starts.\\
 
 % scripts to generates figures & data behind graphs
-The code should also contain all the scripts that were used to produce the figures.
+Published results should come from version of the code where every file has been commited.
+This includes pre-processing, post-processing and plotting code.
 Plotting code may seem mundane, but it is as vulnerable as any other piece of the code to bugs and errors.
-With a code that is re-runnable, repeatable and contains such scripts, a user can quickly check that published figures can be reproduced.
-However, figures by themselves are imprecise and cumbersome, and sometimes plain unusable, when it comes to checking that the reproduced data match the one published in the article.
-A published figure should always be accompanied by its underlying data to allow straightforward numeric comparisons.\\
+With a code that is re-runnable, repeatable and contains plotting scripts, a user can quickly reproduce published figures.
+When it comes to checking that the reproduced data match the one published in the article, however, figures can reveal themselves to be imprecise and cumbersome, and sometimes plain unusable.
+Ideally, published figures should be accompanied by their underlying data (coordinates of the plotted points) in the supplementary data to allow straightforward numeric comparisons.\\
 
@@
 While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable.
-Code \emph{and} input \emph{and} result files \emph{and} figure data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as Figshare or Zenodo\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment.
+Code \emph{and} input data \emph{and} result data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as Figshare or Zenodo\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment.
 Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.}.\\
@@
 To recap, reproducibility implies re-runnability and repeatability and availability, yet imposes additional conditions.
-Dependencies and platforms must be described as precisely and a specifically as possible.
+Dependencies and platforms must be described as precisely and as specifically as possible.
 Parameters values and inputs should accompany the result files.
@@
 Here, comments and documentation can make a significant difference.
-In fact, a source code reflects the results of all the decisions that were made to create it, but not the reasons behind those decisions.
+Source code reflect the results of the decisions that were made during its creation, but not the reasons behind those decisions.
 In science, where the method and its justification matter as much as the results, those reasons are precious knowledge.
 In that context, a comment on how a given parameter was chosen (optimization, experimental data, educated guess), why a library was chosen over another (conceptual or technical reasons?) is valuable information.\\
+
+% Research team typically experience a high turnover of interns, doctoral and post-doctoral students. Can you reproduce the results of member after leave your team?
 
@@
 And, in fact, reusability is not as indispensable a requirement as the other ones presented here.
-Yet, some simple measures can tremendously increase reusability, and at the same time strengthen reproducibility and re-runnability.\\
+Yet, some simple measures can tremendously increase reusability, and at the same time strengthen reproducibility and re-runnability over the long-term.\\
 
@@
 In the R$^3$ code, the seed and the number of steps are respectively hardcoded and magic.\\
-
-Similarly, code behavior should not be changed by commenting/uncommenting code \citep{Wilson:2017}.
-Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user.
-This improves reproducibility in two ways: it allows those conditions to be explicitly tracked in the result files since they contain the parameters used, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
-With a documention explaining which script or configuration file correspond to which experiment, reproducing the different figures becomes straightforward.\\
-
-Documentation, in fact, is one of the most potent tools for reusability.
-A proper documentation on how to install and run the software often makes the difference for other researchers whether they manage to use the code or not.\\
-
-Furthermore, a comment describing what each function does, however evident, can avoid hours of head-scratching.
-Great code may need few comments.
-Scientists, however, are not always brilliant developers.
-Code quality, in science, averages on mediocre.
-Of course, bad, complicated code should be rewritten until is simple enough to explain itself.
-But realistically, this is hardly ever going to be done: there is simply not enough incentive for it.
-There, a comment that explains the intention and scientific reasons behind a block of code can tremendously useful.\\
 
@@
 
-Re-usability is not an absolute requirement for scientific code. But it has many benefits, and a few simple measures can foster it considerably.
+Similarly, code behavior should not be changed by commenting/uncommenting code \citep{Wilson:2017}.
+Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user.
+This improves reproducibility in two ways: it allows those conditions to be explicitly tracked in the result files since they contain the parameters used, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
+With a documention explaining which script or configuration file correspond to which experiment, reproducing the different figures becomes straightforward.\\
+
+Documentation is one of the most potent tools for reusability.
+A proper documentation on how to install and run the software often makes the difference whether other researchers manage to use the code or not.
+A comment describing what each function does, however evident, can avoid hours of head-scratching.
+Great code may need few comments.
+Scientists, however, are not always brilliant developers.
+Code quality, in science, averages on mediocre.
+Of course, bad, complicated code should be rewritten until is simple enough to explain itself.
+But realistically, this is not always going to be done: there is simply not enough incentive for it.
+There, a comment that explains the intention and scientific reasons behind a block of code can tremendously useful.\\
+
+Reusability is not a strict requirement for scientific code. But it has many benefits, and a few simple measures can foster it considerably.
+
+% But a few good practices, observed during the development of the code can have significant benefits for you, your team and any other reasearche desirous to use your code.
 
@@
 % reusable is not foolproof
-Having made a software reusable offers an additional way to find bugs, especially if your scientific contribution is popular.
+Having made a software reusable offers an additional way to find errors, especially if your scientific contribution is popular.
 Unfortunately, this is not always effective, and some recent cases have dramatically illustrated the tremendous impact a bug can have in science
@@
 This is because most papers strive to communicate the main ideas behind their contribution is terms as simple and as clear as possible, so that the reader may be able to easily understand them and the results that are presented to him.
-Trying to ensure replicability in the main text adds a myriad of esoteric details that are not conceptually significant, and clutter the explanations.
+Trying to ensure replicability in the main text adds a myriad of esoteric details that are not conceptually significant and clutter the explanations.
 Therefore, unless the writer dedicates an addendum or a section of the supplementary information for technical details specifically aimed at replicability, the information will not be there because there are incentives not to do so.\\
@@
 % errors may always remain: reproducibility does not make reproducibility useless.
-But even when those details are present, the best efforts may fall short because
-an oversight, a typo or a difference between what is evident for the writer
-and for the reader \citep{Mesnard:2016}. In that case, a reproducible code is a crucial backup. Hence,
-replicability does not negate the necessity of reproducibility. To illustrate
-this, let us consider what could be the textual description of the random
-walker (as it would be written in an article describing it):
+But even when those details are present, the best efforts may fall short because an oversight, a typo or a difference between what is evident for the writer and for the reader \citep{Mesnard:2016}.
+Minute changes in the numerical estimation of a common first-order differential equation can have significant impact \citep{Crook:2013}.
+A reproducible code plays a role that cannot be adequately fulfilled by the conceptual description of it found in the article.
+Hence, replicability does not negate the necessity of reproducibility.
+To illustrate this, let us consider what could be the textual description of the random walker (as it would be written in an article describing it):
 %
@@
 \href{https://docs.python.org/3.6/library/random.html}{Python documentation},
-states that {\em Python uses the Mersenne Twister as the core generator. It produces
-53-bit precision floats and has a period of 2**19937-1.} Someone trying to replicate the work however might choose to use the RNG from the \href{http://www.numpy.org/}{NumPy library}.
+states that "Python uses the Mersenne Twister as the core generator. It produces
+53-bit precision floats and has a period of 2**19937-1".
+Someone trying to replicate the work however might choose to use the RNG from the \href{http://www.numpy.org/}{NumPy library}.
 The NumPy library is extensively used in the science community, and it provides an implementation of the Mersene Twister generator too.

### 2017-08-21 draft.tex
Rewrote conclusion to sum up the main rationales.

@@
 This also allows, if some errors are found and corrected after some results have been obtained, to identify which ones should be recomputed.
-In R$^3$, the code records the last git commit hash, and whether the repository holds uncommited changes when the computation starts.\\
+In R$^3$, the code records the last git commit hash, and whether the repository holds uncommitted changes when the computation starts.\\
 
 % scripts to generates figures & data behind graphs
-Published results should come from version of the code where every file has been commited.
+Published results should come from version of the code where every file has been committed.
 This includes pre-processing, post-processing and plotting code.
@@
 def git_info():
-    """Retrieve the current git hash, and if the repository has uncommited changes."""
+    """Retrieve the current git hash, and if the repository has uncommitted changes."""
     repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
@@
 This improves reproducibility in two ways: it allows those conditions to be explicitly tracked in the result files since they contain the parameters used, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
-With a documention explaining which script or configuration file correspond to which experiment, reproducing the different figures becomes straightforward.\\
+With a documentation explaining which script or configuration file corresponds to which experiment, reproducing the different figures becomes straightforward.\\
 
@@
 
-%FIXME: I feel like this paragraph is openning me to attack in a way that is probably unecessary.
-Striving, as we do here, for a perfect quantitative match may seem unnecessary.
+Here we are able to replicate exactly\footnote{Striving, as we do here, for a perfect quantitative match may seem unnecessary.
 Yet, in replication projects, quantitative comparisons are an effective way to verify that the behavior has been reproduced.
-Moreover, they are particularly helpful to track exactly where the code of a tentative replication fails to reproduce the published results.\\
-
-Here we are able to replicate exactly the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details (the use of the {\tt random} module of the standard Python library of CPython 3.6.1), or to the code itself.\\
-
-But there are still more subtle problems with the description given above.
-If we look more closely at it, we can realize
-that nothing is said about the specific case of {\tt 0} when generating a step.
-Do we have to consider {\tt 0} to be a positive or a negative step? Without
-further information and without the original code, it is up to the reader to
-decide.
-Likewise, the description is ambiguous regarding the first element of the walk.
-Is the initialization value included (it was not in our codes so far)?
-This slight difference affects statistics, especially for short runs.\\
-
-There is no better way to ferret out the ambiguities, big and small, of an
-article than to replicate it. This is one of the reasons why the ReScience
-journal \citep{Rougier:2017} has been created. This journal {\em targets
-  computational research and encourages the explicit replication of already
-  published research, promoting new and open-source implementations in order to
-  ensure that the original research is reproducible}.\\
+Moreover, they are particularly helpful to track exactly where the code of a tentative replication fails to reproduce the published results.} the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details (the use of the {\tt random} module of the standard Python library of CPython 3.6.1), or to the code itself.\\
 
@@
 
+But there are still more subtle problems with the description given above.
+If we look more closely at it, we can realize
+that nothing is said about the specific case of {\tt 0} when generating a step.
+Do we have to consider {\tt 0} to be a positive or a negative step? Without
+further information and without the original code, it is up to the reader to
+decide.
+Likewise, the description is ambiguous regarding the first element of the walk.
+Is the initialization value included (it was not in our codes so far)?
+This slight difference affects the statistics of short runs.\\
+
+All these ambiguities in the description of an algorithm pile up; some are inconsequential (the 0 case has null probability), but some may affect the results in important ways.
+They are mostly inconspicuous to the reader and oftentimes, to the writer as well.
+In fact,  there is no better way to ferret out the ambiguities, big and small, of an article than to replicate it.
+This is one of the reasons why the ReScience journal \citep{Rougier:2017} has been created.
+This journal {\em targets  computational research and encourages the explicit replication of already published research, promoting new and open-source implementations in order to ensure that the original research is reproducible}.\\
+
 % \clearpage
@@
 
-%FIXME: can still be improved
-Through a small random walk example implemented in Python, we've seen that writing good scientific code is not trivial. The
-R$^0$ form, while easy to write and simple to understand is not a good
-scientific code. Adding a single comment, we transformed this code into the
-R$^1$ version (rerunnable) that partially solved the common dependencies
-problem and will allow us to re-run the program two months or two years from
-now. In the meantime, the R$^2$ version (repeatable) guarantees us (to some
-extent) we can faithfully repeat our experiment from one run to the other.  Of
-course, the full-fledge R$^4$ (reusable) form is much preferred but it requires
-a fair amount of work that might be too much of a burden. R$^3$ (reproducible)
-form seems to be an acceptable compromise and should be accepted as the minimum
+Throughout the evolution of a small random walk example implemented in Python, we illustrated some of the issue that may plague code that work and produce correct results, and reduce its value as an element of scientific knowledge.
+Further, we articulated characteristics that a code should possess to be a useful part of a scientific publication: it should be rerunnable, repeatable, reproducible, reusable and replicable.\\
+
+Running old code on tomorrow’s computer and software stacks may not be possible.
+But recreating the old code’s execution environment may be: to ensure that the long-term rerunnability of a code, its execution environment must be documented.
+For our example, a single comment went a long way to go from the R$^0$ to the R$^1$ (rerunnable) code.\\
+
+Science is built on verifying the results of others.
+This is harder to do if each execution of the code produce a different result.
+While for complex parallel workflow this may not be possible, in all instances where it is feasible the code should be repeatable.
+This allows future researchers to examine exactly how a specific result was produced.
+Most of the time, what is needed is to set or record the initial state of the pseudo-random number generator, as what done in the R$^2$ (repeatable) version.\\
+
+Even more care is needed to make a code reproducible.
+The exact execution environment, code and parameters used must be recorded and embedded in the results files.
+The R$^3$ (reproducible) version implements provenance methods to that end.
+Furthermore, the code must be made available as supplementary data with the whole computational workflow, from preprocessing steps to plotting scripts.\\
+
+Making code reusable is a stretch goal that can yield tremendous benefits for you, your team and other researchers.
+Taken into account during development rather than as an afterthought, simple measures can avoid hours of head-scratching for others, and for yourself—in a few years.
+Documentation is paramount here, even a single comment per function, as was done in the R$^4$ (reusable) version.\\
+
+Finally, there is the belief that an article should suffice itself: the description of the algorithms present in the paper should suffice to re-obtain (to replicate) the published results.
+For well-written papers that precisely dissociate conceptually significant details from irrelevant implementation details, that may be.
+But science should not assume the best of cases.
+Science assumes that errors are ubiquitous.
+Every paper is a mistake or a forgotten parameter away from irreproducibility.
+Replication efforts use the paper first, and then the reproducible code that comes along with it whenever the paper falls short of being precise enough to be reimplemented.\\
+
+In conclusion, the R$^3$ (reproducible) form should be accepted as the minimum
 scientific standard \citep{Wilson:2017}. This means this should be actually

### 2017-08-23 draft.tex
Updated and simplified sources

@@
 
-walk, total = [], 0
+x =  0
+path = []
 for i in range(10):
     step = random.choice([-1,+1])
-    total += step
-    walk.append(total)
+    x += step
+    path.append(x)
 
-print(walk)
-with open("results-R2.txt", "w") as file: # Saving output to disk
-    file.write(str(walk))
+print(path)
+# Saving output to disk
+with open("results-R2.txt", "w") as fd:
+    fd.write(str(path))
+
 \end{code}
@@
 # Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
-import platform, datetime
-import random
-import pip
-import git
-
-def git_info(): # Retrieve the code version
-    repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
-    return {'hash': repo.head.object.hexsha, 'dirty': repo.is_dirty(),
-            'git_version': str(pip.git.Git().get_git_version())}
-
-def provenance(): # Get info about the execution environment
-    return {'python'   : {'implementation': platform.python_implementation(),
-                          'version'       : platform.python_version_tuple(),
-                          'compiler'      : platform.python_compiler(),
-                          'branch'        : platform.python_branch(),
-                          'revision'      : platform.python_revision()},
-            'platform' : platform.platform(),
-            'packages'  : list(pip.commands.freeze.freeze()), # list of installed packages
-            'git_info' : git_info(),
-            'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC
+import sys, subprocess, datetime, random
 
 def walk():
-    promenade, total = [], 0
+    path = []
+    x = 0
     for i in range(10):
-        step = +1 if random.uniform(-1,+1) > 0 else -1
-        total += step
-        promenade.append(total)
-    return promenade
+        if random.uniform(-1,+1) > 0:
+            x = x + 1
+        else:
+            x = x - 1
+        path.append(x)
+    return path
+
+# If repository is dirty, don't run anything
+if subprocess.call(("git", "diff-index", "--quiet", "HEAD")):
+    print("Repository is dirty, please commit first")
+    sys.exit(1)
+
+# Get git hash if any
+revision = subprocess.check_output(("git", "rev-parse", "HEAD"))
 
@@
 random.seed(seed)
-w = walk()
+path = walk()
 
 # Display & save results
-print(w)
-results = {'data': w, 'seed': seed, 'provenance': provenance()}
+print(path)
+results = { 'data':      path,
+            'seed':      seed,
+            'timestamp': str(datetime.datetime.utcnow()),
+            'revision':  revision,
+            'system' :   sys.version }
 with open("results-R3.txt", "w") as fd:
@@
 # Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
-import platform, datetime, json
-import random
-import pip
-import git
+import sys, subprocess, datetime, random
 
-def git_info():
-    """Retrieve the current git hash, and if the repository has uncommitted changes."""
-    repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
-    return {'hash': repo.head.object.hexsha, 'dirty': repo.is_dirty(),
-            'git_version': str(pip.git.Git().get_git_version())}
-
-def provenance():
-    """Get info about the execution environment and the code."""
-    return {'python'   : {'implementation': platform.python_implementation(),
-                          'version'       : platform.python_version_tuple(),
-                          'compiler'      : platform.python_compiler(),
-                          'branch'        : platform.python_branch(),
-                          'revision'      : platform.python_revision()},
-            'platform' : platform.platform(),
-            'packages'  : list(pip.commands.freeze.freeze()), # list of installed packages
-            'git_info' : git_info(),
-            'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC
-
-def walk(n, seed):
-    """Return a `n` steps random walk.
-    `seed` is used to initialize the Mersene Twister random number generator."""
+def walk(x0=0, step=1, count=10, seed=0):
+    """ Random walk
+        x0   : initial position (default 0)
+        step : step size (default 1)
+        count: number of steps (default 10)
+        seed : seed for the initialization of the random generator (default 0)
+    """
+    path, x = [], x0
     random.seed(seed)
-    rndwalk, total = [], 0
-    for i in range(10):
-        step = +1 if random.uniform(-1,+1) > 0 else -1
-        total += step
-        rndwalk.append(total)
-    return rndwalk
-
-def generate_results(n, seed, filename='results-R4'):
-    """Generate results and save them to disk as a JSON file"""
-    w = walk(n, seed)
-    results = {'data': w, 'n': n, 'seed': seed, 'provenance': provenance()}
-    with open(filename + '.json', 'w') as fd:
-        json.dump(results, fd)
-    return results
+    for i in range(count):
+        if random.uniform(-1,+1) > 0:
+            x = x + step
+        else:
+            x = x - step
+        path.append(x)
+    return path
 
 if __name__ == '__main__':
+    # If repository is dirty, don't do anything
+    if subprocess.call(("git", "diff-index", "--quiet", "HEAD")):
+        print("Repository is dirty, please commit first")
+        sys.exit(1)
+
+    # Get git hash if any
+    revision = subprocess.check_output(("git", "rev-parse", "HEAD"))
+
     # Unit test checking reproducibility
-    assert walk(10, 1) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+    assert walk(0, 1, 10, 1) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
 
-    print(generate_results(10, 1)['data'])
+    # Simulation parameters
+    parameters = { 'x0':    0,
+                   'step':  1,
+                   'count': 10,
+                   'seed' : 1 }
+    path = walk(**parameters)
+    results = {'data':       path,
+               'parameters': parameters,
+               'timestamp':  str(datetime.datetime.utcnow()),
+               'revision':   revision,
+               'system':     sys.version}
+
+    # Save & display results
+    with open("results-R4.txt", "w") as fd:
+        fd.write(str(results))
+
+    print(path)
 \end{code}
@@
 
-def walk(rng, n):
-    """Random walk for n steps."""
-    # Note: We consider 0 to be a positive step
-    steps = 2*(rng.uniform(-1,+1,n) > 0) - 1
-    return steps.cumsum().tolist()
-
-def rng(seed):
-    """Return a random number generator initialized with seed."""
+def _rng(seed):
+    """ Return a numpy random number generator initialized with seed
+        as it would be with python random generator.
+    """
     rng = random.Random()
@@
 
+def walk(n, seed):
+    """ Random walk for n steps """
+
+    rng = _rng(seed)
+    steps = 2*(rng.uniform(-1,+1,n) > 0) - 1
+    return steps.cumsum().tolist()
+
 if __name__ == '__main__':
     # Unit test
-    assert walk(rng(seed=1), 10) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+    assert walk(n=10, seed=1) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
 
-    # Random walk for 10 steps
+    # Random walk for 10 steps, initialization with seed=1
     seed = 1
-    x = walk(rng(seed=seed), 10)
+    path = walk(n=10, seed=seed)
 
-    # Display & save results
-    print(x)
-    with open("results-R5-%d.txt" % seed, "w") as file:
-        file.write(str(x))
+    # Save & display results
+    results = {'data': path, 'seed': seed}
+    with open("results-R5.txt", "w") as fd:
+        fd.write(str(results))
+    print(path)
+
 \end{code}

### 2017-08-24 draft.tex
Update codes from random-walk repository

@@
 for i in xrange(10):
-    step = random.choice([-1,+1])
-    x += step
+    if random.uniform(-1, +1) > 0:
+        x += 1
+    else:
+        x -= 1
     print x,
@@
 \begin{code}{Output}
--1 0 -1 -2 -1 -2 -1 -2 -3 -4 # with the steps being -1, +1, -1, -1, +1, -1, +1, -1, -1,  -1
+-1 0 -1 0 -1 0 -1 0 1 2 # with the steps being -1, +1, -1, +1, -1, +1, -1, +1, +1, +1
 \end{code}
@@
 
-walk, total = [], 0
+x =  0
+walk = []
 for i in range(10):
-    step = random.choice([-1,+1])
-    total += step
-    walk.append(total)
+    if random.uniform(-1, +1) > 0:
+        x += 1
+    else:
+        x -= 1
+    walk.append(x)
 
@@
 x =  0
-path = []
+walk = []
 for i in range(10):
-    step = random.choice([-1,+1])
-    x += step
-    path.append(x)
+    if random.uniform(-1, +1) > 0:
+        x += 1
+    else:
+        x -= 1
+    walk.append(x)
 
-print(path)
+print(walk)
 # Saving output to disk
-with open("results-R2.txt", "w") as fd:
-    fd.write(str(path))
-
+with open('results-R2.txt', 'w') as fd:
+    fd.write(str(walk))
 \end{code}
@@
 # Release under the BSD 2-clause license
-# Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
+# Tested with CPython 3.6.2 / macOS 10.12.6 / 64 bits architecture
 import sys, subprocess, datetime, random
 
-def walk():
-    path = []
+def generate_walk():
+    walk = []
     x = 0
     for i in range(10):
-        if random.uniform(-1,+1) > 0:
-            x = x + 1
+        if random.uniform(-1, +1) > 0:
+            x += 1
         else:
-            x = x - 1
-        path.append(x)
-    return path
+            x -= 1
+        walk.append(x)
+    return walk
 
 # If repository is dirty, don't run anything
-if subprocess.call(("git", "diff-index", "--quiet", "HEAD")):
-    print("Repository is dirty, please commit first")
+if subprocess.call(('git', 'diff-index', '--quiet', 'HEAD')):
+    print('Repository is dirty, please commit first')
     sys.exit(1)
@@
 # Get git hash if any
-revision = subprocess.check_output(("git", "rev-parse", "HEAD"))
+revision = subprocess.check_output(('git', 'rev-parse', 'HEAD'))
 
 # Unit test
-random.seed(1)
-assert walk() == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+random.seed(42)
+assert generate_walk() == [1, 0, -1, -2, -1, 0, 1, 0, -1, -2]
 
@@
 random.seed(seed)
-path = walk()
+walk = generate_walk()
 
 # Display & save results
-print(path)
-results = { 'data':      path,
-            'seed':      seed,
+print(walk)
+results = {'data'     : walk,
+           'seed'     : seed,
            'timestamp': str(datetime.datetime.utcnow()),
-            'revision':  revision,
-            'system' :   sys.version }
-with open("results-R3.txt", "w") as fd:
+           'revision' : revision,
+           'system'   : sys.version}
+with open('results-R3.txt', 'w') as fd:
     fd.write(str(results))
@@
 \begin{code}{\textbf{\textsc{Listing 5:}} Re-runnable, repeatable, reproducible, reusable random walk (R$^4$)}
-# Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
+# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
 # Release under the BSD 2-clause license
-# Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
+# Tested with CPython 3.6.2 / macOS 10.12.6 / 64 bits architecture
 import sys, subprocess, datetime, random
 
-def walk(x0=0, step=1, count=10, seed=0):
+def generate_walk(count, x0=0, step=1, seed=0):
     """ Random walk
+        count: number of steps
         x0   : initial position (default 0)
         step : step size (default 1)
-        count: number of steps (default 10)
         seed : seed for the initialization of the random generator (default 0)
     """
-    path, x = [], x0
     random.seed(seed)
+    x = x0
+    walk = []
     for i in range(count):
-        if random.uniform(-1,+1) > 0:
-            x = x + step
+        if random.uniform(-1, +1) > 0:
+            x += 1
         else:
-            x = x - step
-        path.append(x)
-    return path
+            x -= 1
+        walk.append(x)
+    return walk
 
-if __name__ == '__main__':
+def generate_results(count, x0=0, step=1, seed=0):
+    """Compute a walk and return it alongside its context"""
     # If repository is dirty, don't do anything
-    if subprocess.call(("git", "diff-index", "--quiet", "HEAD")):
-        print("Repository is dirty, please commit first")
+    if subprocess.call(('git', 'diff-index', '--quiet', 'HEAD')):
+        print('Repository is dirty, please commit first')
         sys.exit(1)
@@
     # Get git hash if any
-    revision = subprocess.check_output(("git", "rev-parse", "HEAD"))
+    revision = subprocess.check_output(('git', 'rev-parse', 'HEAD'))
 
-    # Unit test checking reproducibility
-    assert walk(0, 1, 10, 1) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+    # Compute results
+    walk = generate_walk(count=count, x0=x0, step=step, seed=seed)
+    return {'data'      : walk,
+            'parameters': {'x0':x0, 'step':step, 'count':count, 'seed':seed},
+            'timestamp' : str(datetime.datetime.utcnow()),
+            'revision'  : revision,
+            'system'    : sys.version}
+
+if __name__ == '__main__':
+    # Unit test checking reproducibility (will fail with Python<=3.2)
+    assert generate_walk(10, 0, 1, 42) == [1, 0, -1, -2, -1, 0, 1, 0, -1, -2]
 
     # Simulation parameters
-    parameters = { 'x0':    0,
-                   'step':  1,
-                   'count': 10,
-                   'seed' : 1 }
-    path = walk(**parameters)
-    results = {'data':       path,
-               'parameters': parameters,
-               'timestamp':  str(datetime.datetime.utcnow()),
-               'revision':   revision,
-               'system':     sys.version}
+    count, x0, seed = 10, 0, 1
+    results = generate_results(count, x0=x0, seed=seed)
 
     # Save & display results
-    with open("results-R4.txt", "w") as fd:
+    with open('results-R4.txt', 'w') as fd:
         fd.write(str(results))
-
-    print(path)
+    print(results['data'])
 \end{code}
@@
 # Release under the BSD 2-clause license
-# Tested with Python 3.6.1 / Numpy 1.12.0 / macOS 10.12.4 / 64 bits architecture
+# Tested with CPython 3.6.2 / NumPy 1.12.0 / macOS 10.12.6 / 64 bits architecture
 import random
@@
     # Unit test
-    assert walk(n=10, seed=1) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+    assert walk(n=10, seed=42) == [1, 0, -1, -2, -1, 0, 1, 0, -1, -2]
 
@@
     print(path)
-
 \end{code}

### 2017-08-24 codes/r0.py

@@ -1,7 +1,9 @@
+# Random walk (R0: worst you can do)
 import random
-
 x = 0
 for i in xrange(10):
-    step = random.choice([-1,+1])
-    x += step
+    if random.uniform(-1, +1) > 0:
+        x += 1
+    else:
+        x -= 1
     print x,

### 2017-08-24 codes/r1.py

@@ -1,10 +1,14 @@
+# Random walk (R1: re-runnable)
 # Tested with Python 3
 import random
 
-walk, total = [], 0
+x =  0
+walk = []
 for i in range(10):
-    step = random.choice([-1,+1])
-    total += step
-    walk.append(total)
+    if random.uniform(-1, +1) > 0:
+        x += 1
+    else:
+        x -= 1
+    walk.append(x)
 
 print(walk)

### 2017-08-24 codes/r2.py

@@ -1,14 +1,19 @@
+# Random walk (R2: repeatable)
 # Tested with Python 3
 import random
 
 random.seed(0) # RNG initialization
 
-walk, total = [], 0
+x =  0
+walk = []
 for i in range(10):
-    step = random.choice([-1,+1])
-    total += step
-    walk.append(total)
+    if random.uniform(-1, +1) > 0:
+        x += 1
+    else:
+        x -= 1
+    walk.append(x)
 
 print(walk)
-with open("results-R2.txt", "w") as file: # Saving output to disk
-    file.write(str(walk))
+# Saving output to disk
+with open('results-R2.txt', 'w') as fd:
+    fd.write(str(walk))

### 2017-08-24 codes/r3.py

@@ -1,46 +1,43 @@
+# Random walk (R3: reproducible)
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
 # Release under the BSD 2-clause license
-# Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
-import platform, datetime
-import random
-import pip
-import git
+# Tested with CPython 3.6.2 / macOS 10.12.6 / 64 bits architecture
+import sys, subprocess, datetime, random
 
-def git_info(): # Retrieve the code version
-    repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
-    return {'hash': repo.head.object.hexsha, 'dirty': repo.is_dirty(),
-            'git_version': str(pip.git.Git().get_git_version())}
-
-def provenance(): # Get info about the execution environment
-    return {'python'   : {'implementation': platform.python_implementation(),
-                          'version'       : platform.python_version_tuple(),
-                          'compiler'      : platform.python_compiler(),
-                          'branch'        : platform.python_branch(),
-                          'revision'      : platform.python_revision()},
-            'platform' : platform.platform(),
-            'packages'  : list(pip.commands.freeze.freeze()), # list of installed packages
-            'git_info' : git_info(),
-            'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC
-
-def walk():
-    promenade, total = [], 0
+def generate_walk():
+    walk = []
+    x = 0
     for i in range(10):
-        step = +1 if random.uniform(-1,+1) > 0 else -1
-        total += step
-        promenade.append(total)
-    return promenade
+        if random.uniform(-1, +1) > 0:
+            x += 1
+        else:
+            x -= 1
+        walk.append(x)
+    return walk
+
+# If repository is dirty, don't run anything
+if subprocess.call(('git', 'diff-index', '--quiet', 'HEAD')):
+    print('Repository is dirty, please commit first')
+    sys.exit(1)
+
+# Get git hash if any
+revision = subprocess.check_output(('git', 'rev-parse', 'HEAD'))
 
 # Unit test
-random.seed(1)
-assert walk() == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+random.seed(42)
+assert generate_walk() == [1, 0, -1, -2, -1, 0, 1, 0, -1, -2]
 
 # Random walk for 10 steps
 seed = 1
[…]

### 2017-08-24 codes/r4.py

@@ -1,50 +1,54 @@
-# Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
+# Random walk (R4: re-usable)
+# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
 # Release under the BSD 2-clause license
-# Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
-import platform, datetime, json
-import random
-import pip
-import git
+# Tested with CPython 3.6.2 / macOS 10.12.6 / 64 bits architecture
+import sys, subprocess, datetime, random
 
-def git_info():
-    """Retrieve the current git hash, and if the repository has uncommited changes."""
-    repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
-    return {'hash': repo.head.object.hexsha, 'dirty': repo.is_dirty(),
-            'git_version': str(pip.git.Git().get_git_version())}
-
-def provenance():
-    """Get info about the execution environment and the code."""
-    return {'python'   : {'implementation': platform.python_implementation(),
-                          'version'       : platform.python_version_tuple(),
-                          'compiler'      : platform.python_compiler(),
-                          'branch'        : platform.python_branch(),
-                          'revision'      : platform.python_revision()},
-            'platform' : platform.platform(),
-            'packages'  : list(pip.commands.freeze.freeze()), # list of installed packages
-            'git_info' : git_info(),
-            'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC
-
-def walk(n, seed):
-    """Return a `n` steps random walk.
-    `seed` is used to initialize the Mersene Twister random number generator."""
+def generate_walk(count, x0=0, step=1, seed=0):
+    """ Random walk
+        count: number of steps
+        x0   : initial position (default 0)
+        step : step size (default 1)
+        seed : seed for the initialization of the random generator (default 0)
+    """
     random.seed(seed)
-    rndwalk, total = [], 0
-    for i in range(10):
-        step = +1 if random.uniform(-1,+1) > 0 else -1
-        total += step
-        rndwalk.append(total)
-    return rndwalk
+    x = x0
+    walk = []
+    for i in range(count):
+        if random.uniform(-1, +1) > 0:
+            x += 1
+        else:
+            x -= 1
[…]

### 2017-08-24 codes/r5.py

@@ -1,17 +1,14 @@
+# Random walk (R5: replicable)
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
 # Release under the BSD 2-clause license
-# Tested with Python 3.6.1 / Numpy 1.12.0 / macOS 10.12.4 / 64 bits architecture
+# Tested with CPython 3.6.2 / NumPy 1.12.0 / macOS 10.12.6 / 64 bits architecture
 import random
 import numpy as np
 
-def walk(rng, n):
-    """Random walk for n steps."""
-    # Note: We consider 0 to be a positive step
-    steps = 2*(rng.uniform(-1,+1,n) > 0) - 1
-    return steps.cumsum().tolist()
-
-def rng(seed):
-    """Return a random number generator initialized with seed."""
+def _rng(seed):
+    """ Return a numpy random number generator initialized with seed
+        as it would be with python random generator.
+    """
     rng = random.Random()
     rng.seed(seed)
     _, keys, _ = rng.getstate()
@@ -20,15 +17,23 @@ def rng(seed):
     rng.set_state((state[0], keys[:-1], state[2], state[3], state[4]))
     return rng
 
+def walk(n, seed):
+    """ Random walk for n steps """
+
+    rng = _rng(seed)
+    steps = 2*(rng.uniform(-1,+1,n) > 0) - 1
+    return steps.cumsum().tolist()
+
 if __name__ == '__main__':
     # Unit test
-    assert walk(rng(seed=1), 10) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]
+    assert walk(n=10, seed=42) == [1, 0, -1, -2, -1, 0, 1, 0, -1, -2]
 
-    # Random walk for 10 steps
+    # Random walk for 10 steps, initialization with seed=1
     seed = 1
-    x = walk(rng(seed=seed), 10)
+    path = walk(n=10, seed=seed)
 
-    # Display & save results
-    print(x)
-    with open("results-R5-%d.txt" % seed, "w") as file:
-        file.write(str(x))
+    # Save & display results
+    results = {'data': path, 'seed': seed}
+    with open("results-R5.txt", "w") as fd:
+        fd.write(str(results))
+    print(path)

### 2017-08-24 draft.tex
many little improvements, corrections.

@@
 that computer science would mostly be shielded from such issues, since a
-computer program intrinsically describes precisely what it does and is easily
+computer program describes precisely what it does and is easily
 disseminated to other researchers without alteration.
@@
 \citep{Sandve:2013,Schwab:2000}. Computer science is plagued by replication
-problems, in part, precisely because it seems impervious to them. A program can
+problems, in part, because it seems impervious to them. A program can
 fail as a scientific contribution in many different ways for many different
@@
 \begin{code}{Output}
--1 0 -1 0 -1 0 -1 0 1 2 # with the steps being -1, +1, -1, +1, -1, +1, -1, +1, +1, +1
+-1, 0, -1, 0, -1, 0, -1, 0, 1, 2 # with the steps being -1, +1, -1, +1, -1, +1, -1, +1, +1, +1
 \end{code}
@@
 Have you ever tried to re-run a program you wrote some years ago?
-Frustratingly, it can often be surprisingly hard.
+It can often be frustratingly hard.
 Part of the problem is that technology is evolving at a fast pace and you cannot know in advance how the system, the software and the libraries your program depends on will evolve.
@@
 To be clear, it is impossible to write future-proof code, and the best efforts can be stymied by the smallest change in one of the dependencies.
-At the same time, modernizing an unmaintained ten-year-old code can reveal itself to be an arduous and expensive undertaking, and precarious as that, since each change risks affecting the semantics of the program.
-Rather trying to predict the future or painstakingly dusting off old code, an often more straightforward and less risky solution is to recreate the old execution environment\footnote{To be clear, and although virtual machines are often a great help here, this is not always possible.
-It is, however, \emph{always} more difficult when the original execution environment is not documented.}.
+At the same time, modernizing an unmaintained ten-year-old code can reveal itself to be an arduous and expensive undertaking---and precarious, since each change risks affecting the semantics of the program.
+Rather than trying to predict the future or painstakingly dusting off old code, an often more straightforward solution is to recreate the old execution environment\footnote{To be clear, and although virtual machines are often a great help here, this is not always possible.
+It is, however, \emph{always} more difficult when the original execution environment is unknown.}.
 For this to happen however, the dependencies in terms of systems, software and libraries must be made clear enough.\\
@@
 It is important to notice that the re-runnability of a code is not an intrinsic property.
-Rather, it depends on the context, and re-runnability becomes increasingly difficult as the code ages.
-Therefore, to be re-runnable in other researchers' computers and through time, a re-runnable code should describe---with enough details to be recreated---an execution environment in which it is executable.
+Rather, it depends on the context, and becomes increasingly difficult as the code ages.
+Therefore, to be and remain re-runnable in other researchers' computers, a re-runnable code should describe---with enough details to be recreated---an execution environment in which it is executable.
 As shown by \citep{Collberg:2016}, this is far from being either obvious or easy.\\
@@
 Such repeatability is useful.
-If a run of the program produces a particularly puzzling result, it allows you to scrutinize any step of the execution of the program by re-running it again with extraneous prints, or inside a debugger.
+If a run of the program produces a particularly puzzling result, repeatability allows you to scrutinize any step of the execution of the program by re-running it again with extraneous prints, or inside a debugger.
 It is also the easiest way to prove that the program did indeed produce the published results.
@@
 
-Setting seeds is should be done carefully.
+Setting seeds should be done carefully.
 Using 439 as a seed in the previous program would result in ten consecutive +1 steps\footnote{With CPython 3.3-3.6.
-See the next section for details.}, which---although a perfectly valid random walk---lend itself to a gross misinterpretation of the overall dynamic of the algorithm.
+See the next section for details.}, which---although a perfectly valid random walk---lend itself to a gross misinterpretation of the overall dynamics of the algorithm.
 Verifying that the qualitative aspects of the results and the conclusions that are made are not tied to a specific initialization of the pseudo-random generator is an integral part of any scientific undertaking in computational science;
@@
 As explained by \citeauthor{Donoho:2009} \parencite{Donoho:2009}, scientific practice must expect that {\em errors are ubiquitous}, and therefore be robust to them.
-Ensuring reproducibility is an fundamental step toward this: it provides other researchers the means to verify that your code does indeed produce the published results, and to scrutinize the procedures it used to produce them.
+Ensuring reproducibility is a fundamental step toward this: it provides other researchers the means to verify that the code does indeed produce the published results, and to scrutinize the procedures it used to produce them.
 As demonstrated by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility is hard.\\
@@
 The cause is to be found in a change that occurred in the pseudo-random number generator between Python 3.2 and Python 3.3.
-Executed with Python 2.7 to 3.2, the code will produce the sequence 1, 2, 1, 2, 3, 4, 5, 6, 7, 6.
-But with Python 3.3 to 3.6, it will produce 1, 2, 1, 0, 1, 0, 1, 0, -1, 0.
-With future version of the language, it may change still.\\
+Executed with Python 2.7 to 3.2, the code will produce the sequence -1, 0, 1, 0, -1, -2, -1, 0, -1, -2.
+But with Python 3.3 to 3.6, it will produce -1, -2, -1, -2, -1, 0, 1, 2, 1, 0.
+With future versions of the language, it may change still.\\
 
@@
 % if code changes, parameters are lost
-Recording the execution environment is not enough though.
+Recording the execution environment is only the first step.
 The R$^2$ program uses a random seed but does not keep a trace of it except in the code.
 Should the code change after the production of the results, someone provided with the last version of the code will not be able to know which seed was used to produce the results, and would need to iterate through all possible random seeds, an impossible task in practice\footnote{Here, with $2^{10}$ possibilities for a 10-step random walk, a seed matching the results could obviously be found.
-For instance, seed 1123581321 matches the results of R$^0$ with Python 2.7.
+For instance, seed 11235813 matches the results of R$^0$ with Python 2.7.
 This is not possible for a 100-step walk.}.\\
@@
 % 1. parameters and 2. code version with results
-This is why results files should come alongside their context, i.e.
+This is why result files should come alongside their context, i.e.
 an exhaustive list of the parameters used as well as a precise description of the execution environment, as the R$^3$ code does.
-The code itself is part of that context: the version of the code must be recorded as part of the provenance data.
+The code itself is part of that context: the version of the code must be recorded.
 It is common for different results or different figures to have been generated by different versions of the code.
@@
 This also allows, if some errors are found and corrected after some results have been obtained, to identify which ones should be recomputed.
-In R$^3$, the code records the last git commit hash, and whether the repository holds uncommitted changes when the computation starts.\\
+In R$^3$, the code records the git revision, and whether the repository holds uncommitted changes when the computation starts.\\
 
 % scripts to generates figures & data behind graphs
-Published results should come from version of the code where every file has been committed.
+Published results should obviously come from version of the code where every change and every file has been committed.
 This includes pre-processing, post-processing and plotting code.
 Plotting code may seem mundane, but it is as vulnerable as any other piece of the code to bugs and errors.
-With a code that is re-runnable, repeatable and contains plotting scripts, a user can quickly reproduce published figures.
+%With a code that is re-runnable, repeatable and contains plotting scripts, a user can quickly reproduce published figures.
 When it comes to checking that the reproduced data match the one published in the article, however, figures can reveal themselves to be imprecise and cumbersome, and sometimes plain unusable.
-Ideally, published figures should be accompanied by their underlying data (coordinates of the plotted points) in the supplementary data to allow straightforward numeric comparisons.\\
+To avoid having to manually overlay pixelated plots, published figures should be accompanied by their underlying data (coordinates of the plotted points) in the supplementary data to allow straightforward numeric comparisons.\\
 
@@
 % reproducibility implies availability
-It should be obvious by now that \emph{reproducibility implies availability}.
+It is obvious that \emph{reproducibility implies availability}.
 As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request.
 While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable.
-Code \emph{and} input data \emph{and} result data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as Figshare or Zenodo\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment.
-Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.}.\\
+Code \emph{and} input data \emph{and} result data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as \href{https://figshare.com}{Figshare} or \href{https://zenodo.org}{Zenodo}\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment.
+Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.}.
+The codes presented in this article are available in the GitHub repository \href{https://github.com/rougier/random-walk}{github.com/rougier/random-walk} and at \href{https://doi.org/10.5281/zenodo.848221}{doi.org/10.5281/zenodo.848221}.\\
 
@@
 Here, comments and documentation can make a significant difference.
-Source code reflect the results of the decisions that were made during its creation, but not the reasons behind those decisions.
+Source code reflects the results of the decisions that were made during its creation, but not the reasons behind those decisions.
 In science, where the method and its justification matter as much as the results, those reasons are precious knowledge.
@@
 Scientists are rarely trained in software engineering, and reusability can represent an expensive endeavour if undertaken as an afterthought, for little tangible short-term benefits, for a codebase that might, after all, see only a single use.
-And, in fact, reusability is not as indispensable a requirement as the other ones presented here.
+And, in fact, reusability is not as indispensable a requirement as re-runnability, repeatability and reproducibility.
 Yet, some simple measures can tremendously increase reusability, and at the same time strengthen reproducibility and re-runnability over the long-term.\\
@@
 Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user.
-This improves reproducibility in two ways: it allows those conditions to be explicitly tracked in the result files since they contain the parameters used, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
+This improves reproducibility in two ways: it allows those conditions to be recorded as parameters in the result files, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
 With a documentation explaining which script or configuration file corresponds to which experiment, reproducing the different figures becomes straightforward.\\
@@
 Documentation is one of the most potent tools for reusability.
-A proper documentation on how to install and run the software often makes the difference whether other researchers manage to use the code or not.
+A proper documentation on how to install and run the software often makes the difference whether other researchers manage to use it or not.
 A comment describing what each function does, however evident, can avoid hours of head-scratching.
@@
 Scientists, however, are not always brilliant developers.
-Code quality, in science, averages on mediocre.
 Of course, bad, complicated code should be rewritten until is simple enough to explain itself.
 But realistically, this is not always going to be done: there is simply not enough incentive for it.
-There, a comment that explains the intention and scientific reasons behind a block of code can tremendously useful.\\
+There, a comment that explains the intentions and reasons behind a block of code can tremendously useful.\\
 
@@
 Having made a software reusable offers an additional way to find errors, especially if your scientific contribution is popular.
-Unfortunately, this is not always effective, and some recent cases have dramatically illustrated the tremendous impact a bug can have in science
+Unfortunately, this is not always effective, and some recent cases have dramatically illustrated the considerable impact a bug can have in science
 \citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}. This is
-why, as explained by Peng et al. \cite{Peng:2006}, {\em the replication of
+why, as explained by Peng et al. (\cite{Peng:2006}), {\em the replication of
   important findings by multiple independent investigators is fundamental to
@@
 \begin{quotation}
-{\em The model uses the Mersene Twister generator initialized with the seed 0.
+{\em The model uses the Mersene Twister generator initialized with the seed 1.
 At each iteration, a uniform number between -1 (included) and +1 (excluded) is drawn and the sign of the result is used for generating a positive or negative step.}
@@
 Here we are able to replicate exactly\footnote{Striving, as we do here, for a perfect quantitative match may seem unnecessary.
-Yet, in replication projects, quantitative comparisons are an effective way to verify that the behavior has been reproduced.
+Yet, in replication projects, quantitative comparisons are a simple and effective way to verify that the behavior has been reproduced.
 Moreover, they are particularly helpful to track exactly where the code of a tentative replication fails to reproduce the published results.} the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details (the use of the {\tt random} module of the standard Python library of CPython 3.6.1), or to the code itself.\\
@@
 Is the initialization value included (it was not in our codes so far)?
-This slight difference affects the statistics of short runs.\\
+This slight difference might affect the statistics of short runs.\\
 
@@
 They are mostly inconspicuous to the reader and oftentimes, to the writer as well.
-In fact,  there is no better way to ferret out the ambiguities, big and small, of an article than to replicate it.
+In fact, the best way to ferret out the ambiguities, big and small, of an article is to replicate it.
 This is one of the reasons why the ReScience journal \citep{Rougier:2017} has been created.
@@
 
-Throughout the evolution of a small random walk example implemented in Python, we illustrated some of the issue that may plague code that work and produce correct results, and reduce its value as an element of scientific knowledge.
-Further, we articulated characteristics that a code should possess to be a useful part of a scientific publication: it should be rerunnable, repeatable, reproducible, reusable and replicable.\\
+Throughout the evolution of a small random walk example implemented in Python, we illustrated some of the issues that may plague code that works and produces correct results, and reduce its value as an element of scientific knowledge.
+Further, we articulated five characteristics that a code should possess to be a useful part of a scientific publication: it should be rerunnable, repeatable, reproducible, reusable and replicable.\\
 
@@
 But recreating the old code’s execution environment may be: to ensure that the long-term rerunnability of a code, its execution environment must be documented.
-For our example, a single comment went a long way to go from the R$^0$ to the R$^1$ (rerunnable) code.\\
+For our example, a single comment went a long way to transform the R$^0$ code into the R$^1$ (rerunnable) one.\\
 
@@
 Even more care is needed to make a code reproducible.
-The exact execution environment, code and parameters used must be recorded and embedded in the results files.
-The R$^3$ (reproducible) version implements provenance methods to that end.
+The exact execution environment, code and parameters used must be recorded and embedded in the results files, as the R$^3$ (reproducible) version does.
 Furthermore, the code must be made available as supplementary data with the whole computational workflow, from preprocessing steps to plotting scripts.\\
@@
 Taken into account during development rather than as an afterthought, simple measures can avoid hours of head-scratching for others, and for yourself—in a few years.
-Documentation is paramount here, even a single comment per function, as was done in the R$^4$ (reusable) version.\\
+Documentation is paramount here, even if it is a single comment per function, as was done in the R$^4$ (reusable) version.\\
 
@@
 But making sure that these solutions are adopted will not be solved by articles such as this one.
-Just like in psychology and biology, we have to modify the incentive of the researchers, by adopting exigences, enforced domain-wide, on what constitutes an acceptable scientific computational work.
+Just like in other fields, we have to modify the incentives for the researchers to publish by adopting exigences, enforced domain-wide, on what constitutes an acceptable scientific computational work.
 

### 2017-08-24 codes/r0.py

@@ -2,8 +2,6 @@
 import random
 x = 0
 for i in xrange(10):
-    if random.uniform(-1, +1) > 0:
-        x += 1
-    else:
-        x -= 1
+    step = random.choice([-1,+1])
+    x += step
     print x,

### 2017-08-24 codes/r2.py

@@ -7,10 +7,8 @@ random.seed(0) # RNG initialization
 x =  0
 walk = []
 for i in range(10):
-    if random.uniform(-1, +1) > 0:
-        x += 1
-    else:
-        x -= 1
+    step = random.choice([-1,+1])
+    x += step
     walk.append(x)
 
 print(walk)

### 2017-08-24 draft.tex
conceptual DOI for random-walk rather than the version-specific one.

@@
 Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.}.
-The codes presented in this article are available in the GitHub repository \href{https://github.com/rougier/random-walk}{github.com/rougier/random-walk} and at \href{https://doi.org/10.5281/zenodo.848221}{doi.org/10.5281/zenodo.848221}.\\
+The codes presented in this article are available in the GitHub repository \href{https://github.com/rougier/random-walk}{github.com/rougier/random-walk} and at \href{https://doi.org/10.5281/zenodo.848217}{doi.org/10.5281/zenodo.848217}.\\
 

### 2017-08-24 draft.tex
update to latest random-walk codes

@@
 for i in xrange(10):
-    if random.uniform(-1, +1) > 0:
-        x += 1
-    else:
-        x -= 1
+    step = random.choice([-1,+1])
+    x += step
     print x,
@@
 for i in range(10):
-    if random.uniform(-1, +1) > 0:
-        x += 1
-    else:
-        x -= 1
+    step = random.choice([-1,+1])
+    x += step
     walk.append(x)
@@
 
-random.seed(0) # RNG initialization
+random.seed(1) # RNG initialization
 
@@
 def generate_walk():
+    x = 0
     walk = []
-    x = 0
     for i in range(10):

### 2017-08-24 draft.tex
finishing touches

@@
 re-obtained by an independent party, it merely becomes, at best, an observation
-that may inspire future research (\cite{Mesirov:2010,osc:2015}). Replication
-issues have received increased attention in recent years, with a particular focus on medicine and psychology (\cite{Iqbal:2016}).  One could think
+that may inspire future research \citep{Mesirov:2010,osc:2015}. Replication
+issues have received increased attention in recent years, with a particular focus on medicine and psychology \citep{Iqbal:2016}.  One could think
 that computer science would mostly be shielded from such issues, since a
@@
 reusable (R$^4$) and replicable (R$^5$). Let us illustrate this with a small
-example, a random walk written in Python:\\
+example, a random walk \citep{Hughes:1995} written in Python:\\
 
@@
 Rather, it depends on the context, and becomes increasingly difficult as the code ages.
-Therefore, to be and remain re-runnable in other researchers' computers, a re-runnable code should describe---with enough details to be recreated---an execution environment in which it is executable.
+Therefore, to be and remain re-runnable on other researchers' computers, a re-runnable code should describe---with enough details to be recreated---an execution environment in which it is executable.
 As shown by \citep{Collberg:2016}, this is far from being either obvious or easy.\\
@@
 In other words, the next step is to make your program deterministic, producing {\em repeatable} output.
-Such repeatability is useful.
+Repeatability is valuable.
 If a run of the program produces a particularly puzzling result, repeatability allows you to scrutinize any step of the execution of the program by re-running it again with extraneous prints, or inside a debugger.
-It is also the easiest way to prove that the program did indeed produce the published results.
-Such repeatability is not always possible or easy \citep{Diethelm:2012, Courtes:2015}.
+Repeatability is also useful to prove that the program did indeed produce the published results.
+Repeatability is not always possible or easy \citep{Diethelm:2012, Courtes:2015}.
 But for sequential programs not depending on analog inputs, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\
@@
 But with Python 3.3 to 3.6, it will produce -1, -2, -1, -2, -1, 0, 1, 2, 1, 0.
-With future versions of the language, it may change still.\\
+With future versions of the language, it may change still.
+For the R$^3$ version, we abandon the use of the {\tt \href{https://docs.python.org/3.6/library/random.html#random.choice}{random.choice}} function in favor of the {\tt \href{https://docs.python.org/3.6/library/random.html#random.uniform}{random.uniform}} function, whose behavior is consistent across versions 2.7 to 3.6 of Python.\\
 
@@
 The exact execution environment used to produce the results must also be specified---rather than the broadest set of environments where the code can be effectively run.
-In other words, assertions such as "the results were obtained with CPython 3.6.1" are more valuable, in a scientific context, than "the program works with Python 3.x and above".
+In other words, assertions such as \enquote{the results were obtained with CPython 3.6.1} are more valuable, in a scientific context, than \enquote{the program works with Python 3.x and above}.
 With the increasing complexity of computational stacks, retrieving and deciding what is pertinent (CPU architecture?
@@
 
-% However, making your program reusable requires a little work.
-If it is natural, when one writes a program for oneself, to make implicit choices that suit one needs, such implicit choices won't resist the will of a user community.
-For example, having magic numbers or ill-defined constants in your program might be cumbersome for others.
-Similarly, it might be acceptable to not have a proper documentation when you're the only user (actually it might not be acceptable because the you now and the you in six months are two different persons), but other users want to know precisely how to use this or that function and what it does exactly.
-%FIXME: reword
-%
-% If we look back at our random walker, we can list all the implicit choices that have been made:
-% \begin{itemize}
-% \item Step number has been hard-coded (magic number 10)
-% \item Step size has been fixed to 1
-% \item Initial position has been set to 0
-% \item No possibility of using a user-specified random seed
-% \item Only one run can be started at once
-% \item Walker walk along a single dimension
-% \item ...
-% \end{itemize}
-% Here, we won't try to implement all these features at once, but only giving the possibility for a user to implement them if needed.\\
+Similarly, code behavior should not be changed by commenting/uncommenting code \citep{Wilson:2017}.
+Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user.
+This improves reproducibility in two ways: it allows those conditions to be recorded as parameters in the result files, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
+With a documentation explaining which script or configuration file corresponds to which experiment, reproducing the different figures becomes straightforward.\\
 
@@
 
-Similarly, code behavior should not be changed by commenting/uncommenting code \citep{Wilson:2017}.
-Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user.
-This improves reproducibility in two ways: it allows those conditions to be recorded as parameters in the result files, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
-With a documentation explaining which script or configuration file corresponds to which experiment, reproducing the different figures becomes straightforward.\\
-
 Documentation is one of the most potent tools for reusability.
@@
 
-Reusability is not a strict requirement for scientific code. But it has many benefits, and a few simple measures can foster it considerably.
-
-% But a few good practices, observed during the development of the code can have significant benefits for you, your team and any other reasearche desirous to use your code.
+Reusability is not a strict requirement for scientific code.
+But it has many benefits, and a few simple measures can foster it considerably.
+To complement the R$^4$ version provided here, we provide an example repository of a re-runnable, repeatable, reproducible and reusable random walk code.
+The repository is available on GitHub \href{https://github.com/benureau/r5}{github.com/benureau/r5} and here \href{https://doi.org/10.5281/zenodo.848284}{doi.org/10.5281/zenodo.848284}.
 
@@
 \citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}. This is
-why, as explained by Peng et al. (\cite{Peng:2006}), {\em the replication of
+why, as explained by Peng et al. \citep{Peng:2006}, {\em the replication of
   important findings by multiple independent investigators is fundamental to
@@
 Minute changes in the numerical estimation of a common first-order differential equation can have significant impact \citep{Crook:2013}.
-A reproducible code plays a role that cannot be adequately fulfilled by the conceptual description of it found in the article.
-Hence, replicability does not negate the necessity of reproducibility.
+Hence, a reproducible code plays an important role alongside its article: it is a objective catalog of all the implementation details.\\
+
+A researcher seeking to replicate published results might first consider only the article.
+If she fails to replicate the results, she will consult the original code, and be able to pinpoint why her code and the code of the authors differ in behavior.
+Because a mistake on their part?
+Hers?
+Or a difference in seemingly innocuous implementation details?
+A fine analysis of why a particular algorithmic description is lacking or ambiguous or why a minor implementation decision is in fact crucial to obtain the published results is of great  scientific value.
+Such an analysis can only be done with access to both the article and the code.
+With only the article, the researcher will often be unable to understand why she failed to replicate the results, and will naturally be inclined to only report replication successes.\\
+
+Replicability, therefore, does not negate the necessity of reproducibility.
+In fact, it often relies on it.
 To illustrate this, let us consider what could be the textual description of the random walker (as it would be written in an article describing it):
@@
 \href{https://docs.python.org/3.6/library/random.html}{Python documentation},
-states that "Python uses the Mersenne Twister as the core generator. It produces
-53-bit precision floats and has a period of 2**19937-1".
+states that \enquote{Python uses the Mersenne Twister as the core generator. It produces
+53-bit precision floats and has a period of 2**19937-1}.
 Someone trying to replicate the work however might choose to use the RNG from the \href{http://www.numpy.org/}{NumPy library}.
@@
 
+Code is an integral part of any submission to the ReScience journal.
+During the review process, reviewers run the submitted code, may criticise its quality and its ease-of-use, and verify the reproduciblity of the replication.
+The Journal of Open Source Software \citep{Smith:2017} functions similarly: testing the code is a fundamental part of the review process.
+
 % \clearpage
@@
 
-Throughout the evolution of a small random walk example implemented in Python, we illustrated some of the issues that may plague code that works and produces correct results, and reduce its value as an element of scientific knowledge.
-Further, we articulated five characteristics that a code should possess to be a useful part of a scientific publication: it should be rerunnable, repeatable, reproducible, reusable and replicable.\\
+Throughout the evolution of a small random walk example implemented in Python, we illustrated some of the issues that may plague scientific code.
+The code may be correct and of good quality, but still many problems may reduce its contribution to scientific knowledge.
+To make these problems explicit, we articulated five characteristics that a code should possess to be a useful part of a scientific publication: it should be re-runnable, repeatable, reproducible, reusable and replicable.\\
 
 Running old code on tomorrow’s computer and software stacks may not be possible.
-But recreating the old code’s execution environment may be: to ensure that the long-term rerunnability of a code, its execution environment must be documented.
-For our example, a single comment went a long way to transform the R$^0$ code into the R$^1$ (rerunnable) one.\\
+But recreating the old code’s execution environment may be: to ensure that the long-term re-runnability of a code, its execution environment must be documented.
+For our example, a single comment went a long way to transform the R$^0$ code into the R$^1$ (re-runnable) one.\\
 

### 2017-08-25 draft.tex
Typos, more compact layout.

@@
   \textbf{Fabien C. Y. Benureau}$^{1,2,3,*}$,
-  \textbf{Nicolas P. Rougier}$^{1,2,3}$\\ \begin{minipage}{\textwidth}
+  \textbf{Nicolas P. Rougier}$^{1,2,3}$\\ \begin{minipage}{0.8\textwidth}
     \begin{center}
       \vspace{2mm}
+      \small
       $^{1}$INRIA Bordeaux Sud-Ouest Talence, France $^{2}$ Institut des
@@
 But realistically, this is not always going to be done: there is simply not enough incentive for it.
-There, a comment that explains the intentions and reasons behind a block of code can tremendously useful.\\
+There, a comment that explains the intentions and reasons behind a block of code can be tremendously useful.\\
 
@@
 A researcher seeking to replicate published results might first consider only the article.
-If she fails to replicate the results, she will consult the original code, and be able to pinpoint why her code and the code of the authors differ in behavior.
+If she fails to replicate the results, she will consult the original code, and with it be able to pinpoint why her code and the code of the authors differ in behavior.
 Because a mistake on their part?
 Hers?
-Or a difference in seemingly innocuous implementation details?
+Or a difference in a seemingly innocuous implementation detail?
 A fine analysis of why a particular algorithmic description is lacking or ambiguous or why a minor implementation decision is in fact crucial to obtain the published results is of great scientific value.
@@
 This description, while somewhat precise, forgoes---as it is common---the initialization of the variables (here the starting value of the walk: {\tt 0}), and the technical details about which implementation of the RNG is used.\\
-
-% NumPy and Python: differences in the Mersene Twister
-It may look innocuous. After all, the
-\href{https://docs.python.org/3.6/library/random.html}{Python documentation},
-states that \enquote{Python uses the Mersenne Twister as the core generator. It produces
-53-bit precision floats and has a period of 2**19937-1}.
-Someone trying to replicate the work however might choose to use the RNG from the \href{http://www.numpy.org/}{NumPy library}.
-The NumPy library is extensively used in the science community, and it provides an implementation of the Mersene Twister generator too.
-Unfortunately, the way the seed is interpreted by the two implementations is different, yielding different random sequences.\\
-
-Here we are able to replicate exactly\footnote{Striving, as we do here, for a perfect quantitative match may seem unnecessary.
-Yet, in replication projects, quantitative comparisons are a simple and effective way to verify that the behavior has been reproduced.
-Moreover, they are particularly helpful to track exactly where the code of a tentative replication fails to reproduce the published results.} the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details (the use of the {\tt random} module of the standard Python library of CPython 3.6.1), or to the code itself.\\
 
@@
 
+% NumPy and Python: differences in the Mersene Twister
+It may look innocuous. After all, the
+\href{https://docs.python.org/3.6/library/random.html}{Python documentation},
+states that \enquote{Python uses the Mersenne Twister as the core generator. It produces
+53-bit precision floats and has a period of 2**19937-1}.
+Someone trying to replicate the work however might choose to use the RNG from the \href{http://www.numpy.org/}{NumPy library}.
+The NumPy library is extensively used in the science community, and it provides an implementation of the Mersene Twister generator too.
+Unfortunately, the way the seed is interpreted by the two implementations is different, yielding different random sequences.\\
+
+Here we are able to replicate exactly\footnote{Striving, as we do here, for a perfect quantitative match may seem unnecessary.
+Yet, in replication projects, quantitative comparisons are a simple and effective way to verify that the behavior has been reproduced.
+Moreover, they are particularly helpful to track exactly where the code of a tentative replication fails to reproduce the published results.} the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details (the use of the {\tt random} module of the standard Python library of CPython 3.6.1), or to the code itself.\\
+
 But there are still more subtle problems with the description given above.
@@
 Taken into account during development rather than as an afterthought, simple measures can avoid hours of head-scratching for others, and for yourself—in a few years.
-Documentation is paramount here, even if it is a single comment per function, as was done in the R$^4$ (reusable) version.\\
+Documentation is paramount here, even if it is a single comment per function, as it was done in the R$^4$ (reusable) version.\\
 
-Finally, there is the belief that an article should suffice itself: the description of the algorithms present in the paper should suffice to re-obtain (to replicate) the published results.
-For well-written papers that precisely dissociate conceptually significant details from irrelevant implementation details, that may be.
-But science should not assume the best of cases.
-Science assumes that errors are ubiquitous.
+Finally, there is the belief that an article should suffice by itself: the descriptions of the algorithms present in the paper should suffice to re-obtain (to replicate) the published results.
+For well-written papers that precisely dissociate conceptually significant aspects from irrelevant implementation details, that may be.
+But scientific practice should not assume the best of cases.
+Science assumes that errors can crop up everywhere.
 Every paper is a mistake or a forgotten parameter away from irreproducibility.

### 2017-08-26 draft.tex
Add OrcIDs

@@
   commentstyle=\color{gray},
-  columns=fullflexible,
+  columns=flexible,
 }
@@
   \textbf{Fabien C. Y. Benureau}$^{1,2,3,*}$,
-  \textbf{Nicolas P. Rougier}$^{1,2,3}$\\ \begin{minipage}{0.8\textwidth}
+  \textbf{Nicolas P. Rougier}$^{1,2,3}$\\
+  {\small \href{https://orcid.org/0000-0003-4083-4512}{0000-0003-4083-4512},
+  \href{https://orcid.org/0000-0002-6972-589X}{0000-0002-6972-589X}}\\
+  \begin{minipage}{0.8\textwidth}
     \begin{center}
       \vspace{2mm}
-      \small
-      $^{1}$INRIA Bordeaux Sud-Ouest Talence, France $^{2}$ Institut des
-      Maladies Neurodégénératives, Université de Bordeaux, CNRS UMR 5293,
-      Bordeaux, France $^{3}$ LaBRI, Université de Bordeaux, Bordeaux INP, CNRS
-      UMR 5800, Talence, France\\
+      \scriptsize
+      $^{1}$INRIA Bordeaux Sud-Ouest, Talence, France\\
+      $^{2}$ Institut des Maladies Neurodégénératives, Université de Bordeaux, CNRS UMR 5293, Bordeaux, France\\
+      $^{3}$ LaBRI, Université de Bordeaux, Bordeaux INP, CNRS UMR 5800, Talence, France\\
       \vspace{2mm}
@@
 
+In the code above, the {\tt \href{https://docs.python.org/3.6/library/random.html#random.choice}{random.choice}} function randomly returns either +1 or -1.
+The instruction \enquote{\tt for i in xrange(10):}  executes the next three indented lines ten times.
 Executed, this program would display:

### 2017-10-16 draft.tex
Changes following review

@@
 issues have received increased attention in recent years, with a particular focus on medicine and psychology \citep{Iqbal:2016}.  One could think
-that computer science would mostly be shielded from such issues, since a
+that computational research would mostly be shielded from such issues, since a
 computer program describes precisely what it does and is easily
@@
 working code into meaningful scientific contributions are rarely undertaken
-\citep{Sandve:2013,Schwab:2000}. Computer science is plagued by replication
-problems, in part, because it seems impervious to them. A program can
+\citep{Collberg:2016,Sandve:2013,Schwab:2000}. Computational research is plagued by replication
+problems, in part, because it seems impervious to them.
+Contrary to production software who provide a service geared towards a practical outcome , the motivation behind scientific code is to test an hypothesis.
+While in some instance production software and scientific code are indistinguishable, the reason why they were created is different, and, therefore, so are the criteria to evaluate their success.
+A program can
 fail as a scientific contribution in many different ways for many different
@@
 The R$^2$ program uses a random seed but does not keep a trace of it except in the code.
-Should the code change after the production of the results, someone provided with the last version of the code will not be able to know which seed was used to produce the results, and would need to iterate through all possible random seeds, an impossible task in practice\footnote{Here, with $2^{10}$ possibilities for a 10-step random walk, a seed matching the results could obviously be found.
+Should the code change after the production of the results, someone provided with the last version of the code will not be able to know which seed was used to produce the results, and would need to iterate through all possible random seeds, an impossible task in practice\footnote{Here, with $2^{10}$ possibilities for a 10-step random walk, a seed matching the any possible sequence could certainly be found.
 For instance, seed 11235813 matches the results of R$^0$ with Python 2.7.
-This is not possible for a 100-step walk.}.\\
+Such a search becomes intractable for a 100-step walk.}.\\
 
@@
 This improves reproducibility in two ways: it allows those conditions to be recorded as parameters in the result files, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
-With a documentation explaining which script or configuration file corresponds to which experiment, reproducing the different figures becomes straightforward.\\
+With documentation explaining which script or configuration file corresponds to which experiment, reproducing the different figures becomes straightforward.\\
 
@@
 Having made a software reusable offers an additional way to find errors, especially if your scientific contribution is popular.
-Unfortunately, this is not always effective, and some recent cases have dramatically illustrated the considerable impact a bug can have in science
-\citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}. This is
-why, as explained by Peng et al. \citep{Peng:2006}, {\em the replication of
+Unfortunately, this is not always effective, and some recent cases have shown that bugs can lurk in well-used open-source code, impacting the false positive rates of fMRI studies \citep{Eklund:2016}, or the encryption of communications over the Internet \citep{Durumeric:2014}.
+Let's be clear: the goal here is not to remove all bugs and mistakes from science.
+The goal is to have methods and practices in place that make possible for the inevitable errors that will be made to be caught and corrected by motivated investigators.
+This is why, as explained by Peng et al.
+\citep{Peng:2006}, {\em the replication of
   important findings by multiple independent investigators is fundamental to
@@
 \emph{Replicability} is the implicit
-assumption that any article that does not provide the code source makes: that
+assumption that any article that does not provide the source code makes: that
 the description it provides of the algorithms is sufficiently precise and
-complete to re-obtain the results it presents. While every published article
+complete to re-obtain the results it presents.
+Therefore, replicating a result means coding a new implementation from scratch that match the conceptual description of the article.
+This affords robustness to the results because, should the original code contain an error, a different codebase creates the possibility that this error will not be repeated, in the same way replicating an laboratory experiment can ferret out hidden biases.
+While every published article
 should strive for replicability, it is seldom obtained. In fact, absent an explicit effort to make an algorithmic description replicable, there is little probability that it will be.\\
@@
 In fact, the best way to ferret out the ambiguities, big and small, of an article is to replicate it.
-This is one of the reasons why the ReScience journal \citep{Rougier:2017} has been created.
-This journal {\em targets computational research and encourages the explicit replication of already published research, promoting new and open-source implementations in order to ensure that the original research is reproducible}.\\
+This is one of the reasons why the ReScience journal \citep{Rougier:2017} was created (the second author, Nicolas Rougier, is one of the editor-in-chief of ReScience).
+This journal, run entirely by volunteers on GitHub, {\em targets computational research and encourages the explicit replication of already published research, promoting new and open-source implementations in order to ensure that the original research is reproducible}.\\
 

### 2017-10-17 draft.tex
second part of changes. add links in listings

@@
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\textbf{\textsc{Listing 1:}} Random walk (R$^0$)}
+\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 1:}} Random walk (R$^0$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R0.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
 import random
@@
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\textbf{\textsc{Listing 2:}} Re-runnable random walk (R$^1$)}
+\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 2:}} Re-runnable random walk (R$^1$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R1.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
 # Tested with Python 3
@@
 Repeatability is not always possible or easy \citep{Diethelm:2012, Courtes:2015}.
-But for sequential programs not depending on analog inputs, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\
+But for sequential and deterministically parallels programs \citep{Hines:2008, Collange:2015} not depending on analog inputs, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\
 
@@
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\textbf{\textsc{Listing 3:}} Re-runnable, repeatable random walk (R$^2$)}
+\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 3:}} Re-runnable, repeatable random walk (R$^2$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R2.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
 # Tested with Python 3
@@
 While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable.
-Code \emph{and} input data \emph{and} result data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as \href{https://figshare.com}{Figshare} or \href{https://zenodo.org}{Zenodo}\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment.
-Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.}.
+Code \emph{and} input data \emph{and} result data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as \href{https://figshare.com}{Figshare}, \href{https://zenodo.org}{Zenodo}\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment.
+Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.} or a domain specific database, such as \href{https://senselab.med.yale.edu/modeldb/}{ModelDB} for computational neuroscience.
 The codes presented in this article are available in the GitHub repository \href{https://github.com/rougier/random-walk}{github.com/rougier/random-walk} and at \href{https://doi.org/10.5281/zenodo.848217}{doi.org/10.5281/zenodo.848217}.\\
@@
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\textbf{\textsc{Listing 4:}} Re-runnable, repeatable, reproducible random walk (R$^3$)}
+\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 4:}} Re-runnable, repeatable, reproducible random walk (R$^3$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R3.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
@@
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\textbf{\textsc{Listing 5:}} Re-runnable, repeatable, reproducible, reusable random walk (R$^4$)}
+\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 5:}} Re-runnable, repeatable, reproducible, reusable random walk (R$^4$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R4.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
@@
 \emph{Replicability} is the implicit
-assumption that any article that does not provide the source code makes: that
+assumption that an article that does not provide the source code makes: that
 the description it provides of the algorithms is sufficiently precise and
 complete to re-obtain the results it presents.
-Therefore, replicating a result means coding a new implementation from scratch that match the conceptual description of the article.
-This affords robustness to the results because, should the original code contain an error, a different codebase creates the possibility that this error will not be repeated, in the same way replicating an laboratory experiment can ferret out hidden biases.
-While every published article
-should strive for replicability, it is seldom obtained. In fact, absent an explicit effort to make an algorithmic description replicable, there is little probability that it will be.\\
+Here, replicating implies writing a new code matching the conceptual description of the article, in order to reobtain the same results.
+Replication affords robustness to the results because, should the original code contain an error, a different codebase creates the possibility that this error will not be repeated, in the same way that replicating an laboratory experiment in a different laboratory can ferret out subtle biases.
+While every published article should strive for replicability, it is seldom obtained.
+In fact, absent an explicit effort to make an algorithmic description replicable, there is little probability that it will be.\\
 
 % conceptual description != replicable description
-This is because most papers strive to communicate the main ideas behind their contribution is terms as simple and as clear as possible, so that the reader may be able to easily understand them and the results that are presented to him.
+This is because most papers strive to communicate the main ideas behind their contribution in terms as simple and as clear as possible, so that the reader may be able to easily understand them and the results that are presented.
 Trying to ensure replicability in the main text adds a myriad of esoteric details that are not conceptually significant and clutter the explanations.
@@
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\textbf{\textsc{Listing 6:}} Replicated random walk (R$^5$)}
+\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 6:}} Replicated random walk (R$^5$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R5.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau

### 2017-10-18 draft.tex
final changes...

@@
 \begin{document}
+\vfill
 \maketitle
+
+\begin{abstract}
+  Scientific code is different from production software.
+Scientific code, by producing results that are then analyzed and interpreted, participates in the elaboration of scientific conclusions.
+This imposes specific constraints on the code that are often overlooked in practice.
+We articulate, with a small example, five characteristics that a scientific code in computational science should possess: re-runnable, repeatable, reproducible, reusable and replicable.
+The code should be executable (re-runnable) and produce the same result more than once (repeatable); it should allow an investigator to reobtain the published results (reproducible) while being easy to use, understand and modify (reusable), and it should act as an available reference for any ambiguity in the algorithmic descriptions of the article (replicable).
+\end{abstract}
 % -------------------------
+\vfill
+
+\pagebreak
 \section*{Introduction (R$^{\mathbf 0}$)}
 
-Replicability is a cornerstone of science.  If an experimental result cannot be
+Replicability\footnote{\emph{Reproducibility} and \emph{replicability} are employed differently by different authors and in different domains (see for instance the report from the U.S.
+\cite{Schwalbe:2016}).
+Here, we place ourselves in the context of computational works, where data is produced by a program.
+A result is \emph{reproducible} if one can take the original source code, re-execute it and reobtain the original results.
+A result is \emph{replicable} if one can create a code that matches the algorithmic descriptions given in the published article and reobtain the original results.} is a cornerstone of science.
+If an experimental result cannot be
 re-obtained by an independent party, it merely becomes, at best, an observation
@@
 problems, in part, because it seems impervious to them.
-Contrary to production software who provide a service geared towards a practical outcome , the motivation behind scientific code is to test an hypothesis.
-While in some instance production software and scientific code are indistinguishable, the reason why they were created is different, and, therefore, so are the criteria to evaluate their success.
+Contrary to production software who provides a service geared towards a practical outcome, the motivation behind scientific code is to test an hypothesis.
+While in some instance production software and scientific code are indistinguishable, the reasons why they were created are different, and, therefore, so are the criteria to evaluate their success.
 A program can
@@
 As explained by \citeauthor{Donoho:2009} \parencite{Donoho:2009}, scientific practice must expect that {\em errors are ubiquitous}, and therefore be robust to them.
-Ensuring reproducibility is a fundamental step toward this: it provides other researchers the means to verify that the code does indeed produce the published results, and to scrutinize the procedures it used to produce them.
+Ensuring reproducibility is a fundamental step toward this: it provides other researchers the means to verify that the code does indeed produce the published results, and to scrutinize the procedures it employed to produce them.
 As demonstrated by \citeauthor{Mesnard:2016} \citep{Mesnard:2016}, reproducibility is hard.\\
@@
 The R$^2$ program uses a random seed but does not keep a trace of it except in the code.
-Should the code change after the production of the results, someone provided with the last version of the code will not be able to know which seed was used to produce the results, and would need to iterate through all possible random seeds, an impossible task in practice\footnote{Here, with $2^{10}$ possibilities for a 10-step random walk, a seed matching the any possible sequence could certainly be found.
-For instance, seed 11235813 matches the results of R$^0$ with Python 2.7.
+Should the code change after the production of the results, someone provided with the last version of the code will not be able to know which seed was used to produce the results, and would need to iterate through all possible random seeds, an impossible task in practice\footnote{Here, with $2^{10}$ possibilities for a 10-step random walk, the seed used or another matching the generated sequence could certainly be found.
+For instance, 436 is the smallest positive integer seed to reproduce the results of R$^0$ with Python 2.7 (1151, 3800, 4717 or 11235813 work as well).
 Such a search becomes intractable for a 100-step walk.}.\\
@@
 Of course, passing the test does not guarantee anything.\\
-
-% reproducibility implies availability
-It is obvious that \emph{reproducibility implies availability}.
-As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request.
-While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable.
-Code \emph{and} input data \emph{and} result data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as \href{https://figshare.com}{Figshare}, \href{https://zenodo.org}{Zenodo}\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment.
-Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.} or a domain specific database, such as \href{https://senselab.med.yale.edu/modeldb/}{ModelDB} for computational neuroscience.
-The codes presented in this article are available in the GitHub repository \href{https://github.com/rougier/random-walk}{github.com/rougier/random-walk} and at \href{https://doi.org/10.5281/zenodo.848217}{doi.org/10.5281/zenodo.848217}.\\
 
@@
 
+% reproducibility implies availability
+It is obvious that \emph{reproducibility implies availability}.
+As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request.
+While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable.
+Code \emph{and} input data \emph{and} result data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as \href{https://figshare.com}{Figshare}, \href{https://zenodo.org}{Zenodo}\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment.
+Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.} or a domain specific database, such as \href{https://senselab.med.yale.edu/modeldb/}{ModelDB} for computational neuroscience.
+The codes presented in this article are available in the GitHub repository \href{https://github.com/rougier/random-walk}{github.com/rougier/random-walk} and at \href{https://doi.org/10.5281/zenodo.848217}{doi.org/10.5281/zenodo.848217}.\\
+
 % Summary
@@
 Dependencies and platforms must be described as precisely and as specifically as possible.
-Parameters values and inputs should accompany the result files.
+Parameters values, the version of the code, and inputs should accompany the result files.
 The data and scripts behind the graphs must be published.
@@
 Despite all this, reusability is often overlooked, and it is not hard to see why.
-Scientists are rarely trained in software engineering, and reusability can represent an expensive endeavour if undertaken as an afterthought, for little tangible short-term benefits, for a codebase that might, after all, see only a single use.
+Scientists are rarely trained in software engineering, and reusability can represent an expensive endeavour if undertaken as an afterthought, for little tangible short-term benefits, for a codebase that might, after all, only see a single use.
 And, in fact, reusability is not as indispensable a requirement as re-runnability, repeatability and reproducibility.
@@
 In the R$^3$ code, the seed and the number of steps are respectively hardcoded and magic.\\
-
-Similarly, code behavior should not be changed by commenting/uncommenting code \citep{Wilson:2017}.
-Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user.
-This improves reproducibility in two ways: it allows those conditions to be recorded as parameters in the result files, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
-With documentation explaining which script or configuration file corresponds to which experiment, reproducing the different figures becomes straightforward.\\
 
@@
 
+Similarly, code behavior should not be changed by commenting/uncommenting code \citep{Wilson:2017}.
+Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user.
+This improves reproducibility in two ways: it allows those conditions to be recorded as parameters in the result files, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
+With documentation explaining which script or configuration file corresponds to which experiment, reproducing the different figures becomes straightforward.\\
+
 Documentation is one of the most potent tools for reusability.
@@
 % conceptual description != replicable description
-This is because most papers strive to communicate the main ideas behind their contribution in terms as simple and as clear as possible, so that the reader may be able to easily understand them and the results that are presented.
+This is because most papers strive to communicate the main ideas behind their contribution in terms as simple and as clear as possible, so that the reader may be able to easily understand them.
 Trying to ensure replicability in the main text adds a myriad of esoteric details that are not conceptually significant and clutter the explanations.
-Therefore, unless the writer dedicates an addendum or a section of the supplementary information for technical details specifically aimed at replicability, the information will not be there because there are incentives not to do so.\\
+Therefore, unless the writer dedicates an addendum or a section of the supplementary information for technical details specifically aimed at replicability, the information will not be there because clarity and concision represent enticing incentives not to do so.\\
 
@@
 In fact, it often relies on it.
-To illustrate this, let us consider what could be the textual description of the random walker (as it would be written in an article describing it):
+To illustrate this, let us consider what could be the description of the random walker, as it would be written in an article describing it:
 %
@@
 This description, while somewhat precise, forgoes---as it is common---the initialization of the variables (here the starting value of the walk: {\tt 0}), and the technical details about which implementation of the RNG is used.\\
+
+% NumPy and Python: differences in the Mersene Twister
+It may look innocuous. After all, the
+\href{https://docs.python.org/3.6/library/random.html}{Python documentation},
+states that \enquote{Python uses the Mersenne Twister as the core generator. It produces
+53-bit precision floats and has a period of 2**19937-1}.
+Someone trying to replicate the work however might choose to use the RNG from the \href{http://www.numpy.org/}{NumPy library}.
+The NumPy library is extensively used in the science community, and it provides an implementation of the Mersene Twister generator too.
+Unfortunately, the way the seed is interpreted by the two implementations is different, yielding different random sequences.\\
+
+Here we are able to replicate exactly\footnote{Striving, as we do here, for a perfect quantitative match may seem unnecessary.
+Yet, in replication projects, in particular in computational research, quantitative comparisons are a simple and effective way to verify that the behavior has been reproduced.
+Moreover, they are particularly helpful to track exactly where the code of a tentative replication fails to reproduce the published results.
+For a discussion about statistical ways to assess replication see the report of the U.S.
+\cite{Schwalbe:2016}.} the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details of the original code (the use of the {\tt random} module of the standard Python library of CPython 3.6.1), or to the code itself.\\
+
+But there are still more subtle problems with the description given above.
+If we look more closely at it, we can realize
+that nothing is said about the specific case of {\tt 0} when generating a step.
+Do we have to consider {\tt 0} to be a positive or a negative step? Without
+further information and without the original code, it is up to the reader to
+decide.
+Likewise, the description is ambiguous regarding the first element of the walk.
+Is the initialization value included (it was not in our codes so far)?
+This slight difference might affect the statistics of short runs.\\
 
@@
 
-% NumPy and Python: differences in the Mersene Twister
-It may look innocuous. After all, the
-\href{https://docs.python.org/3.6/library/random.html}{Python documentation},
-states that \enquote{Python uses the Mersenne Twister as the core generator. It produces
-53-bit precision floats and has a period of 2**19937-1}.
-Someone trying to replicate the work however might choose to use the RNG from the \href{http://www.numpy.org/}{NumPy library}.
-The NumPy library is extensively used in the science community, and it provides an implementation of the Mersene Twister generator too.
-Unfortunately, the way the seed is interpreted by the two implementations is different, yielding different random sequences.\\
-
-Here we are able to replicate exactly\footnote{Striving, as we do here, for a perfect quantitative match may seem unnecessary.
-Yet, in replication projects, quantitative comparisons are a simple and effective way to verify that the behavior has been reproduced.
-Moreover, they are particularly helpful to track exactly where the code of a tentative replication fails to reproduce the published results.} the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details (the use of the {\tt random} module of the standard Python library of CPython 3.6.1), or to the code itself.\\
-
-But there are still more subtle problems with the description given above.
-If we look more closely at it, we can realize
-that nothing is said about the specific case of {\tt 0} when generating a step.
-Do we have to consider {\tt 0} to be a positive or a negative step? Without
-further information and without the original code, it is up to the reader to
-decide.
-Likewise, the description is ambiguous regarding the first element of the walk.
-Is the initialization value included (it was not in our codes so far)?
-This slight difference might affect the statistics of short runs.\\
-
 All these ambiguities in the description of an algorithm pile up; some are inconsequential (the 0 case has null probability), but some may affect the results in important ways.
@@
 This is one of the reasons why the ReScience journal \citep{Rougier:2017} was created (the second author, Nicolas Rougier, is one of the editor-in-chief of ReScience).
-This journal, run entirely by volunteers on GitHub, {\em targets computational research and encourages the explicit replication of already published research, promoting new and open-source implementations in order to ensure that the original research is reproducible}.\\
+This open-access journal, run by volunteers, {\em targets computational research and encourages the explicit replication of already published research, promoting new and open-source implementations in order to ensure that the original research is reproducible}.\\
 
-Code is an integral part of any submission to the ReScience journal.
+Code is an central part of a submission to the ReScience journal.
 During the review process, reviewers run the submitted code, may criticise its quality and its ease-of-use, and verify the reproduciblity of the replication.
@@
 Throughout the evolution of a small random walk example implemented in Python, we illustrated some of the issues that may plague scientific code.
-The code may be correct and of good quality, but still many problems may reduce its contribution to scientific knowledge.
+The code may be correct and of good quality, but still many problems may reduce its contribution to the scientific discourse.
 To make these problems explicit, we articulated five characteristics that a code should possess to be a useful part of a scientific publication: it should be re-runnable, repeatable, reproducible, reusable and replicable.\\
@@
 Running old code on tomorrow’s computer and software stacks may not be possible.
-But recreating the old code’s execution environment may be: to ensure that the long-term re-runnability of a code, its execution environment must be documented.
+But recreating the old code’s execution environment may be: to ensure the long-term re-runnability of a code, its execution environment must be documented.
 For our example, a single comment went a long way to transform the R$^0$ code into the R$^1$ (re-runnable) one.\\
@@
 This is harder to do if each execution of the code produce a different result.
-While for complex parallel workflow this may not be possible, in all instances where it is feasible the code should be repeatable.
+While for complex concurrent workflows this may not be possible, in all instances where it is feasible the code should be repeatable.
 This allows future researchers to examine exactly how a specific result was produced.
@@
 
-Finally, there is the belief that an article should suffice by itself: the descriptions of the algorithms present in the paper should suffice to re-obtain (to replicate) the published results.
+Finally, there is the belief that an article should suffice by itself: the descriptions of the algorithms present in the paper should suffice to reobtain (to replicate) the published results.
 For well-written papers that precisely dissociate conceptually significant aspects from irrelevant implementation details, that may be.
@@
 Every paper is a mistake or a forgotten parameter away from irreproducibility.
-Replication efforts use the paper first, and then the reproducible code that comes along with it whenever the paper falls short of being precise enough to be reimplemented.\\
+Replication efforts use the paper first, and then the reproducible code that comes along with it whenever the paper falls short of being precise enough.\\
 

### 2017-10-19 draft.tex
insisting

@@
 Here, we place ourselves in the context of computational works, where data is produced by a program.
-A result is \emph{reproducible} if one can take the original source code, re-execute it and reobtain the original results.
-A result is \emph{replicable} if one can create a code that matches the algorithmic descriptions given in the published article and reobtain the original results.} is a cornerstone of science.
+In this paper, we call a result \emph{reproducible} if one can take the original source code, re-execute it and reobtain the original results.
+Conversely, a result is \emph{replicable} if one can create a code that matches the algorithmic descriptions given in the published article and reobtain the original results.} is a cornerstone of science.
 If an experimental result cannot be
@@
 
-Code is an central part of a submission to the ReScience journal.
+Code is a key part of a submission to the ReScience journal.
 During the review process, reviewers run the submitted code, may criticise its quality and its ease-of-use, and verify the reproduciblity of the replication.

### 2017-10-19 draft.tex
last spelling corrections

@@
 problems, in part, because it seems impervious to them.
-Contrary to production software who provides a service geared towards a practical outcome, the motivation behind scientific code is to test an hypothesis.
+Contrary to production software who provides a service geared towards a practical outcome, the motivation behind scientific code is to test a hypothesis.
 While in some instance production software and scientific code are indistinguishable, the reasons why they were created are different, and, therefore, so are the criteria to evaluate their success.
@@
 Repeatability is not always possible or easy \citep{Diethelm:2012, Courtes:2015}.
-But for sequential and deterministically parallels programs \citep{Hines:2008, Collange:2015} not depending on analog inputs, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\
+But for sequential and deterministically parallel programs \citep{Hines:2008, Collange:2015} not depending on analog inputs, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\
 
@@
 
+% reproducibility implies availability
+It is obvious that \emph{reproducibility implies availability}.
+As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request.
+While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable.
+Code \emph{and} input data \emph{and} result data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as \href{https://figshare.com}{Figshare}, \href{https://zenodo.org}{Zenodo}\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment.
+Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.} or a domain specific database, such as \href{https://senselab.med.yale.edu/modeldb/}{ModelDB} for computational neuroscience.
+The codes presented in this article are available in the GitHub repository \href{https://github.com/rougier/random-walk}{github.com/rougier/random-walk} and at \href{https://doi.org/10.5281/zenodo.848217}{doi.org/10.5281/zenodo.848217}.\\
+
 \noindent \begin{minipage}[c]{\linewidth}
@@
 
-% reproducibility implies availability
-It is obvious that \emph{reproducibility implies availability}.
-As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request.
-While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable.
-Code \emph{and} input data \emph{and} result data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as \href{https://figshare.com}{Figshare}, \href{https://zenodo.org}{Zenodo}\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment.
-Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.} or a domain specific database, such as \href{https://senselab.med.yale.edu/modeldb/}{ModelDB} for computational neuroscience.
-The codes presented in this article are available in the GitHub repository \href{https://github.com/rougier/random-walk}{github.com/rougier/random-walk} and at \href{https://doi.org/10.5281/zenodo.848217}{doi.org/10.5281/zenodo.848217}.\\
-
 % Summary
@@
 In the R$^3$ code, the seed and the number of steps are respectively hardcoded and magic.\\
+
+Similarly, code behavior should not be changed by commenting/uncommenting code \citep{Wilson:2017}.
+Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user.
+This improves reproducibility in two ways: it allows those conditions to be recorded as parameters in the result files, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
+With documentation explaining which script or configuration file corresponds to which experiment, reproducing the different figures becomes straightforward.\\
 
@@
 
-Similarly, code behavior should not be changed by commenting/uncommenting code \citep{Wilson:2017}.
-Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user.
-This improves reproducibility in two ways: it allows those conditions to be recorded as parameters in the result files, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
-With documentation explaining which script or configuration file corresponds to which experiment, reproducing the different figures becomes straightforward.\\
-
 Documentation is one of the most potent tools for reusability.
@@
 Here, replicating implies writing a new code matching the conceptual description of the article, in order to reobtain the same results.
-Replication affords robustness to the results because, should the original code contain an error, a different codebase creates the possibility that this error will not be repeated, in the same way that replicating an laboratory experiment in a different laboratory can ferret out subtle biases.
+Replication affords robustness to the results because, should the original code contain an error, a different codebase creates the possibility that this error will not be repeated, in the same way that replicating a laboratory experiment in a different laboratory can ferret out subtle biases.
 While every published article should strive for replicability, it is seldom obtained.
@@
 This description, while somewhat precise, forgoes---as it is common---the initialization of the variables (here the starting value of the walk: {\tt 0}), and the technical details about which implementation of the RNG is used.\\
-
-% NumPy and Python: differences in the Mersene Twister
-It may look innocuous. After all, the
-\href{https://docs.python.org/3.6/library/random.html}{Python documentation},
-states that \enquote{Python uses the Mersenne Twister as the core generator. It produces
-53-bit precision floats and has a period of 2**19937-1}.
-Someone trying to replicate the work however might choose to use the RNG from the \href{http://www.numpy.org/}{NumPy library}.
-The NumPy library is extensively used in the science community, and it provides an implementation of the Mersene Twister generator too.
-Unfortunately, the way the seed is interpreted by the two implementations is different, yielding different random sequences.\\
-
-Here we are able to replicate exactly\footnote{Striving, as we do here, for a perfect quantitative match may seem unnecessary.
-Yet, in replication projects, in particular in computational research, quantitative comparisons are a simple and effective way to verify that the behavior has been reproduced.
-Moreover, they are particularly helpful to track exactly where the code of a tentative replication fails to reproduce the published results.
-For a discussion about statistical ways to assess replication see the report of the U.S.
-\cite{Schwalbe:2016}.} the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details of the original code (the use of the {\tt random} module of the standard Python library of CPython 3.6.1), or to the code itself.\\
-
-But there are still more subtle problems with the description given above.
-If we look more closely at it, we can realize
-that nothing is said about the specific case of {\tt 0} when generating a step.
-Do we have to consider {\tt 0} to be a positive or a negative step? Without
-further information and without the original code, it is up to the reader to
-decide.
-Likewise, the description is ambiguous regarding the first element of the walk.
-Is the initialization value included (it was not in our codes so far)?
-This slight difference might affect the statistics of short runs.\\
 
@@
 \end{minipage}
+
+% NumPy and Python: differences in the Mersene Twister
+It may look innocuous. After all, the
+\href{https://docs.python.org/3.6/library/random.html}{Python documentation},
+states that \enquote{Python uses the Mersenne Twister as the core generator. It produces
+53-bit precision floats and has a period of 2**19937-1}.
+Someone trying to replicate the work however might choose to use the RNG from the \href{http://www.numpy.org/}{NumPy library}.
+The NumPy library is extensively used in the science community, and it provides an implementation of the Mersene Twister generator too.
+Unfortunately, the way the seed is interpreted by the two implementations is different, yielding different random sequences.\\
+
+Here we are able to replicate exactly\footnote{Striving, as we do here, for a perfect quantitative match may seem unnecessary.
+Yet, in replication projects, in particular in computational research, quantitative comparisons are a simple and effective way to verify that the behavior has been reproduced.
+Moreover, they are particularly helpful to track exactly where the code of a tentative replication fails to reproduce the published results.
+For a discussion about statistical ways to assess replication see the report of the U.S.
+\cite{Schwalbe:2016}.} the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details of the original code (the use of the {\tt random} module of the standard Python library of CPython 3.6.1), or to the code itself.\\
+
+But there are still more subtle problems with the description given above.
+If we look more closely at it, we can realize
+that nothing is said about the specific case of {\tt 0} when generating a step.
+Do we have to consider {\tt 0} to be a positive or a negative step? Without
+further information and without the original code, it is up to the reader to
+decide.
+Likewise, the description is ambiguous regarding the first element of the walk.
+Is the initialization value included (it was not in our codes so far)?
+This slight difference might affect the statistics of short runs.\\
 

### 2017-10-19 draft.tex
final adjustments to placement of code

@@
 
+\renewcommand{\floatpagefraction}{.9}%
+
 % Hyperref
@@
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 1:}} Random walk (R$^0$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R0.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
+\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 0:}} Random walk (R$^0$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R0.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
 import random
@@
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 2:}} Re-runnable random walk (R$^1$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R1.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
+\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 1:}} Re-runnable random walk (R$^1$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R1.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
 # Tested with Python 3
@@
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 3:}} Re-runnable, repeatable random walk (R$^2$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R2.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
+\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 2:}} Re-runnable, repeatable random walk (R$^2$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R2.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
 # Tested with Python 3
@@
 
-% reproducibility implies availability
-It is obvious that \emph{reproducibility implies availability}.
-As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request.
-While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable.
-Code \emph{and} input data \emph{and} result data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as \href{https://figshare.com}{Figshare}, \href{https://zenodo.org}{Zenodo}\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment.
-Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.} or a domain specific database, such as \href{https://senselab.med.yale.edu/modeldb/}{ModelDB} for computational neuroscience.
-The codes presented in this article are available in the GitHub repository \href{https://github.com/rougier/random-walk}{github.com/rougier/random-walk} and at \href{https://doi.org/10.5281/zenodo.848217}{doi.org/10.5281/zenodo.848217}.\\
-
+\begin{figure}[t!]
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 4:}} Re-runnable, repeatable, reproducible random walk (R$^3$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R3.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
+\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 3:}} Re-runnable, repeatable, reproducible random walk (R$^3$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R3.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
@@
 \end{minipage}\\
+\end{figure}
+
+% reproducibility implies availability
+It is obvious that \emph{reproducibility implies availability}.
+As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request.
+While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable.
+Code \emph{and} input data \emph{and} result data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as \href{https://figshare.com}{Figshare}, \href{https://zenodo.org}{Zenodo}\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment.
+Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.} or a domain specific database, such as \href{https://senselab.med.yale.edu/modeldb/}{ModelDB} for computational neuroscience.
+The codes presented in this article are available in the GitHub repository \href{https://github.com/rougier/random-walk}{github.com/rougier/random-walk} and at \href{https://doi.org/10.5281/zenodo.848217}{doi.org/10.5281/zenodo.848217}.\\
 
@@
 
-For you, first.
-Because the you now and the you in two years are two different persons.
-Details on how to use the code, its limitations, its quirks, may be present to your mind now, but will probably escape you in six months \parencite{Donoho:2009}.
-Here, comments and documentation can make a significant difference.
-Source code reflects the results of the decisions that were made during its creation, but not the reasons behind those decisions.
-In science, where the method and its justification matter as much as the results, those reasons are precious knowledge.
-In that context, a comment on how a given parameter was chosen (optimization, experimental data, educated guess), why a library was chosen over another (conceptual or technical reasons?) is valuable information.\\
-
-% Research team typically experience a high turnover of interns, doctoral and post-doctoral students. Can you reproduce the results of member after leave your team?
-
-Reusability of course directly benefits other researchers from your team and outside of it.
-The easier it is to use your code, the lower the threshold is for other to study, modify and extend it.
-Scientists constantly face the constraint of time: if a model is available, documented, and can be installed, run and understood all in a few hours, it will be preferred over another that would require weeks to reach the same stage.
-A reproducible and reusable code offers a platform both \emph{verifiable} and easy-to-use, fostering the development of derivative works by other researchers on solid foundations.
-Those derivative works contribute to the impact of your original contribution.\\
-
-Having more people examining and using your code also means that potential errors have a higher chance to be caught.
-If people start using your program, they will most likely report bugs or malfunctions they encounter.
-If you're lucky enough, they might even propose either bug fixes or improvements, hence improving the overall quality of your software.
-This process contributes to the long-term reproducibility to the extent people continue to use and maintain the program.\\
-
-Despite all this, reusability is often overlooked, and it is not hard to see why.
-Scientists are rarely trained in software engineering, and reusability can represent an expensive endeavour if undertaken as an afterthought, for little tangible short-term benefits, for a codebase that might, after all, only see a single use.
-And, in fact, reusability is not as indispensable a requirement as re-runnability, repeatability and reproducibility.
-Yet, some simple measures can tremendously increase reusability, and at the same time strengthen reproducibility and re-runnability over the long-term.\\
-
-Avoid hardcoded or magic numbers.
-Magic numbers are numbers present directly in the source code, that do not have a name and therefore can be difficult to interpret semantically.
-Hardcoded values are variables that cannot be changed through a function argument or a parameter configuration file.
-To be modified, they involve editing the code, which is cumbersome and error-prone.
-In the R$^3$ code, the seed and the number of steps are respectively hardcoded and magic.\\
-
-Similarly, code behavior should not be changed by commenting/uncommenting code \citep{Wilson:2017}.
-Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user.
-This improves reproducibility in two ways: it allows those conditions to be recorded as parameters in the result files, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
-With documentation explaining which script or configuration file corresponds to which experiment, reproducing the different figures becomes straightforward.\\
-
+\begin{figure}
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 5:}} Re-runnable, repeatable, reproducible, reusable random walk (R$^4$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R4.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
+\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 4:}} Re-runnable, repeatable, reproducible, reusable random walk (R$^4$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R4.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
@@
 \end{minipage}\\
+\end{figure}
+
+For you, first.
+Because the you now and the you in two years are two different persons.
+Details on how to use the code, its limitations, its quirks, may be present to your mind now, but will probably escape you in six months \parencite{Donoho:2009}.
+Here, comments and documentation can make a significant difference.
+Source code reflects the results of the decisions that were made during its creation, but not the reasons behind those decisions.
+In science, where the method and its justification matter as much as the results, those reasons are precious knowledge.
+In that context, a comment on how a given parameter was chosen (optimization, experimental data, educated guess), why a library was chosen over another (conceptual or technical reasons?) is valuable information.\\
+
+% Research team typically experience a high turnover of interns, doctoral and post-doctoral students. Can you reproduce the results of member after leave your team?
+
+Reusability of course directly benefits other researchers from your team and outside of it.
+The easier it is to use your code, the lower the threshold is for other to study, modify and extend it.
+Scientists constantly face the constraint of time: if a model is available, documented, and can be installed, run and understood all in a few hours, it will be preferred over another that would require weeks to reach the same stage.
+A reproducible and reusable code offers a platform both \emph{verifiable} and easy-to-use, fostering the development of derivative works by other researchers on solid foundations.
+Those derivative works contribute to the impact of your original contribution.\\
+
+Having more people examining and using your code also means that potential errors have a higher chance to be caught.
+If people start using your program, they will most likely report bugs or malfunctions they encounter.
+If you're lucky enough, they might even propose either bug fixes or improvements, hence improving the overall quality of your software.
+This process contributes to the long-term reproducibility to the extent people continue to use and maintain the program.\\
+
+Despite all this, reusability is often overlooked, and it is not hard to see why.
+Scientists are rarely trained in software engineering, and reusability can represent an expensive endeavour if undertaken as an afterthought, for little tangible short-term benefits, for a codebase that might, after all, only see a single use.
+And, in fact, reusability is not as indispensable a requirement as re-runnability, repeatability and reproducibility.
+Yet, some simple measures can tremendously increase reusability, and at the same time strengthen reproducibility and re-runnability over the long-term.\\
+
+Avoid hardcoded or magic numbers.
+Magic numbers are numbers present directly in the source code, that do not have a name and therefore can be difficult to interpret semantically.
+Hardcoded values are variables that cannot be changed through a function argument or a parameter configuration file.
+To be modified, they involve editing the code, which is cumbersome and error-prone.
+In the R$^3$ code, the seed and the number of steps are respectively hardcoded and magic.\\
+
+Similarly, code behavior should not be changed by commenting/uncommenting code \citep{Wilson:2017}.
+Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user.
+This improves reproducibility in two ways: it allows those conditions to be recorded as parameters in the result files, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper.
+With documentation explaining which script or configuration file corresponds to which experiment, reproducing the different figures becomes straightforward.\\
 
@@
 
-A researcher seeking to replicate published results might first consider only the article.
-If she fails to replicate the results, she will consult the original code, and with it be able to pinpoint why her code and the code of the authors differ in behavior.
-Because a mistake on their part?
-Hers?
-Or a difference in a seemingly innocuous implementation detail?
-A fine analysis of why a particular algorithmic description is lacking or ambiguous or why a minor implementation decision is in fact crucial to obtain the published results is of great scientific value.
-Such an analysis can only be done with access to both the article and the code.
-With only the article, the researcher will often be unable to understand why she failed to replicate the results, and will naturally be inclined to only report replication successes.\\
-
-Replicability, therefore, does not negate the necessity of reproducibility.
-In fact, it often relies on it.
-To illustrate this, let us consider what could be the description of the random walker, as it would be written in an article describing it:
-%
-\begin{quotation}
-{\em The model uses the Mersene Twister generator initialized with the seed 1.
-At each iteration, a uniform number between -1 (included) and +1 (excluded) is drawn and the sign of the result is used for generating a positive or negative step.}
-\end{quotation}
-%
-This description, while somewhat precise, forgoes---as it is common---the initialization of the variables (here the starting value of the walk: {\tt 0}), and the technical details about which implementation of the RNG is used.\\
-
+\begin{figure}[b!]
 \noindent \begin{minipage}[c]{\linewidth}
-\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 6:}} Replicated random walk (R$^5$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R5.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
+\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 5:}} Replicated random walk (R$^5$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R5.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
 # Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
@@
 \end{minipage}
+\end{figure}
+
+A researcher seeking to replicate published results might first consider only the article.
+If she fails to replicate the results, she will consult the original code, and with it be able to pinpoint why her code and the code of the authors differ in behavior.
+Because a mistake on their part?
+Hers?
+Or a difference in a seemingly innocuous implementation detail?
+A fine analysis of why a particular algorithmic description is lacking or ambiguous or why a minor implementation decision is in fact crucial to obtain the published results is of great scientific value.
+Such an analysis can only be done with access to both the article and the code.
+With only the article, the researcher will often be unable to understand why she failed to replicate the results, and will naturally be inclined to only report replication successes.\\
+
+Replicability, therefore, does not negate the necessity of reproducibility.
+In fact, it often relies on it.
+To illustrate this, let us consider what could be the description of the random walker, as it would be written in an article describing it:
+%
+\begin{quotation}
+{\em The model uses the Mersene Twister generator initialized with the seed 1.
+At each iteration, a uniform number between -1 (included) and +1 (excluded) is drawn and the sign of the result is used for generating a positive or negative step.}
+\end{quotation}
+%
+This description, while somewhat precise, forgoes---as it is common---the initialization of the variables (here the starting value of the walk: {\tt 0}), and the technical details about which implementation of the RNG is used.\\
 

### 2017-10-19 draft.tex
add space in intro
another spelling mistake!

@@
 
+\vspace{1.5em}
+
 In the code above, the {\tt \href{https://docs.python.org/3.6/library/random.html#random.choice}{random.choice}} function randomly returns either +1 or -1.
@@
 This not the case: it uses the print {\em instruction} and the {\tt xrange} operator, both specific to Python 2.
-The print {\em instruction}, available in Python 2 (a version still widely used; support is scheduled to stop in 2020), has been deprecated in Python 3 (first released in 2008, almost a decade ago) in favor or a  print {\em function}, while the {\tt xrange} operator has been replaced by the {\tt range} operator in Python 3.
+The print {\em instruction}, available in Python 2 (a version still widely used; support is scheduled to stop in 2020), has been deprecated in Python 3 (first released in 2008, almost a decade ago) in favor of a  print {\em function}, while the {\tt xrange} operator has been replaced by the {\tt range} operator in Python 3.
 In order to try to future-proof the code a bit, we might as well target Python 3, as is done in the R$^1$ version.

### 2017-12-11 draft.tex
update codes

@@
 \begin{code}{Output}
--1, 0, -1, 0, -1, 0, -1, 0, 1, 2 # with the steps being -1, +1, -1, +1, -1, +1, -1, +1, +1, +1
+-1, 0, -1, 0, -1, 0, -1, 0, 1, 2
+# with the steps being -1,+1,-1,+1,-1,+1,-1,+1,+1,+1
 \end{code}
@@
 \begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 3:}} Re-runnable, repeatable, reproducible random walk (R$^3$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R3.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
-# Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
+# Copyright (c) 2017 N.P. Rougier and F.C.Y. Benureau
 # Release under the BSD 2-clause license
-# Tested with CPython 3.6.2 / macOS 10.12.6 / 64 bits architecture
+# Tested with 64-bit CPython 3.6.2 / macOS 10.12.6
 import sys, subprocess, datetime, random
 
-def generate_walk():
+def compute_walk():
     x = 0
@@
 
-# If repository is dirty, don't run anything
-if subprocess.call(('git', 'diff-index', '--quiet', 'HEAD')):
-    print('Repository is dirty, please commit first')
+# If repository is dirty, don"t run anything
+if subprocess.call(("git", "diff-index",
+                    "--quiet", "HEAD")):
+    print("Repository is dirty, please commit first")
     sys.exit(1)
@@
 # Get git hash if any
-revision = subprocess.check_output(('git', 'rev-parse', 'HEAD'))
+hash_cmd = ("git", "rev-parse", "HEAD")
+revision = subprocess.check_output(hash_cmd)
 
@@
 random.seed(42)
-assert generate_walk() == [1, 0, -1, -2, -1, 0, 1, 0, -1, -2]
+assert compute_walk() == [1,0,-1,-2,-1,0,1,0,-1,-2]
 
@@
 random.seed(seed)
-walk = generate_walk()
+walk = compute_walk()
 
@@
 print(walk)
-results = {'data'       : walk,
-            'seed'      : seed,
-            'timestamp': str(datetime.datetime.utcnow()),
-            'revision' : revision,
-            'system'    : sys.version}
-with open('results-R3.txt', 'w') as fd:
+results = {
+    "data"     : walk,
+    "seed"     : seed,
+    "timestamp": str(datetime.datetime.utcnow()),
+    "revision" : revision,
+    "system"   : sys.version}
+with open("results-R3.txt", "w") as fd:
     fd.write(str(results))
@@
 \begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 4:}} Re-runnable, repeatable, reproducible, reusable random walk (R$^4$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R4.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
-# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
+# Copyright (c) 2017 N.P. Rougier and F.C.Y. Benureau
 # Release under the BSD 2-clause license
-# Tested with CPython 3.6.2 / macOS 10.12.6 / 64 bits architecture
+# Tested with 64-bit CPython 3.6.2 / macOS 10.12.6
 import sys, subprocess, datetime, random
 
-def generate_walk(count, x0=0, step=1, seed=0):
-    """ Random walk
+def compute_walk(count, x0=0, step=1, seed=0):
+    """Random walk
        count: number of steps
@@
        step : step size (default 1)
-        seed : seed for the initialization of the random generator (default 0)
+       seed : seed for the initialization of the
+	     random generator (default 0)
     """
@@
 
-def generate_results(count, x0=0, step=1, seed=0):
-    """Compute a walk and return it alongside its context"""
-    # If repository is dirty, don't do anything
-    if subprocess.call(('git', 'diff-index', '--quiet', 'HEAD')):
-        print('Repository is dirty, please commit first')
+def compute_results(count, x0=0, step=1, seed=0):
+    """Compute a walk and return it with context"""
+    # If repository is dirty, don"t do anything
+    if subprocess.call(("git", "diff-index",
+                        "--quiet", "HEAD")):
+        print("Repository is dirty, please commit")
         sys.exit(1)
@@
     # Get git hash if any
-    revision = subprocess.check_output(('git', 'rev-parse', 'HEAD'))
+    hash_cmd = ("git", "rev-parse", "HEAD")
+    revision = subprocess.check_output(hash_cmd)
 
     # Compute results
-    walk = generate_walk(count=count, x0=x0, step=step, seed=seed)
-    return {'data'      : walk,
-            'parameters': {'x0':x0, 'step':step, 'count':count, 'seed':seed},
-            'timestamp' : str(datetime.datetime.utcnow()),
-            'revision'  : revision,
-            'system'    : sys.version}
+    walk = compute_walk(count=count, x0=x0,
+                        step=step, seed=seed)
+    return {
+        "data"      : walk,
+        "parameters": {"count": count, "x0": x0,
+                       "step": step, "seed": seed},
+        "timestamp" : str(datetime.datetime.utcnow()),
+        "revision"  : revision,
+        "system"    : sys.version}
 
-if __name__ == '__main__':
-    # Unit test checking reproducibility (will fail with Python<=3.2)
-    assert generate_walk(10, 0, 1, 42) == [1, 0, -1, -2, -1, 0, 1, 0, -1, -2]
+if __name__ == "__main__":
+    # Unit test checking reproducibility
+    # (will fail with Python<=3.2)
+    assert (compute_walk(10, 0, 1, 42) ==
+	        [1,0,-1,-2,-1,0,1,0,-1,-2])
 
@@
     count, x0, seed = 10, 0, 1
-    results = generate_results(count, x0=x0, seed=seed)
+    results = compute_results(count, x0=x0, seed=seed)
 
     # Save & display results
-    with open('results-R4.txt', 'w') as fd:
+    with open("results-R4.txt", "w") as fd:
         fd.write(str(results))
-    print(results['data'])
+    print(results["data"])
 \end{code}
@@
 \begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 5:}} Replicated random walk (R$^5$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R5.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
-# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
+# Copyright (c) 2017 N.P. Rougier and F.C.Y. Benureau
 # Release under the BSD 2-clause license
-# Tested with CPython 3.6.2 / NumPy 1.12.0 / macOS 10.12.6 / 64 bits architecture
+# 64-bit CPython 3.6.2 / NumPy 1.12.0 / macOS 10.12.6
 import random
@@
 def _rng(seed):
-    """ Return a numpy random number generator initialized with seed
-        as it would be with python random generator.
+    """Return a numpy random number generator
+       initialized with seed as it would be with
+       a python random generator.
     """
@@
     state = rng.get_state()
-    rng.set_state((state[0], keys[:-1], state[2], state[3], state[4]))
+    rng.set_state((state[0], keys[:-1], state[2],
+		           state[3], state[4]))
     return rng
@@
 def walk(n, seed):
-    """ Random walk for n steps """
-
+    """Random walk for n steps"""
     rng = _rng(seed)
-    steps = 2*(rng.uniform(-1,+1,n) > 0) - 1
+    steps = 2 * (rng.uniform(-1, +1, n) > 0) - 1
     return steps.cumsum().tolist()
 
-if __name__ == '__main__':
+if __name__ == "__main__":
     # Unit test
-    assert walk(n=10, seed=42) == [1, 0, -1, -2, -1, 0, 1, 0, -1, -2]
+    assert (walk(n=10, seed=42) ==
+            [1,0,-1,-2,-1,0,1,0,-1,-2])
 
-    # Random walk for 10 steps, initialization with seed=1
+    # Random walk for 10 steps, with seed=1
     seed = 1
@@
     # Save & display results
-    results = {'data': path, 'seed': seed}
+    results = {"data": path, "seed": seed}
     with open("results-R5.txt", "w") as fd:

### 2017-12-11 codes/r3.py

@@ -1,43 +1,0 @@
-# Random walk (R3: reproducible)
-# Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
-# Release under the BSD 2-clause license
-# Tested with CPython 3.6.2 / macOS 10.12.6 / 64 bits architecture
-import sys, subprocess, datetime, random
-
-def generate_walk():
-    walk = []
-    x = 0
-    for i in range(10):
-        if random.uniform(-1, +1) > 0:
-            x += 1
-        else:
-            x -= 1
-        walk.append(x)
-    return walk
-
-# If repository is dirty, don't run anything
-if subprocess.call(('git', 'diff-index', '--quiet', 'HEAD')):
-    print('Repository is dirty, please commit first')
-    sys.exit(1)
-
-# Get git hash if any
-revision = subprocess.check_output(('git', 'rev-parse', 'HEAD'))
-
-# Unit test
-random.seed(42)
-assert generate_walk() == [1, 0, -1, -2, -1, 0, 1, 0, -1, -2]
-
-# Random walk for 10 steps
-seed = 1
-random.seed(seed)
-walk = generate_walk()
-
-# Display & save results
-print(walk)
-results = {'data'     : walk,
-           'seed'     : seed,
-           'timestamp': str(datetime.datetime.utcnow()),
-           'revision' : revision,
-           'system'   : sys.version}
-with open('results-R3.txt', 'w') as fd:
-    fd.write(str(results))

### 2017-12-11 codes/r4.py

@@ -1,54 +1,0 @@
-# Random walk (R4: re-usable)
-# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
-# Release under the BSD 2-clause license
-# Tested with CPython 3.6.2 / macOS 10.12.6 / 64 bits architecture
-import sys, subprocess, datetime, random
-
-def generate_walk(count, x0=0, step=1, seed=0):
-    """ Random walk
-        count: number of steps
-        x0   : initial position (default 0)
-        step : step size (default 1)
-        seed : seed for the initialization of the random generator (default 0)
-    """
-    random.seed(seed)
-    x = x0
-    walk = []
-    for i in range(count):
-        if random.uniform(-1, +1) > 0:
-            x += 1
-        else:
-            x -= 1
-        walk.append(x)
-    return walk
-
-def generate_results(count, x0=0, step=1, seed=0):
-    """Compute a walk and return it alongside its context"""
-    # If repository is dirty, don't do anything
-    if subprocess.call(('git', 'diff-index', '--quiet', 'HEAD')):
-        print('Repository is dirty, please commit first')
-        sys.exit(1)
-
-    # Get git hash if any
-    revision = subprocess.check_output(('git', 'rev-parse', 'HEAD'))
-
-    # Compute results
-    walk = generate_walk(count=count, x0=x0, step=step, seed=seed)
-    return {'data'      : walk,
-            'parameters': {'x0':x0, 'step':step, 'count':count, 'seed':seed},
-            'timestamp' : str(datetime.datetime.utcnow()),
-            'revision'  : revision,
-            'system'    : sys.version}
-
-if __name__ == '__main__':
-    # Unit test checking reproducibility (will fail with Python<=3.2)
-    assert generate_walk(10, 0, 1, 42) == [1, 0, -1, -2, -1, 0, 1, 0, -1, -2]
-
-    # Simulation parameters
-    count, x0, seed = 10, 0, 1
-    results = generate_results(count, x0=x0, seed=seed)
-
-    # Save & display results
-    with open('results-R4.txt', 'w') as fd:
-        fd.write(str(results))
-    print(results['data'])

### 2017-12-11 codes/r5.py

@@ -1,39 +1,0 @@
-# Random walk (R5: replicable)
-# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
-# Release under the BSD 2-clause license
-# Tested with CPython 3.6.2 / NumPy 1.12.0 / macOS 10.12.6 / 64 bits architecture
-import random
-import numpy as np
-
-def _rng(seed):
-    """ Return a numpy random number generator initialized with seed
-        as it would be with python random generator.
-    """
-    rng = random.Random()
-    rng.seed(seed)
-    _, keys, _ = rng.getstate()
-    rng = np.random.RandomState()
-    state = rng.get_state()
-    rng.set_state((state[0], keys[:-1], state[2], state[3], state[4]))
-    return rng
-
-def walk(n, seed):
-    """ Random walk for n steps """
-
-    rng = _rng(seed)
-    steps = 2*(rng.uniform(-1,+1,n) > 0) - 1
-    return steps.cumsum().tolist()
-
-if __name__ == '__main__':
-    # Unit test
-    assert walk(n=10, seed=42) == [1, 0, -1, -2, -1, 0, 1, 0, -1, -2]
-
-    # Random walk for 10 steps, initialization with seed=1
-    seed = 1
-    path = walk(n=10, seed=seed)
-
-    # Save & display results
-    results = {'data': path, 'seed': seed}
-    with open("results-R5.txt", "w") as fd:
-        fd.write(str(results))
-    print(path)

### 2017-12-11 codes/random-walk-R0.py

@@ -1,5 +1,6 @@
 # Random walk (R0: worst you can do)
 import random
+
 x = 0
 for i in xrange(10):
     step = random.choice([-1,+1])

### 2017-12-11 codes/random-walk-R0_output.py

@@ -1,0 +1,2 @@
+-1, 0, -1, 0, -1, 0, -1, 0, 1, 2 
+# with the steps being -1,+1,-1,+1,-1,+1,-1,+1,+1,+1

### 2017-12-11 codes/random-walk-R3.py

@@ -1,0 +1,45 @@
+# Copyright (c) 2017 N.P. Rougier and F.C.Y. Benureau
+# Release under the BSD 2-clause license
+# Tested with 64-bit CPython 3.6.2 / macOS 10.12.6
+import sys, subprocess, datetime, random
+
+def compute_walk():
+    x = 0
+    walk = []
+    for i in range(10):
+        if random.uniform(-1, +1) > 0:
+            x += 1
+        else:
+            x -= 1
+        walk.append(x)
+    return walk
+
+# If repository is dirty, don"t run anything
+if subprocess.call(("git", "diff-index",
+                    "--quiet", "HEAD")):
+    print("Repository is dirty, please commit first")
+    sys.exit(1)
+
+# Get git hash if any
+hash_cmd = ("git", "rev-parse", "HEAD")
+revision = subprocess.check_output(hash_cmd)
+
+# Unit test
+random.seed(42)
+assert compute_walk() == [1,0,-1,-2,-1,0,1,0,-1,-2]
+
+# Random walk for 10 steps
+seed = 1
+random.seed(seed)
+walk = compute_walk()
+
+# Display & save results
+print(walk)
+results = {
+    "data"     : walk,
+    "seed"     : seed,
+    "timestamp": str(datetime.datetime.utcnow()),
+    "revision" : revision,
+    "system"   : sys.version}
+with open("results-R3.txt", "w") as fd:
+    fd.write(str(results))

### 2017-12-11 codes/random-walk-R4.py

@@ -1,0 +1,61 @@
+# Copyright (c) 2017 N.P. Rougier and F.C.Y. Benureau
+# Release under the BSD 2-clause license
+# Tested with 64-bit CPython 3.6.2 / macOS 10.12.6
+import sys, subprocess, datetime, random
+
+def compute_walk(count, x0=0, step=1, seed=0):
+    """Random walk
+       count: number of steps
+       x0   : initial position (default 0)
+       step : step size (default 1)
+       seed : seed for the initialization of the
+	     random generator (default 0)
+    """
+    random.seed(seed)
+    x = x0
+    walk = []
+    for i in range(count):
+        if random.uniform(-1, +1) > 0:
+            x += 1
+        else:
+            x -= 1
+        walk.append(x)
+    return walk
+
+def compute_results(count, x0=0, step=1, seed=0):
+    """Compute a walk and return it with context"""
+    # If repository is dirty, don"t do anything
+    if subprocess.call(("git", "diff-index",
+                        "--quiet", "HEAD")):
+        print("Repository is dirty, please commit")
+        sys.exit(1)
+
+    # Get git hash if any
+    hash_cmd = ("git", "rev-parse", "HEAD")
+    revision = subprocess.check_output(hash_cmd)
+
+    # Compute results
+    walk = compute_walk(count=count, x0=x0,
+                        step=step, seed=seed)
+    return {
+        "data"      : walk,
+        "parameters": {"count": count, "x0": x0,
+                       "step": step, "seed": seed},
+        "timestamp" : str(datetime.datetime.utcnow()),
+        "revision"  : revision,
+        "system"    : sys.version}
+
+if __name__ == "__main__":
+    # Unit test checking reproducibility
+    # (will fail with Python<=3.2)
+    assert (compute_walk(10, 0, 1, 42) ==
+	        [1,0,-1,-2,-1,0,1,0,-1,-2])
+
+    # Simulation parameters
+    count, x0, seed = 10, 0, 1
+    results = compute_results(count, x0=x0, seed=seed)
+
+    # Save & display results
+    with open("results-R4.txt", "w") as fd:
+        fd.write(str(results))
+    print(results["data"])

### 2017-12-11 codes/random-walk-R5.py

@@ -1,0 +1,40 @@
+# Copyright (c) 2017 N.P. Rougier and F.C.Y. Benureau
+# Release under the BSD 2-clause license
+# 64-bit CPython 3.6.2 / NumPy 1.12.0 / macOS 10.12.6
+import random
+import numpy as np
+
+def _rng(seed):
+    """Return a numpy random number generator
+       initialized with seed as it would be with
+       a python random generator.
+    """
+    rng = random.Random()
+    rng.seed(seed)
+    _, keys, _ = rng.getstate()
+    rng = np.random.RandomState()
+    state = rng.get_state()
+    rng.set_state((state[0], keys[:-1], state[2],
+		           state[3], state[4]))
+    return rng
+
+def walk(n, seed):
+    """Random walk for n steps"""
+    rng = _rng(seed)
+    steps = 2 * (rng.uniform(-1, +1, n) > 0) - 1
+    return steps.cumsum().tolist()
+
+if __name__ == "__main__":
+    # Unit test
+    assert (walk(n=10, seed=42) ==
+            [1,0,-1,-2,-1,0,1,0,-1,-2])
+
+    # Random walk for 10 steps, with seed=1
+    seed = 1
+    path = walk(n=10, seed=seed)
+
+    # Save & display results
+    results = {"data": path, "seed": seed}
+    with open("results-R5.txt", "w") as fd:
+        fd.write(str(results))
+    print(path)

### 2017-12-11 draft.tex
Correct some mistakes, better whitespaces in code

@@
   commentstyle=\color{gray},
+  keepspaces=true,
   columns=flexible,
@@
       $^{*}$Corresponding author:
-      \href{mailto:fabien.benureau@gmail.com}{fabien.benureau@gmail.com}
+      \href{mailto:fabien@benureau.com}{fabien@benureau.com}
     \end{center}
@@
 
-Replicability\footnote{\emph{Reproducibility} and \emph{replicability} are employed differently by different authors and in different domains (see for instance the report from the U.S.
-\cite{Schwalbe:2016}).
+Replicability\footnote{\emph{Reproducibility} and \emph{replicability} are employed differently by different authors and in different domains (see for instance the report from the  \cite{Schwalbe:2016}).
 Here, we place ourselves in the context of computational works, where data is produced by a program.
-In this paper, we call a result \emph{reproducible} if one can take the original source code, re-execute it and reobtain the original results.
-Conversely, a result is \emph{replicable} if one can create a code that matches the algorithmic descriptions given in the published article and reobtain the original results.} is a cornerstone of science.
+In this paper, we call a result \emph{reproducible} if one can take the original source code, re-execute it and reobtain the original result.
+Conversely, a result is \emph{replicable} if one can create a code that matches the algorithmic descriptions given in the published article and reobtain the original result.} is a cornerstone of science.
 If an experimental result cannot be
@@
 Rather, it depends on the context, and becomes increasingly difficult as the code ages.
-Therefore, to be and remain re-runnable on other researchers' computers, a re-runnable code should describe---with enough details to be recreated---an execution environment in which it is executable.
+Therefore, to be and remain re-runnable on the computers of other researchers, a re-runnable code should describe---with enough details to be recreated---an execution environment in which it is executable.
 As shown by \citep{Collberg:2016}, this is far from being either obvious or easy.\\
@@
 % precise execution environment recording
-Because any dependency of a program---to the most basic one, the language itself---can change its behavior from one version to another, executability (R$^1$) and determinism (R$^2$) are necessary but not sufficient for reproducibility.
+Because any dependency of a program---to the most basic one, the language itself---can change its behavior from one version to the other, executability (R$^1$) and determinism (R$^2$) are necessary but not sufficient for reproducibility.
 The exact execution environment used to produce the results must also be specified---rather than the broadest set of environments where the code can be effectively run.
@@
 This also allows, if some errors are found and corrected after some results have been obtained, to identify which ones should be recomputed.
-In R$^3$, the code records the git revision, and whether the repository holds uncommitted changes when the computation starts.\\
+In R$^3$, the code records the git revision, and prevents execution if the repository holds uncommitted changes when the computation starts.\\
 
 % scripts to generates figures & data behind graphs
-Published results should obviously come from version of the code where every change and every file has been committed.
+Published results should obviously come from versions of the code where every change and every file has been committed.
 This includes pre-processing, post-processing and plotting code.
@@
     """Compute a walk and return it with context"""
-    # If repository is dirty, don"t do anything
-    if subprocess.call(("git", "diff-index",
-                        "--quiet", "HEAD")):
+    # If repository is dirty, don't do anything
+    if subprocess.call(("git", "diff-index", "--quiet", "HEAD")):
         print("Repository is dirty, please commit")
@@
     # Compute results
-    walk = compute_walk(count=count, x0=x0,
-                        step=step, seed=seed)
-    return {
-        "data"      : walk,
+    walk = compute_walk(count=count, x0=x0, step=step, seed=seed)
+    return {"data"      : walk,
             "parameters": {"count": count, "x0": x0,
@@
 if __name__ == "__main__":
-    # Unit test checking reproducibility
-    # (will fail with Python<=3.2)
-    assert (compute_walk(10, 0, 1, 42) ==
-	        [1,0,-1,-2,-1,0,1,0,-1,-2])
+    # Unit test checking reproducibility (will fail with Python<=3.2)
+    assert compute_walk(10, 0, 1, 42) == [1,0,-1,-2,-1,0,1,0,-1,-2]
 
@@
 To complement the R$^4$ version provided here, we provide an example repository of a re-runnable, repeatable, reproducible and reusable random walk code.
-The repository is available on GitHub \href{https://github.com/benureau/r5}{github.com/benureau/r5} and here \href{https://doi.org/10.5281/zenodo.848284}{doi.org/10.5281/zenodo.848284}.
+The repository is available on GitHub \href{https://github.com/benureau/r5}{github.com/benureau/r5} and at \href{https://doi.org/10.5281/zenodo.848284}{doi.org/10.5281/zenodo.848284}.
 
@@
 Moreover, they are particularly helpful to track exactly where the code of a tentative replication fails to reproduce the published results.
-For a discussion about statistical ways to assess replication see the report of the U.S.
-\cite{Schwalbe:2016}.} the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details of the original code (the use of the {\tt random} module of the standard Python library of CPython 3.6.1), or to the code itself.\\
+For a discussion about statistical ways to assess replication see the report of the \cite{Schwalbe:2016}.} the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details of the original code (the use of the {\tt random} module of the standard Python library of CPython 3.6.1), or to the code itself.\\
 
@@
 Throughout the evolution of a small random walk example implemented in Python, we illustrated some of the issues that may plague scientific code.
-The code may be correct and of good quality, but still many problems may reduce its contribution to the scientific discourse.
+The code may be correct and of good quality and still possess many problems that reduce its contribution to the scientific discourse.
 To make these problems explicit, we articulated five characteristics that a code should possess to be a useful part of a scientific publication: it should be re-runnable, repeatable, reproducible, reusable and replicable.\\
@@
 Science assumes that errors can crop up everywhere.
-Every paper is a mistake or a forgotten parameter away from irreproducibility.
+Every paper is a mistake or a forgotten parameter away from irreplicability.
 Replication efforts use the paper first, and then the reproducible code that comes along with it whenever the paper falls short of being precise enough.\\
@@
 checked by reviewers and publishers when code is part of a work worth being
-published. But it's hardly the case today.\\
+published. It's hardly the case today.\\
 
